### 2021-08-26 paper/paper.tex
add paper template, switch to bibtex

\documentclass{aci}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\usepackage{cite}
\def\typeofarticle{Article type} \def\currentvolume{1} \def\currentissue{1}
\def\currentyear{2021} \def\currentmonth{} \def\ppages{1--30} \def\DOI{to
appear} \def\Received{June 2022} \def\Accepted{July 2022} \def\Published{October
2022 }

\newcommand{\ep}{\varepsilon}
\newcommand{\eps}[1]{{#1}_{\varepsilon}}

%\numberwithin{equation}{section}
\DeclareMathOperator*{\essinf}{ess\,inf}

\begin{document}

\title{Title of the paper}

\author{%
  First name Last name\affil{1}, First name Last name\affil{2} and First name
  Last name\affil{3,}\corrauth}

% \shortauthors is used in copyright information in the end of the paper
\shortauthors{the Author(s)}

\address{%
  \addr{\affilnum{1}}{College of Computing and Software Engineering, Kennesaw
    State University, 1000 Chastain Road, Kennesaw, GA 30144, USA; email}
  \addr{\affilnum{2}}{Affiliation email} \addr{\affilnum{3}}{Affiliation email}}
% corresponding author
\corraddr{Email address of corresponding author; Tel: +1-111-111-1111; Fax:\\
  +1-111-111-1111.}

\editor{ First-name Last-name}

\begin{abstract}
  An abstract is normally of 150 to 250 words. List 4 to 8 keywords all with
  lowercase (except special cases) for both highlighting the focuses of your
  study and indexing/searching in databases. Article type can be one of the
  following: regular article, case study, express letter, review or overview,
  survey, opinion, perspective, communication, lecture notes, tutorial notes,
  and editorial (special issue only).

\end{abstract}

\keywords{(4 to 8 keywords)}

\maketitle

\section{Introduction}

Submission of a manuscript implies that the work described has not been
published before, and not been under consideration for publishing anywhere else.
Your submission also means that the manuscript has been approved by all
co-authors with appropriate permission if required. The publisher will not be
held legally responsible should any dispute be raised for any reason. AIMS-ACI
is an OA publication published quarterly by AIMS Press in February, May, August,
and November. Publishing is currently free of charge. Readers can access all
published articles free of charge through the journal website
https://aimspress.com/journal/aci

\section{Layout and fonts}
The page size is A4 ($8.5 \times 11.0$ in or $21.59 \times 27.94$ cm). The
margin is set to moderate. Font for all text, except symbols and formulae,
should use Times New Roman. Article title is bold with a font size 16 and
aligned to the left. Authors are bold with a font size 12. Using Times New Roman
size 12 for the main text, including figure captions and table titles. However,
text size for contents in a table can be chosen from 8 to 12 to fit the table
within the width of the page.

\section{Headline and sub-headline}
The first-level headline is numbered and bold with a font size 12, and has a
space of 12 $pt$ before and after the headline. Each paragraph begins with an
indentation of 0.63 $cm$ (the same by using the Tab key once) in the first line.
A paragraph does not have a space before and after that paragraph.

\subsection{The second-level headline}
The second-level headline is numbered with a font size 12. The second-level
headline has a default space of 6 $pt$ before and after it.

\subsubsection{The third-level headline (if necessary)}
If really necessary, the third-level headline should be numbered and italicized
with a font size 12, indented by 0.63 $cm$ (the same by using the Tab key once).
The third-level headline has a default space of 6 $pt$ before and after it. Any
lower headline should be avoided.

\subsection{Footnotes}
Although not encouraged, footnotes can be used to provide additional information
for an informal reference that cannot be included in the reference list. A
footnote should be numbered and cannot contain any figure or table. Use font
size 8 for footnotes.

\section{Figures or images, captions, tables, table titles, and formulae}
\subsection{Figure/image and caption}
Leave a space of one line before and after a figure or an image, i.e., one line
between the main text and the top of the figure or image, and one line between
the bottom of the figure or image and the caption. The caption has a default
space of 12 $pt$ after it so the main text can continue below the caption. If no
text follows the figure caption, do not leave any space between the caption and
the next headline (the headline has a default space).

\begin{figure}[ht]
  \centering
  \vskip 0.2in
  %\includegraphics[width=0.85\textwidth]{{logo_new.pdf}}
  \vspace {12pt}
  \caption{ Insert the figure/image here}
  \label{fig:res}
\end{figure}

Figures should be numbered consecutively in the text. A figure can be referred
explicitly in the text as Figure~\ref{fig:res}, or implicitly at the end of a
sentence in parentheses (Fig. 1).

Use black and white graphic for line drawings. All lines should be at least 0.1
mm (0.3 $pt$) wide. Line drawings or scanned line drawings should have a
resolution at least 600 dpi. Images with a large volume should be properly
compressed within 1 MB each. Scanned images should be properly edited to balance
the volume size and clarity of the image.

Figure captions begin with \textbf{Figure x.} in bold, followed by the text. Do
not add a full stop at the end of the caption.

\subsection{Tables and table titles}
\begin{table}[H]
  \begin{center}
    \caption{Publication schedule for AIMS-ACI in 2022}
    \begin{tabular}{lll} \hline
      Issue        & Publishing date   & Submission deadline \\\hline
      First issue  & February 25, 2022 & January 15, 2022    \\
      Second issue & May 25, 2022      & April 15, 2022      \\
      Third issue  & August 25, 2022   & July 15, 2022       \\
      Fourth issue & November 25, 2022 & October 15, 2022    \\\hline
    \end{tabular}
    \label{table}
  \end{center}
\end{table}

Table title should be placed above the table, numbered consecutively, and
referred in the text like Table~\ref{table} or (Table~\ref{table}). There is no
space between the table and table title. Leave a space of one line above the
table title and one line below the bottom of the table. If a new first-level
headline follows the table immediately, do not leave the space below the table.

The font of the table title is size 12 but the font of the table contents can be
sizes 8 to 12 depending on the best fit for the table. Table title begins with
\textbf{Table x}. in bold, followed by the text. Do not add a full stop at the
end of the table title.

\subsection{Mathematical formulae}
Mathematical formulae should be numbered consecutively by placing the number in
parentheses like (1), one TAB space away from the right line. Mathematical
formulae should be typed using MS Word Equation or better using \emph{MathType}.
Use size 10--12 for the main characters and size 6--8 for subscripts or
superscripts. Leave a space of 6 $pt$ between two formula in separate lines.
Some examples are shown below.

\begin{equation}
  \lim_{x \mapsto a^+ }f\left ( x \right )= \pm \infty \qquad \text{or} \qquad \lim_{x \mapsto a^- }f\left ( x \right )= \pm \infty
  \label{equation1}
\end{equation}

\begin{equation}
  \lim_{x \mapsto \infty }f\left ( x \right )=  a \qquad \text{or} \qquad \lim_{x \mapsto -\infty}f\left ( x \right )=  b
  \label{equation2}
\end{equation}

To cite a formula in the text, use the number like Equation~(\ref{equation1}),
or Eq.~(\ref{equation2}), or alike.

\section{Submission, review, and decision}
In the first year, submission, review and decisions of manuscripts is conducted
through email communications. Please send your manuscript to one of the relevant
Editors-in-Chief by their email available on the journal Website. With the
increase in submissions from 2022, we would use the AIMS Editorial System to
handle the process of submission, review, and decisions.

Once a submission is received, the receiving Editor will pre-screen the article
for its suitability and quality for the journal. After passing this
pre-screening, the Editor may act as the “Academic Editor” or assign this
manuscript to another Editor from the Editorial Board as the “Academic Editor”
to handle the review of this paper, communicate with the authors on revisions,
and inform the final decision after consultation with the relevant
Editor-in-Chief.

AIMS-ACI holds a high standard in peer review. A submitted manuscript should be
reviewed by at least three independent reviewers who are scholars and
knowledgeable of the area of research. Standard review forms are available for
reviewers but they could choose own appropriate way of reporting to the Academic
Editor.

\section{Reference style, citation, and cross-reference}
Test citation \cite{Jensen1996}.

\section*{Acknowledgments}
We would like to thank the constructive feedback provided by the reviewers.

\bibliographystyle{unsrt}
\bibliography{paper}

\end{document}

### 2021-08-26 cuticle_analysis/core/__init__.py
fix data_spread script for rough/smooth

@@ -33,8 +33,6 @@ def init():
 
 
 def start_app():
-    init()
-
     args = get_args()
 
     if args.download_dataset:

### 2021-08-26 cuticle_analysis/datasets/sub.py

@@ -36,6 +36,16 @@ class SubDataset(Dataset):
                          rebuild=rebuild,
                          save=save)
 
+    def get_label_name(self, _id: int) -> str:
+        """
+        Args:   
+            _id (int): Label id.
+
+        Returns:
+            str: Label name (rough, smooth, or _background_).
+        """
+        return const.INT_RS_LABEL_MAP[self.get_label(_id)]
+
     def preprocess(self, img: np.ndarray) -> List[np.ndarray]:
         """Apply preprocessing step to the image
 
@@ -43,7 +53,7 @@ class SubDataset(Dataset):
             img (np.ndarray): Orginal image.
 
         Returns:
-            np.ndarray: Updated image with preprocessing, using a list to allow
+            List[np.ndarray]: Updated image with preprocessing, using a list to allow
             for multiple images to be output by preprocessing.
         """
         return [img]

### 2021-08-26 cuticle_analysis/datasets/utils.py

@@ -5,13 +5,13 @@ from typing import Dict, Tuple
 
 import numpy as np
 import pandas as pd
-import PIL.Image
+from PIL import Image, ImageDraw
 
 
 def shape_to_mask(img_shape, points, shape_type=None, line_width=10, point_size=5):
     mask = np.zeros(img_shape[:2], dtype=np.uint8)
-    mask = PIL.Image.fromarray(mask)
-    draw = PIL.ImageDraw.Draw(mask)
+    mask = Image.fromarray(mask)
+    draw = ImageDraw.Draw(mask)
     xy = [tuple(point) for point in points]
     if shape_type == "circle":
         assert len(xy) == 2, "Shape of shape_type=circle must have 2 points"

### 2021-08-26 scripts/data_spread.py

@@ -1,9 +1,15 @@
 
-import re
-
 import numpy as np
 
-from cuticle_analysis.datasets import RoughSmoothSub
+if __name__ == "__main__":
+    import cuticle_analysis
+    print(cuticle_analysis)
 
-data = RoughSmoothSub((16, 16))
-print(np.uniques(data.labels))
+    # for debugging
+    # from cuticle_analysis.core import init
+    # init()
+
+    from cuticle_analysis.datasets import RoughSmoothSub  # noqa
+
+    data = RoughSmoothSub((16, 16), save=True, rebuild=True)
+    print(np.unique(data.labels))

### 2021-08-26 scripts/data_spread.py
add todo

@@ -2,13 +2,8 @@
 import numpy as np
 
 if __name__ == "__main__":
-    import cuticle_analysis
-    print(cuticle_analysis)
-
-    # for debugging
-    # from cuticle_analysis.core import init
-    # init()
-
+    # TODO: add AllSub, RoughSmoothFull, and AllFull
+    # TODO: add class function for representing data spread
     from cuticle_analysis.datasets import RoughSmoothSub  # noqa
 
     data = RoughSmoothSub((16, 16), save=True, rebuild=True)

### 2021-08-27 cuticle_analysis/datasets/__init__.py
adjust datasets

@@ -1,97 +1,11 @@
 
-from typing import List
-
-import numpy as np
-
 from .. import const
 
 # register datasets
 from .dataset import Dataset
 from .full import FullDataset
 from .sub import SubDataset
+from .rs import RoughSmoothFull
+from .rs import RoughSmoothSub
 from .gabor import GaborRoughSmoothFull
 from .gabor import GaborRoughSmoothSub
-
-
-class AllFull(FullDataset):
-    'Full sized image dataset with all original labels.'
-
-    def __init__(self,
-                 size: tuple,
-                 excludes: list = None,
-                 random_seed: int = None,
-                 rebuild: bool = False,
-                 save: bool = False):
-        name = f'all_augmented'
-        d_type = const.DATASET_ALL  # used for converting labels
-        super().__init__(size,
-                         name=name,
-                         d_type=d_type,
-                         excludes=excludes,
-                         random_seed=random_seed,
-                         rebuild=rebuild,
-                         save=save)
-
-
-class AllFullAugmented(FullDataset):
-    'Full sized image dataset with all original labels and augmented data.'
-
-    def __init__(self,
-                 size: tuple,
-                 excludes: list = None,
-                 random_seed: int = None,
-                 rebuild: bool = False,
-                 save: bool = False):
-        name = f'all_augmented'
-        d_type = const.DATASET_ALL  # used for converting labels
-        super().__init__(size,
-                         name=name,
-                         d_type=d_type,
-                         excludes=excludes,
-                         random_seed=random_seed,
-                         rebuild=rebuild,
-                         save=save)
-        self.augment()
-
-    def augment(self):
-        self.images, self.labels = self.images, self.labels
-
-
-class RoughSmoothFull(FullDataset):
-    'Full sized image dataset with rough and smooth labels only.'
-
-    def __init__(self,
-                 size: tuple,
-                 excludes: list = None,
-                 random_seed: int = None,
-                 rebuild: bool = False,
-                 save: bool = False):
-        name = f'rs'
-        d_type = const.DATASET_RS  # used for converting labels to rough or smooth
-        super().__init__(size,
-                         name=name,
-                         d_type=d_type,
-                         excludes=excludes,
-                         random_seed=random_seed,
-                         rebuild=rebuild,
-                         save=save)
-
-
-class RoughSmoothSub(SubDataset):
-    'Subimage dataset with rough and smooth labels only.'
-
-    def __init__(self,
-                 size: tuple,
-                 excludes: list = None,
[…]

### 2021-08-27 cuticle_analysis/datasets/all.py

@@ -1,0 +1,47 @@
+
+from .. import const
+from .full import FullDataset
+
+
+class AllFull(FullDataset):
+    'Full sized image dataset with all original labels.'
+
+    def __init__(self,
+                 size: tuple,
+                 excludes: list = None,
+                 random_seed: int = None,
+                 rebuild: bool = False,
+                 save: bool = False):
+        name = f'all_augmented'
+        d_type = const.DATASET_ALL  # used for converting labels
+        super().__init__(size,
+                         name=name,
+                         d_type=d_type,
+                         excludes=excludes,
+                         random_seed=random_seed,
+                         rebuild=rebuild,
+                         save=save)
+
+
+class AllFullAugmented(FullDataset):
+    'Full sized image dataset with all original labels and augmented data.'
+
+    def __init__(self,
+                 size: tuple,
+                 excludes: list = None,
+                 random_seed: int = None,
+                 rebuild: bool = False,
+                 save: bool = False):
+        name = f'all_augmented'
+        d_type = const.DATASET_ALL  # used for converting labels
+        super().__init__(size,
+                         name=name,
+                         d_type=d_type,
+                         excludes=excludes,
+                         random_seed=random_seed,
+                         rebuild=rebuild,
+                         save=save)
+        self.augment()
+
+    def augment(self):
+        self.images, self.labels = self.images, self.labels

### 2021-08-27 cuticle_analysis/datasets/dataset.py

@@ -1,7 +1,7 @@
 
 import os
 import logging
-from typing import List, Tuple
+from typing import Dict, List, Tuple
 
 import cv2
 import pandas as pd
@@ -72,9 +72,22 @@ class Dataset():
             f'Total of {len(self.labels)} sub-images built from {len(np.unique(self.ids))} images.')
 
         logger.info(f'Class data:')
+        for k, v in self.class_data().items():
+            logger.info(f'{k}: {v}')
+
+    def class_data(self) -> Dict:
+        """Return the class data for the dataset.
+        Intended to be overridden by the dataset type.
+
+        Returns:
+            Dict: Class data of the dataset.
+        """
         uniques = np.unique(self.labels, return_counts=True)
+        res = {}
         for i in range(len(uniques[0])):
-            logger.info(f'\t{uniques[0][i]}: {uniques[1][i]}')
+            res[uniques[0][i]] = uniques[1][i]
+
+        return res
 
     def _build_labels(self, rebuild: bool, save: bool):
         """Helper for Dataset.build_labels, it will use the rebuild variable to

### 2021-08-27 cuticle_analysis/datasets/rs.py

@@ -1,0 +1,49 @@
+
+from typing import Dict
+
+from .. import const
+from .full import FullDataset
+from .sub import SubDataset
+
+
+class RoughSmoothFull(FullDataset):
+    'Full sized image dataset with rough and smooth labels only.'
+
+    def __init__(self,
+                 size: tuple,
+                 excludes: list = None,
+                 random_seed: int = None,
+                 rebuild: bool = False,
+                 save: bool = False):
+        name = f'rs'
+        d_type = const.DATASET_RS  # used for converting labels to rough or smooth
+        super().__init__(size,
+                         name=name,
+                         d_type=d_type,
+                         excludes=excludes,
+                         random_seed=random_seed,
+                         rebuild=rebuild,
+                         save=save)
+
+
+class RoughSmoothSub(SubDataset):
+    'Subimage dataset with rough and smooth labels only.'
+
+    def __init__(self,
+                 size: tuple,
+                 excludes: list = None,
+                 random_seed: int = None,
+                 rebuild: bool = False,
+                 save: bool = False):
+        name = f'rs'
+        d_type = const.DATASET_RS  # used for converting labels to rough or smooth
+        super().__init__(size,
+                         name=name,
+                         d_type=d_type,
+                         excludes=excludes,
+                         random_seed=random_seed,
+                         rebuild=rebuild,
+                         save=save)
+
+    def class_data(self) -> Dict:
+        return super().class_data()

### 2021-08-27 scripts/data_spread.py

@@ -4,6 +4,7 @@ import numpy as np
 if __name__ == "__main__":
     # TODO: add AllSub, RoughSmoothFull, and AllFull
     # TODO: add class function for representing data spread
+    # TODO: save tables to file for paper
     from cuticle_analysis.datasets import RoughSmoothSub  # noqa
 
     data = RoughSmoothSub((16, 16), save=True, rebuild=True)

### 2021-08-27 cuticle_analysis/datasets/dataset.py
add rsfull

@@ -1,5 +1,4 @@
 
-import os
 import logging
 from typing import Dict, List, Tuple
 
@@ -11,10 +10,6 @@ from numpy.random import default_rng
 from . import utils
 from .. import const
 
-if not os.path.exists('./logs'):
-    os.makedirs('./logs')
-
-logging.basicConfig(filename='./logs/dataset.log', level=logging.INFO)
 logger = logging.getLogger(__name__)
 
 

### 2021-08-27 cuticle_analysis/datasets/rs.py

@@ -25,6 +25,18 @@ class RoughSmoothFull(FullDataset):
                          rebuild=rebuild,
                          save=save)
 
+    def class_data(self) -> Dict:
+        """Return class data for the dataset.
+
+        Returns:
+            Dict: Class data for the dataset.
+        """
+        temp = super().class_data()
+        res = {}
+        for k, v in temp.items():
+            res[const.INT_RS_LABEL_MAP[k]] = v
+        return temp
+
 
 class RoughSmoothSub(SubDataset):
     'Subimage dataset with rough and smooth labels only.'
@@ -46,4 +58,13 @@ class RoughSmoothSub(SubDataset):
                          save=save)
 
     def class_data(self) -> Dict:
-        return super().class_data()
+        """Return class data for the dataset.
+
+        Returns:
+            Dict: Class data for the dataset.
+        """
+        temp = super().class_data()
+        res = {}
+        for k, v in temp.items():
+            res[const.INT_RS_LABEL_MAP[k+1]] = v
+        return res

### 2021-08-27 scripts/data_spread.py

@@ -1,11 +1,20 @@
 
-import numpy as np
+import logging
 
-if __name__ == "__main__":
-    # TODO: add AllSub, RoughSmoothFull, and AllFull
-    # TODO: add class function for representing data spread
-    # TODO: save tables to file for paper
-    from cuticle_analysis.datasets import RoughSmoothSub  # noqa
+from cuticle_analysis.core import init
+init()
 
-    data = RoughSmoothSub((16, 16), save=True, rebuild=True)
-    print(np.unique(data.labels))
+from cuticle_analysis.datasets import RoughSmoothFull  # noqa
+from cuticle_analysis.datasets import RoughSmoothSub  # noqa
+
+logger = logging.getLogger(__name__)
+
+rss = RoughSmoothSub((16, 16), save=True)
+rsf = RoughSmoothFull((512, 512), save=True)
+
+logger.info(f"type: subimage, size: {rss.size}, {rss.class_data()}")
+logger.info(f"type: full, size: {rsf.size}, {rsf.class_data()}")
+
+# TODO: add AllSub, and AllFull
+# TODO: add class function for representing data spread
+# TODO: save tables to file for paper

### 2021-08-27 cuticle_analysis/datasets/__init__.py
all full spread

@@ -5,6 +5,7 @@ from .. import const
 from .dataset import Dataset
 from .full import FullDataset
 from .sub import SubDataset
+from .all import AllFull
 from .rs import RoughSmoothFull
 from .rs import RoughSmoothSub
 from .gabor import GaborRoughSmoothFull

### 2021-08-27 cuticle_analysis/datasets/all.py

@@ -1,4 +1,6 @@
 
+from typing import Dict
+
 from .. import const
 from .full import FullDataset
 
@@ -12,7 +14,7 @@ class AllFull(FullDataset):
                  random_seed: int = None,
                  rebuild: bool = False,
                  save: bool = False):
-        name = f'all_augmented'
+        name = f'all'
         d_type = const.DATASET_ALL  # used for converting labels
         super().__init__(size,
                          name=name,
@@ -21,6 +23,18 @@ class AllFull(FullDataset):
                          random_seed=random_seed,
                          rebuild=rebuild,
                          save=save)
+
+    def class_data(self) -> Dict:
+        """Return class data for the dataset.
+
+        Returns:
+            Dict: Class data for the dataset.
+        """
+        temp = super().class_data()
+        res = {}
+        for k, v in temp.items():
+            res[const.INT_ALL_LABEL_MAP[k]] = v
+        return res
 
 
 class AllFullAugmented(FullDataset):
@@ -44,4 +58,5 @@ class AllFullAugmented(FullDataset):
         self.augment()
 
     def augment(self):
-        self.images, self.labels = self.images, self.labels
+        # self.images, self.labels = self.images, self.labels
+        raise NotImplementedError

### 2021-08-27 cuticle_analysis/datasets/rs.py

@@ -35,7 +35,7 @@ class RoughSmoothFull(FullDataset):
         res = {}
         for k, v in temp.items():
             res[const.INT_RS_LABEL_MAP[k]] = v
-        return temp
+        return res
 
 
 class RoughSmoothSub(SubDataset):

### 2021-08-27 cuticle_analysis/datasets/sub.py

@@ -37,9 +37,10 @@ class SubDataset(Dataset):
                          save=save)
 
     def get_label_name(self, _id: int) -> str:
-        """
+        """Given an image id, return the label name as a string.
+
         Args:   
-            _id (int): Label id.
+            _id (int): Image id.
 
         Returns:
             str: Label name (rough, smooth, or _background_).

### 2021-08-27 cuticle_analysis/datasets/utils.py

@@ -91,7 +91,7 @@ def convert_labels(label: pd.Series) -> pd.Series:
     label = label.replace(
         to_replace=r'^[^0-5].*', value=np.nan, regex=True)
     label = label.replace(
-        to_replace=r'1e', value=np.nan, regex=True)
+        to_replace=r'[1-3][a,e].*', value=np.nan, regex=True)
 
     # convert to dataframe and filter by existing label
     label = label.to_frame('class')

### 2021-08-27 scripts/data_spread.py

@@ -1,20 +1,18 @@
 
-import logging
-
 from cuticle_analysis.core import init
 init()
 
 from cuticle_analysis.datasets import RoughSmoothFull  # noqa
 from cuticle_analysis.datasets import RoughSmoothSub  # noqa
+from cuticle_analysis.datasets import AllFull  # noqa
 
-logger = logging.getLogger(__name__)
 
-rss = RoughSmoothSub((16, 16), save=True)
 rsf = RoughSmoothFull((512, 512), save=True)
+rss = RoughSmoothSub((16, 16), save=True)
+af = AllFull((512, 512), save=True)
 
-logger.info(f"type: subimage, size: {rss.size}, {rss.class_data()}")
-logger.info(f"type: full, size: {rsf.size}, {rsf.class_data()}")
+print(f"type: full, size: {rsf.size}, {rsf.class_data()}")
+print(f"type: subimage, size: {rss.size}, {rss.class_data()}")
+print(f"type: full, size: {af.size}, {af.class_data()}")
 
-# TODO: add AllSub, and AllFull
-# TODO: add class function for representing data spread
 # TODO: save tables to file for paper

### 2021-08-27 scripts/data_spread.py
save data spread tables to file

@@ -1,18 +1,111 @@
 
-from cuticle_analysis.core import init
-init()
+from typing import Dict
 
+import pandas as pd
+
+# for debugging, env logging level must be set to debug/info
+# from cuticle_analysis.core import init
+# init()
+
+from cuticle_analysis.datasets import AllFull  # noqa
 from cuticle_analysis.datasets import RoughSmoothFull  # noqa
 from cuticle_analysis.datasets import RoughSmoothSub  # noqa
-from cuticle_analysis.datasets import AllFull  # noqa
 
 
-rsf = RoughSmoothFull((512, 512), save=True)
-rss = RoughSmoothSub((16, 16), save=True)
+all_cols = [
+    'type',
+    'rough dimpled',
+    'rough netted',
+    'rough ridged',
+    'rough T',
+    'smooth gritty',
+    'smooth smooth',
+    'total'
+]
+rs_cols = [
+    'type',
+    'rough',
+    'smooth',
+    'total'
+]
+
+
+def total_samples(class_data: Dict) -> int:
+    """
+    Total number of samples in the dataset.
+    """
+    res = 0
+    for _, data in class_data.items():
+        res += data
+    return res
+
+
+def all_row(type: str, class_data: Dict) -> pd.DataFrame:
+    row = pd.DataFrame([[
+        type,
+        class_data['rough dimpled'],
+        class_data['rough netted'],
+        class_data['rough ridged'],
+        class_data['rough T'],
+        class_data['smooth gritty'],
+        class_data['smooth smooth'],
+        total_samples(class_data)
+    ]], columns=all_cols)
+    return row
+
+
+def rs_row(type: str, size: tuple, class_data: Dict) -> pd.DataFrame:
+    if type != 'full':
+        type = f'{type}_{size}'
+
+    row = pd.DataFrame([[
+        type,
+        class_data['rough'],
+        class_data['smooth'],
+        total_samples(class_data)
+    ]], columns=rs_cols)
+    return row
+
+
+# init output tables
+all_df = pd.DataFrame(columns=all_cols)
+rs_df = pd.DataFrame(columns=rs_cols)
+
+# add all full dataset
 af = AllFull((512, 512), save=True)
+all_df = all_df.append(all_row('full', af.class_data()), ignore_index=True)
 
-print(f"type: full, size: {rsf.size}, {rsf.class_data()}")
[…]

### 2021-08-27 cuticle_analysis/core/__init__.py
script to generate tables

@@ -33,6 +33,8 @@ def init():
 
 
 def start_app():
+    init()
+
     args = get_args()
 
     if args.download_dataset:

### 2021-08-27 cuticle_analysis/core/dataset.py

@@ -11,21 +11,25 @@ logger = logging.getLogger(__name__)
 def download_dataset():
     import gdown  # type: ignore
 
-    logger.info("Downloading dataset...")
-    url = "https://drive.google.com/uc?id=1xABlWE790uWmT0mMxbDjn9KZlNUB6Y57"
     output = "./dataset.zip"
-    gdown.download(url, output, quiet=False)
 
-    assert os.path.isfile(output)
-    logger.info("Downloaded dataset!")
+    if not os.path.isfile(output):
+        logger.info("Downloading dataset...")
+        url = "https://drive.google.com/uc?id=1xABlWE790uWmT0mMxbDjn9KZlNUB6Y57"
+        gdown.download(url, output, quiet=False)
+        assert os.path.isfile(output)
+        logger.info("Downloaded dataset!")
+
+    else:
+        logger.info("Found dataset!")
 
 
 def unzip_dataset(path="./dataset.zip"):
-    logger.info("Extracting dataset...")
     out_path = os.path.join(os.curdir, "dataset")
 
     try:
         if not os.path.isdir(out_path):
+            logger.info("Extracting dataset...")
             if os.path.isfile(path):
                 from zipfile import PyZipFile
                 zf = PyZipFile(path)
@@ -36,7 +40,7 @@ def unzip_dataset(path="./dataset.zip"):
             else:
                 raise Exception(f"Failed to find {path}.")
         else:
-            raise Exception(f"Directory {out_path} already exists.")
+            logger.info("Skipping extract...")
 
     except Exception as e:
         logger.error(f'Failed to unzip dataset:" {e}')

### 2021-08-27 scripts/gen_assests.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+nix-shell --run "
+cuticle_analysis --download-dataset \
+&& python scripts/data_spread.py
+"

### 2021-08-27 paper/paper.tex

@@
 \author{%
-  First name Last name\affil{1}, First name Last name\affil{2} and First name
-  Last name\affil{3,}\corrauth}
+  Noah Gardner\affil{1} and Chih-Cheng Hung\affil{1}\corrauth}
 
@@
     State University, 1000 Chastain Road, Kennesaw, GA 30144, USA; email}
-  \addr{\affilnum{2}}{Affiliation email} \addr{\affilnum{3}}{Affiliation email}}
+  % \addr{\affilnum{2}}{Affiliation email}
+}
 % corresponding author
-\corraddr{Email address of corresponding author; Tel: +1-111-111-1111; Fax:\\
-  +1-111-111-1111.}
+\corraddr{chung1@kennesaw.edu}
 
@@
 \begin{abstract}
-  An abstract is normally of 150 to 250 words. List 4 to 8 keywords all with
-  lowercase (except special cases) for both highlighting the focuses of your
-  study and indexing/searching in databases. Article type can be one of the
-  following: regular article, case study, express letter, review or overview,
-  survey, opinion, perspective, communication, lecture notes, tutorial notes,
-  and editorial (special issue only).
+  There is a large variety of ant species, and most species are diverse in terms
+  of size, shape, behaviors, and especially skin (cuticle) textures. However,
+  the significance of ant cuticle texture is not widely researched. This
+  research employs modern machine learning methods such as texture analysis and
+  classification with CNN and clustering to automatically group similar ant
+  species to allow for the study of influences cuticle texture on ant
+  ecology.
 
@@
 
-\keywords{(4 to 8 keywords)}
+\keywords{Texture Analysis, Image Processing, Clustering, Machine Learning,
+  Myrmecology, Ecology}
 
@@
 pre-screening, the Editor may act as the “Academic Editor” or assign this
-manuscript to another Editor from the Editorial Board as the “Academic Editor”
-to handle the review of this paper, communicate with the authors on revisions,
-and inform the final decision after consultation with the relevant
-Editor-in-Chief.
+manuscript to another Editor from the Editorial Board as the “Academic Editor”to
+handle the review of this paper, communicate with the authors on revisions, and
+inform the final decision after consultation with the relevant Editor-in-Chief.
 
@@
 \section{Reference style, citation, and cross-reference}
-Test citation \cite{Jensen1996}.
+Test citation \cite{ward_phylogeny_2014}.
 

### 2021-08-31 paper/paper.tex
add intro, start dataset section

@@
 \usepackage{cite}
-\def\typeofarticle{Article type} \def\currentvolume{1} \def\currentissue{1}
+
+\usepackage{hyperref}
+\hypersetup{colorlinks=true}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newcommand{\ep}{\varepsilon}
+\newcommand{\eps}[1]{{#1}_{\varepsilon}}
+
+\def\typeofarticle{Research Article} \def\currentvolume{1} \def\currentissue{1}
 \def\currentyear{2021} \def\currentmonth{} \def\ppages{1--30} \def\DOI{to
@@
 2022 }
-
-\newcommand{\ep}{\varepsilon}
-\newcommand{\eps}[1]{{#1}_{\varepsilon}}
-
 %\numberwithin{equation}{section}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
 
-\title{Title of the paper}
+\title{An Ant Cuticle Texture Classification Algorithm for Ecological Anaylsis}
 
 \author{%
-  Noah Gardner\affil{1} and Chih-Cheng Hung\affil{1}\corrauth}
+  Noah Gardner\affil{1}, John Paul Hellenbrand\affil{2}, and Chih-Cheng
+  Hung\affil{1}\corrauth}
 
@@
     State University, 1000 Chastain Road, Kennesaw, GA 30144, USA; email}
-  % \addr{\affilnum{2}}{Affiliation email}
-}
+  \addr{\affilnum{2}}{College of Science and Mathematics, Kennesaw State
+    University, 1000 Chastain Road, Kennesaw, GA 30144, USA; email}}
 % corresponding author
@@
 
-\editor{ First-name Last-name}
+\editor{First-name Last-name}
 
@@
   classification with CNN and clustering to automatically group similar ant
-  species to allow for the study of influences cuticle texture on ant
-  ecology.
+  species to allow for the study of influences cuticle texture on ant ecology.
 
@@
 \section{Introduction}
+Insects compose half of biodiversity and rank among the most dominant organisms
+in terrestrial ecosystems \cite{sheikh_diverse_2017}. A key factor for the
+ecological success of insects is their exoskeleton, also known as cuticle. The
+cuticle protects insects from predation, provides structural support, prevents
+desiccation, and serves as a canvas for advertising visual and chemical signals
+\cite{gullan_insects_2009}. Research has heavily focused on the macrostructures
+and internal chemical components that make the exoskeleton functional and more
+recent work is being done to understand the functional aspects of external
+cuticle micro sculpturing \cite{muthukrishnan_insect_2020,
+  gunderson_insect_1989, watson_diversity_2017}.
 
-Submission of a manuscript implies that the work described has not been
-published before, and not been under consideration for publishing anywhere else.
-Your submission also means that the manuscript has been approved by all
-co-authors with appropriate permission if required. The publisher will not be
-held legally responsible should any dispute be raised for any reason. AIMS-ACI
-is an OA publication published quarterly by AIMS Press in February, May, August,
-and November. Publishing is currently free of charge. Readers can access all
-published articles free of charge through the journal website
-https://aimspress.com/journal/aci
+We examine ants (\textit{Formicidae}) as they display an extreme diversity of
+cuticle micro sculpturing across all subfamilies. Sculpturing ranges from
+parallel longitudinal ridges to deep oval impressions to erratic protuberances.
+The sculpturing has arisen convergently and independently throughout ant’s
+evolutionary history, which suggests some inherent function. Cuticle sculpturing
+on ants may help increase strength and rigidness, resist abrasion, increase
+internal and external surface area, resist microbial growth, and rear beneficial
+anti-biotic producing bacteria \cite{johnson_effect_2011,
+  bruckner_relationship_2017, currie_coevolved_2006}. These specific functions may
+be associated with certain sculpturing types and the purpose of classification
+is to group similar textures based on proposed function.
 
-\section{Layout and fonts}
-The page size is A4 ($8.5 \times 11.0$ in or $21.59 \times 27.94$ cm). The
-margin is set to moderate. Font for all text, except symbols and formulae,
-should use Times New Roman. Article title is bold with a font size 16 and
-aligned to the left. Authors are bold with a font size 12. Using Times New Roman
-size 12 for the main text, including figure captions and table titles. However,
-text size for contents in a table can be chosen from 8 to 12 to fit the table
-within the width of the page.
+\section{Methods}
 
-\section{Headline and sub-headline}
-The first-level headline is numbered and bold with a font size 12, and has a
-space of 12 $pt$ before and after the headline. Each paragraph begins with an
-indentation of 0.63 $cm$ (the same by using the Tab key once) in the first line.
-A paragraph does not have a space before and after that paragraph.
+\subsection{Sculpture Identification Protocol}
 
-\subsection{The second-level headline}
-The second-level headline is numbered with a font size 12. The second-level
-headline has a default space of 6 $pt$ before and after it.
+\subsection{Dataset}
+In order to classify the cuticle, we used ant head images sourced from AntWeb
+\cite{antweb}. In general, the ant head images are centered in the image, facing
+the front, and share a similar posture. However, some images may not be
+centered, show the ant head in a different orientation, or may have a
+drastically different resolution from the average image. Additionally, images
+can vary in scale. Fortunately, each can have several scales available such that
+the images' scales can be roughly similar.
 
-\subsubsection{The third-level headline (if necessary)}
-If really necessary, the third-level headline should be numbered and italicized
-with a font size 12, indented by 0.63 $cm$ (the same by using the Tab key once).
-The third-level headline has a default space of 6 $pt$ before and after it. Any
-lower headline should be avoided.
+Images were collected and classified by undergraduate students according to the
+sculpturing identification protocol. Overall, there are a total of 2443 images
+in the dataset. The class spread is shown in
 
-\subsection{Footnotes}
-Although not encouraged, footnotes can be used to provide additional information
-for an informal reference that cannot be included in the reference list. A
-footnote should be numbered and cannot contain any figure or table. Use font
-size 8 for footnotes.
-
-\section{Figures or images, captions, tables, table titles, and formulae}
-\subsection{Figure/image and caption}
-Leave a space of one line before and after a figure or an image, i.e., one line
-between the main text and the top of the figure or image, and one line between
-the bottom of the figure or image and the caption. The caption has a default
-space of 12 $pt$ after it so the main text can continue below the caption. If no
-text follows the figure caption, do not leave any space between the caption and
-the next headline (the headline has a default space).
-
-\begin{figure}[ht]
-  \centering
-  \vskip 0.2in
-  %\includegraphics[width=0.85\textwidth]{{logo_new.pdf}}
-  \vspace {12pt}
-  \caption{ Insert the figure/image here}
-  \label{fig:res}
-\end{figure}
-
-Figures should be numbered consecutively in the text. A figure can be referred
-explicitly in the text as Figure~\ref{fig:res}, or implicitly at the end of a
-sentence in parentheses (Fig. 1).
-
-Use black and white graphic for line drawings. All lines should be at least 0.1
-mm (0.3 $pt$) wide. Line drawings or scanned line drawings should have a
-resolution at least 600 dpi. Images with a large volume should be properly
-compressed within 1 MB each. Scanned images should be properly edited to balance
-the volume size and clarity of the image.
-
-Figure captions begin with \textbf{Figure x.} in bold, followed by the text. Do
-not add a full stop at the end of the caption.
-
-\subsection{Tables and table titles}
-\begin{table}[H]
-  \begin{center}
-    \caption{Publication schedule for AIMS-ACI in 2022}
-    \begin{tabular}{lll} \hline
-      Issue        & Publishing date   & Submission deadline \\\hline
-      First issue  & February 25, 2022 & January 15, 2022    \\
-      Second issue & May 25, 2022      & April 15, 2022      \\
-      Third issue  & August 25, 2022   & July 15, 2022       \\
-      Fourth issue & November 25, 2022 & October 15, 2022    \\\hline
-    \end{tabular}
-    \label{table}
-  \end{center}
-\end{table}
-
-Table title should be placed above the table, numbered consecutively, and
-referred in the text like Table~\ref{table} or (Table~\ref{table}). There is no
-space between the table and table title. Leave a space of one line above the
-table title and one line below the bottom of the table. If a new first-level
-headline follows the table immediately, do not leave the space below the table.
-
-The font of the table title is size 12 but the font of the table contents can be
-sizes 8 to 12 depending on the best fit for the table. Table title begins with
-\textbf{Table x}. in bold, followed by the text. Do not add a full stop at the
-end of the table title.
-
-\subsection{Mathematical formulae}
-Mathematical formulae should be numbered consecutively by placing the number in
-parentheses like (1), one TAB space away from the right line. Mathematical
-formulae should be typed using MS Word Equation or better using \emph{MathType}.
-Use size 10--12 for the main characters and size 6--8 for subscripts or
-superscripts. Leave a space of 6 $pt$ between two formula in separate lines.
-Some examples are shown below.
-
-\begin{equation}
-  \lim_{x \mapsto a^+ }f\left ( x \right )= \pm \infty \qquad \text{or} \qquad \lim_{x \mapsto a^- }f\left ( x \right )= \pm \infty
-  \label{equation1}
-\end{equation}
-
-\begin{equation}
-  \lim_{x \mapsto \infty }f\left ( x \right )=  a \qquad \text{or} \qquad \lim_{x \mapsto -\infty}f\left ( x \right )=  b
-  \label{equation2}
-\end{equation}
-
-To cite a formula in the text, use the number like Equation~(\ref{equation1}),
-or Eq.~(\ref{equation2}), or alike.
-
-\section{Submission, review, and decision}
-In the first year, submission, review and decisions of manuscripts is conducted
-through email communications. Please send your manuscript to one of the relevant
-Editors-in-Chief by their email available on the journal Website. With the
-increase in submissions from 2022, we would use the AIMS Editorial System to
-handle the process of submission, review, and decisions.
-
-Once a submission is received, the receiving Editor will pre-screen the article
-for its suitability and quality for the journal. After passing this
-pre-screening, the Editor may act as the “Academic Editor” or assign this
-manuscript to another Editor from the Editorial Board as the “Academic Editor”to
-handle the review of this paper, communicate with the authors on revisions, and
-inform the final decision after consultation with the relevant Editor-in-Chief.
-
-AIMS-ACI holds a high standard in peer review. A submitted manuscript should be
-reviewed by at least three independent reviewers who are scholars and
-knowledgeable of the area of research. Standard review forms are available for
-reviewers but they could choose own appropriate way of reporting to the Academic
-Editor.
-
-\section{Reference style, citation, and cross-reference}
-Test citation \cite{ward_phylogeny_2014}.
+\section{Experimental Results}
 
@@
 
+\bibliography{paper}
 \bibliographystyle{unsrt}
-\bibliography{paper}
 

### 2021-08-31 scripts/data_spread.py
update table gen

@@ -9,29 +9,24 @@ import pandas as pd
 from cuticle_analysis.core import init
 init()
 
+from cuticle_analysis import const  # noqa
 from cuticle_analysis.datasets import AllFull  # noqa
 from cuticle_analysis.datasets import RoughSmoothFull  # noqa
 from cuticle_analysis.datasets import RoughSmoothSub  # noqa
 
+
 logger = logging.getLogger(__name__)
 
+all_labels = list(const.ALL_LABEL_MAP.keys())
+all_labels.remove('_background_')
+rs_labels = list(const.RS_LABEL_MAP.keys())
+rs_labels.remove('_background_')
 
-all_cols = [
-    'type',
-    'rough dimpled',
-    'rough netted',
-    'rough ridged',
-    'rough T',
-    'smooth gritty',
-    'smooth smooth',
-    'total'
-]
-rs_cols = [
-    'type',
-    'rough',
-    'smooth',
-    'total'
-]
+_TOTAL = 'Total'
+_TYPE = 'Type'
+
+all_cols = [_TYPE] + all_labels + [_TOTAL]
+rs_cols = [_TYPE] + rs_labels + [_TOTAL]
 
 
 def total_samples(class_data: Dict) -> int:
@@ -45,16 +40,10 @@ def total_samples(class_data: Dict) -> int:
 
 
 def all_row(type: str, class_data: Dict) -> pd.DataFrame:
-    row = pd.DataFrame([[
-        type,
-        class_data['rough dimpled'],
-        class_data['rough netted'],
-        class_data['rough ridged'],
-        class_data['rough T'],
-        class_data['smooth gritty'],
-        class_data['smooth smooth'],
-        total_samples(class_data)
-    ]], columns=all_cols)
+    data = [class_data[label] for label in all_labels]
+    cols = [type] + data + [total_samples(class_data)]
+    row = pd.DataFrame(
+        [cols], columns=all_cols)
     return row
 
 
@@ -62,12 +51,9 @@ def rs_row(type: str, size: tuple, class_data: Dict) -> pd.DataFrame:
     if type != 'full':
         type = f'{type}_{size}'
 
-    row = pd.DataFrame([[
-        type,
-        class_data['rough'],
-        class_data['smooth'],
-        total_samples(class_data)
-    ]], columns=rs_cols)
[…]

### 2021-08-31 paper/paper.tex
update table gen
add percent col to all_samples

@@
 \usepackage{cite}
+\usepackage{booktabs}
 
@@
 centered, show the ant head in a different orientation, or may have a
-drastically different resolution from the average image. Additionally, images
-can vary in scale. Fortunately, each can have several scales available such that
-the images' scales can be roughly similar.
+drastically different resolution from the average image. Fortunately, each can
+have several scales available such that the images' scales can be roughly
+similar.
+
+\begin{table}[h]
+  \centering
+  \caption{Dataset Class Distribution}
+  \label{table:classdist}
+  \include{tables/all_samples.tex}
+\end{table}
 
@@
 sculpturing identification protocol. Overall, there are a total of 2443 images
-in the dataset. The class spread is shown in
+in the dataset. The class distribution is shown in Table \ref{table:classdist}.
+The relatively small number of images in the dataset paired with the uneven
+class distribution makes it difficult to perform a meaningful classification.
+However, the goal of this project is to develop a classification algorithm that
+can be used to help classify over 200,000 ant images. Therefore, it is desirable
+to use a few representative images from each class to train the algorithm.
+Additionally, data augmentation and other methods are performed to increase the
+number of training images to create a more robust classification algorithm.
 

### 2021-08-31 scripts/data_spread.py

@@ -84,8 +84,13 @@ for size in sizes:
     gc.collect()
 
 # output
-logger.info("number of samples per class per dataset")
-all_df = all_df.transpose().rename(columns={_TYPE: 'Label', 0: 'Samples'})
+# add row for samples as percentage of total
+samples_pct = all_df.loc[:, all_cols[1:]].div(
+    all_df[_TOTAL], axis=0)
+samples_pct = samples_pct.astype(float).round(2)
+all_df = all_df.append(samples_pct, ignore_index=True)
+all_df = all_df.transpose().rename(
+    columns={_TYPE: 'Label', 0: 'Samples (n)', 1: 'Samples (%)'})
 rs_df = rs_df.transpose()
 
 # drop first row
@@ -97,16 +102,5 @@ logger.info(rs_df)
 all_df.to_latex('./paper/tables/all_samples_int.tex')
 rs_df.to_latex('./paper/tables/rs_samples_int.tex')
 
-# logger.info("\n samples per class as percentage")
-# all_df.loc[:, all_cols[1:]] = all_df.loc[:, all_cols[1:]].div(
-#     all_df[_TOTAL], axis=0) * 100
 # rs_df.loc[:, rs_cols[1:]] = rs_df.loc[:, rs_cols[1:]].div(
 #     rs_df[_TOTAL], axis=0) * 100
-# all_df = all_df.drop(columns=[_TOTAL])
-# rs_df = rs_df.drop(columns=[_TOTAL])
-# logger.info(all_df)
-# logger.info(rs_df)
-
-# # save to file
-# all_df.to_latex('./paper/tables/all_samples_pct.tex')
-# rs_df.to_latex('./paper/tables/rs_samples_pct.tex')

### 2021-08-31 scripts/data_spread.py
add overall class distribution

@@ -83,7 +83,7 @@ for size in sizes:
     del temp
     gc.collect()
 
-# output
+# output all full df
 # add row for samples as percentage of total
 samples_pct = all_df.loc[:, all_cols[1:]].div(
     all_df[_TOTAL], axis=0)
@@ -91,16 +91,33 @@ samples_pct = samples_pct.astype(float).round(2)
 all_df = all_df.append(samples_pct, ignore_index=True)
 all_df = all_df.transpose().rename(
     columns={_TYPE: 'Label', 0: 'Samples (n)', 1: 'Samples (%)'})
-rs_df = rs_df.transpose()
 
 # drop first row
 all_df = all_df.iloc[1:]
 
 # save to file
 logger.info(all_df)
-logger.info(rs_df)
-all_df.to_latex('./paper/tables/all_samples_int.tex')
-rs_df.to_latex('./paper/tables/rs_samples_int.tex')
+all_df.to_latex('./paper/tables/all_samples_full.tex')
+
+# output rs full df
+rs_full_df = rs_df.loc[0, :].to_frame()
+
+# drop first row
+rs_full_df = rs_full_df.iloc[1:]
+rs_full_df = rs_full_df.transpose()
+logger.info(rs_full_df)
+
+
+# add row for samples as percentage of total
+samples_pct = rs_full_df.loc[:, rs_cols[1:]].div(
+    rs_full_df[_TOTAL], axis=0)
+samples_pct = samples_pct.astype(float).round(2)
+rs_full_df = rs_full_df.append(samples_pct, ignore_index=True)
+rs_full_df = rs_full_df.transpose().rename(
+    columns={_TYPE: 'Label', 0: 'Samples (n)', 1: 'Samples (%)'})
+
+logger.info(rs_full_df)
+rs_full_df.to_latex('./paper/tables/rs_samples_full.tex')
 
 # rs_df.loc[:, rs_cols[1:]] = rs_df.loc[:, rs_cols[1:]].div(
 #     rs_df[_TOTAL], axis=0) * 100

### 2021-09-01 scripts/data_spread.py
update script

@@ -1,5 +1,5 @@
 
-from typing import Dict
+from typing import Dict, List
 import gc
 import logging
 
@@ -83,41 +83,50 @@ for size in sizes:
     del temp
     gc.collect()
 
-# output all full df
-# add row for samples as percentage of total
-samples_pct = all_df.loc[:, all_cols[1:]].div(
-    all_df[_TOTAL], axis=0)
-samples_pct = samples_pct.astype(float).round(2)
-all_df = all_df.append(samples_pct, ignore_index=True)
-all_df = all_df.transpose().rename(
-    columns={_TYPE: 'Label', 0: 'Samples (n)', 1: 'Samples (%)'})
 
-# drop first row
-all_df = all_df.iloc[1:]
-
-# save to file
-logger.info(all_df)
-all_df.to_latex('./paper/tables/all_samples_full.tex')
-
-# output rs full df
-rs_full_df = rs_df.loc[0, :].to_frame()
-
-# drop first row
-rs_full_df = rs_full_df.iloc[1:]
-rs_full_df = rs_full_df.transpose()
-logger.info(rs_full_df)
+def distribution_pct(df: pd.DataFrame, cols: List) -> pd.DataFrame:
+    """
+    Calculate the distribution of a class in a dataframe as a percent.
+    Returns:
+        df: Transposed dataframe with the columns ['Label', 'Samples (n)', 'Samples (%)']
+    """
+    sample_pct = df.loc[:, cols[1:]].div(
+        df[_TOTAL], axis=0).astype(float).round(2)
+    df = df.append(sample_pct, ignore_index=True)
+    return df.transpose().rename(
+        columns={_TYPE: 'Label', 0: 'Samples (n)', 1: 'Samples (%)'})
 
 
-# add row for samples as percentage of total
-samples_pct = rs_full_df.loc[:, rs_cols[1:]].div(
-    rs_full_df[_TOTAL], axis=0)
-samples_pct = samples_pct.astype(float).round(2)
-rs_full_df = rs_full_df.append(samples_pct, ignore_index=True)
-rs_full_df = rs_full_df.transpose().rename(
-    columns={_TYPE: 'Label', 0: 'Samples (n)', 1: 'Samples (%)'})
+def save_all_full_df(df: pd.DataFrame, cols: List) -> None:
+    """
+    Save the dataframe to a latex file. Add percent samples column.
+    """
[…]

### 2021-09-02 scripts/data_spread.py
add index column name

@@ -93,8 +93,10 @@ def distribution_pct(df: pd.DataFrame, cols: List) -> pd.DataFrame:
     sample_pct = df.loc[:, cols[1:]].div(
         df[_TOTAL], axis=0).astype(float).round(2)
     df = df.append(sample_pct, ignore_index=True)
-    return df.transpose().rename(
+    df = df.transpose().rename(
         columns={_TYPE: 'Label', 0: 'Samples (n)', 1: 'Samples (%)'})
+    df.rename_axis('Label', axis=1, inplace=True)
+    return df
 
 
 def save_all_full_df(df: pd.DataFrame, cols: List) -> None:

### 2021-09-02 cuticle_analysis/antweb/__init__.py
scraper start

@@ -1,0 +1,4 @@
+
+
+# exports
+from .specimen import SpecimenScraper

### 2021-09-02 cuticle_analysis/antweb/specimen.py

@@ -1,0 +1,41 @@
+
+import requests
+from bs4 import BeautifulSoup
+
+from ..datasets.helper import DatasetHelper
+
+
+class SpecimenScraper():
+    def __init__(self):
+        self.dh = DatasetHelper()
+        self.base_url = "https://antweb.org/images.do"
+
+    def get_specimen(self, info: dict) -> str:
+        """
+            Get specimen using taxonomic information.
+        """
+        payload = {
+            "genus": info["genus"],
+            "species": info["species"],
+            "rank": "species",
+            "project": "allantwebants"
+        }
+        r = requests.get(self.base_url, params=payload)
+
+        soup = BeautifulSoup(r.text, "html.parser")
+        div = soup.find("div", {"class": "name"})
+
+        url = ""
+        name = ""
+        if div:
+            if div.has_attr("href"):
+                url = div["href"]
+            name = div.text
+        return name, url
+
+    def get_specimen_by_id(self, _id: int) -> str:
+        """
+            Get specimen by id.
+        """
+        ant_info = self.dh.get_ant_info(_id)
+        return self.get_specimen(ant_info)

### 2021-09-02 cuticle_analysis/datasets/__init__.py

@@ -1,7 +1,5 @@
 
-from .. import const
-
-# register datasets
+# exports
 from .dataset import Dataset
 from .full import FullDataset
 from .sub import SubDataset

### 2021-09-02 cuticle_analysis/datasets/dataset.py

@@ -9,11 +9,12 @@ from numpy.random import default_rng
 
 from . import utils
 from .. import const
+from .helper import DatasetHelper
 
 logger = logging.getLogger(__name__)
 
 
-class Dataset():
+class Dataset(DatasetHelper):
     def __init__(self,
                  size: tuple,
                  name: str = "",
@@ -207,25 +208,6 @@ class Dataset():
                 raise ValueError(
                     f'Class of ID[{_id}] is NA or not considered in this version.')
 
-    def get_image(self, _id: int) -> np.ndarray:
-        """Get image by ID.
-
-        Args:
-            _id (int): ID of the sample.
-
-        Returns:
-            img (np.ndarray): Image as cv2 image object (numpy array).
-        """
-        path = f'./dataset/data/{_id}.jpg'
-        img = cv2.imread(path)
-
-        if img is None:
-            msg = f'Failed to open image {path}'
-            logger.error(msg)
-            raise ValueError(msg)
-
-        return img
-
     def is_included(self, _id: int) -> bool:
         """Given an image id, return if the image is included from the dataset.
 
@@ -238,31 +220,6 @@ class Dataset():
         if _id in self.ids:
             return True
         return False
-
-    def get_ant_info(self, _id: int) -> List[str]:
-        """Get ant species info from original dataset.
-
-        Args:
-            _id (int): ID of the sample.
-
-        Returns:
-            List[str]: List of ant species info.
-        """
-        row = self.ant_data.loc[self.ant_data['Photo_number'] == _id]
-
-        res = []
-        if not row['Sub-species'].isnull().any():
-            res.append(f'Sub-species: {row["Sub-species"].values[0]}')
-        if not row['Species'].isnull().any():
-            res.append(f'Species: {row["Species"].values[0]}')
-        if not row['Subgenus'].isnull().any():
-            res.append(f'Subgenus: {row["Subgenus"].values[0]}')
-        if not row['Genus'].isnull().any():
-            res.append(f'Genus: {row["Genus"].values[0]}')
-        if not row['Sub-Family'].isnull().any():
-            res.append(f'Sub-Family: {row["Sub-Family"].values[0]}')
-
-        return res
 
     def stratified_split(self, n: int) -> Tuple[np.ndarray, np.ndarray]:
         """Stratified split with $n samples per class.

### 2021-09-02 cuticle_analysis/datasets/helper.py

@@ -1,0 +1,64 @@
+
+import logging
+from typing import Dict
+
+import cv2
+import numpy as np
+import pandas as pd
+
+
+logger = logging.getLogger(__name__)
+
+
+class DatasetHelper():
+    """
+        A helper class for the Dataset class for functions relating to basic
+        dataset access.
+    """
+
+    def __init__(self):
+        self.ant_data = pd.read_excel(f'./dataset/labels.xlsx', header=0)
+
+    def get_image(self, _id: int) -> np.ndarray:
+        """Get image by ID.
+
+        Args:
+            _id (int): ID of the sample.
+
+        Returns:
+            img (np.ndarray): Image as cv2 image object (numpy array).
+        """
+        path = f'./dataset/data/{_id}.jpg'
+        img = cv2.imread(path)
+
+        if img is None:
+            msg = f'Failed to open image {path}'
+            logger.error(msg)
+            raise ValueError(msg)
+
+        return img
+
+    def get_ant_info(self, _id: int) -> Dict:
+        """Get ant species info from original dataset.
+
+        Args:
+            _id (int): ID of the sample.
+
+        Returns:
+            Dict[str]: List of ant species info.
+        """
+        row = self.ant_data.loc[self.ant_data['Photo_number'] == _id]
+
+        res = {}
+        if not row['Sub-species'].isnull().any():
+            res['subspecies'] = row["Sub-species"].values[0]
+        if not row['Species'].isnull().any():
+            res['species'] = row["Species"].values[0]
+        if not row['Subgenus'].isnull().any():
+            res['subgenus'] = row["Subgenus"].values[0]
+        if not row['Genus'].isnull().any():
+            res['genus'] = row["Genus"].values[0]
+        if not row['Sub-Family'].isnull().any():
+            res['subfamily'] = row["Sub-Family"].values[0]
+
+        return res

### 2021-09-02 scripts/images_cr.py

@@ -1,0 +1,22 @@
+
+import logging
+
+from bs4 import BeautifulSoup
+
+# for debugging, env logging level must be set to debug/info
+from cuticle_analysis.core import init
+init()
+
+from cuticle_analysis.antweb import SpecimenScraper  # noqa
+
+scraper = SpecimenScraper()
+logger = logging.getLogger(__name__)
+
+ids = [1]
+for _id in ids:
+    logger.info(f"Processing  ID: {_id}")
+    res = scraper.get_specimen_by_id(_id)
+    if res:
+        logger.info(res)
+    else:
+        logger.info(f"No data found for ID: {_id}")

### 2021-09-02 cuticle_analysis/antweb/specimen.py
scrape author

@@ -1,14 +1,22 @@
 
+import re
+import logging
+from typing import Dict
+
 import requests
 from bs4 import BeautifulSoup
 
 from ..datasets.helper import DatasetHelper
 
+logger = logging.getLogger(__name__)
+
 
 class SpecimenScraper():
     def __init__(self):
         self.dh = DatasetHelper()
-        self.base_url = "https://antweb.org/images.do"
+        self.base_url = "https://antweb.org"
+        self.images_url = f"{self.base_url}/images.do"
+        self.pic_url = f"{self.base_url}/bigPicture.do"
 
     def get_specimen(self, info: dict) -> str:
         """
@@ -20,18 +28,14 @@ class SpecimenScraper():
             "rank": "species",
             "project": "allantwebants"
         }
-        r = requests.get(self.base_url, params=payload)
-
+        r = requests.get(self.images_url, params=payload)
         soup = BeautifulSoup(r.text, "html.parser")
+
+        # parse name div
         div = soup.find("div", {"class": "name"})
 
-        url = ""
-        name = ""
-        if div:
-            if div.has_attr("href"):
-                url = div["href"]
-            name = div.text
-        return name, url
+        name = str(div.text).lower()
+        return name
 
     def get_specimen_by_id(self, _id: int) -> str:
         """
@@ -39,3 +43,53 @@ class SpecimenScraper():
         """
         ant_info = self.dh.get_ant_info(_id)
         return self.get_specimen(ant_info)
+
+    def get_meta_by_specimen(self, _id: int) -> Dict:
+        """
+            Get high res image, author and author_url by id.
+        """
+        payload = {
+            "name": self.get_specimen_by_id(_id),
+            "shot": "h",
+            "number": "1"
+        }
+        r = requests.get(self.pic_url, params=payload)
+        soup = BeautifulSoup(r.text, "html.parser")
+
+        # parse photo_metadata span
+        span = soup.find("span", {"id": "photo_metadata"})
+        elements = span.find_all("li")
+
+        res = {
+            "author": "",
+            "author_url": "",
+            "image_url": ""
+        }
+        image_pattern = r"View Highest Resolution"
+        author_pattern = r"Photographer: "
+        for element in elements:
+            if re.match(image_pattern, element.text):
+                res["image_url"] = element.a.get("href")
+            elif re.match(author_pattern, element.text):
+                res["author"] = re.sub(author_pattern, "", element.text)
+                res["author_url"] = element.a.get("href")
[…]

### 2021-09-02 cuticle_analysis/datasets/helper.py

@@ -51,14 +51,14 @@ class DatasetHelper():
 
         res = {}
         if not row['Sub-species'].isnull().any():
-            res['subspecies'] = row["Sub-species"].values[0]
+            res['subspecies'] = str(row["Sub-species"].values[0]).lower()
         if not row['Species'].isnull().any():
-            res['species'] = row["Species"].values[0]
+            res['species'] = str(row["Species"].values[0]).lower()
         if not row['Subgenus'].isnull().any():
-            res['subgenus'] = row["Subgenus"].values[0]
+            res['subgenus'] = str(row["Subgenus"].values[0]).lower()
         if not row['Genus'].isnull().any():
-            res['genus'] = row["Genus"].values[0]
+            res['genus'] = str(row["Genus"].values[0]).lower()
         if not row['Sub-Family'].isnull().any():
-            res['subfamily'] = row["Sub-Family"].values[0]
+            res['subfamily'] = str(row["Sub-Family"].values[0]).lower()
 
         return res

### 2021-09-02 scripts/images_cr.py

@@ -1,4 +1,5 @@
 
+import re
 import logging
 
 from bs4 import BeautifulSoup
@@ -14,9 +15,6 @@ logger = logging.getLogger(__name__)
 
 ids = [1]
 for _id in ids:
-    logger.info(f"Processing  ID: {_id}")
-    res = scraper.get_specimen_by_id(_id)
-    if res:
-        logger.info(res)
-    else:
-        logger.info(f"No data found for ID: {_id}")
+    logger.info(f"Processing ID: {_id}")
+    res = scraper.get_meta_by_specimen(_id)
+    logger.info(f"{res}")

### 2021-09-03 paper/paper.tex
add overall class distribution

@@
 \maketitle
-
 \section{Introduction}
@@
   gunderson_insect_1989, watson_diversity_2017}.
+
+Texture is an important feature in many applications, such as image processing,
+pattern recognition, and computer vision. Analysis of textures can be broken
+into three main categories: texture classification, texture segmentation, and
+texture synthesis. The process of classifying a texture into a set of categories
+and relies on three different approaches: feature-based, model-based, or
+structural \cite{reed_review_1993}. In this paper, we focus on a
+\textit{model-based approach} which attempt to extract parameters to reveal
+common patterns and use those parameters to automatically distinguish between
+different textures \cite{maillard_texture_2003}.
+
+Although there is some work regarding grouping ants into categories of similar
+cuticle, automated classification has yet to become an active area of research.
+Due to the large number of different ant species, the classification of ants
+into categories of similar texture is difficult to accomplish manually. Texture
+analysis has shown promising results in related fields, such as plant
+identification \cite{boudra_plant_2018}. With modern texture analysis methods,
+the classifications of ants can be automated and the results can be used to
+study the influence of cuticle texture on ant ecology.
 
@@
   \centering
+  \caption{Dataset Subclass Distribution}
+  \label{table:classdist_all_full}
+  \include{tables/all_samples_full.tex}
+\end{table}
+
+\begin{table}[h]
+  \centering
   \caption{Dataset Class Distribution}
-  \label{table:classdist}
-  \include{tables/all_samples.tex}
+  \label{table:classdist_rs_full}
+  \include{tables/rs_samples_full.tex}
 \end{table}
@@
 sculpturing identification protocol. Overall, there are a total of 2443 images
-in the dataset. The class distribution is shown in Table \ref{table:classdist}.
-The relatively small number of images in the dataset paired with the uneven
-class distribution makes it difficult to perform a meaningful classification.
-However, the goal of this project is to develop a classification algorithm that
-can be used to help classify over 200,000 ant images. Therefore, it is desirable
-to use a few representative images from each class to train the algorithm.
-Additionally, data augmentation and other methods are performed to increase the
-number of training images to create a more robust classification algorithm.
+in the dataset. The class distribution with all considered subclasses is shown
+in Table \ref{table:classdist_all_full}, and the class distribution with overall
+distribution is shown in Table \ref{table:classdist_rs_full}. The relatively
+small number of images in the dataset paired with the uneven class distribution
+makes it difficult to perform a meaningful classification. However, the goal of
+this project is to develop a classification algorithm that can be used to help
+classify over 200,000 ant images. Therefore, it is desirable to use a few
+representative images from each class to train the algorithm. Additionally, data
+augmentation and other methods are performed to increase the number of training
+images to create a more robust classification algorithm.
 

### 2021-09-03 cuticle_analysis/antweb/specimen.py
auto generate antweb figures

@@ -1,9 +1,13 @@
 
+import os
 import re
 import logging
-from typing import Dict
+from typing import Dict, List
 
+import cv2
 import requests
+import pylatex
+from pylatex.utils import escape_latex, NoEscape
 from bs4 import BeautifulSoup
 
 from ..datasets.helper import DatasetHelper
@@ -11,12 +15,29 @@ from ..datasets.helper import DatasetHelper
 logger = logging.getLogger(__name__)
 
 
+def hyperlink(url: str, text: str) -> str:
+    """
+    Helper function to create a hyperlink.
+
+    Args:
+        url: The url to link to.
+        tex: The text to be displayed in the hyperlink.
+
+    Returns:
+        The hyperlinked text.
+    """
+    text = escape_latex(text)
+    return NoEscape(r'\href{' + url + '}{' + text + '}')
+
+
 class SpecimenScraper():
     def __init__(self):
         self.dh = DatasetHelper()
-        self.base_url = "https://antweb.org"
+        self.base_url = "https://www.antweb.org"
         self.images_url = f"{self.base_url}/images.do"
         self.pic_url = f"{self.base_url}/bigPicture.do"
+        self.assets_path = "./paper/assets"
+        self.images_path = "./paper/assets/images"
 
     def get_specimen(self, info: dict) -> str:
         """
@@ -48,8 +69,9 @@ class SpecimenScraper():
         """
             Get high res image, author and author_url by id.
         """
+        specimen = self.get_specimen_by_id(_id)
         payload = {
-            "name": self.get_specimen_by_id(_id),
+            "name": specimen,
             "shot": "h",
             "number": "1"
         }
@@ -61,6 +83,8 @@ class SpecimenScraper():
         elements = span.find_all("li")
 
         res = {
+            "specimen": specimen.upper(),
+            "specimen_url": r.url,
             "author": "",
             "author_url": "",
             "image_url": ""
@@ -73,6 +97,11 @@ class SpecimenScraper():
             elif re.match(author_pattern, element.text):
                 res["author"] = re.sub(author_pattern, "", element.text)
                 res["author_url"] = element.a.get("href")
+
+        # run backup request if high res image is not available
+        if not res["author_url"]:
+            res["author_url"] = self.get_image_by_specimen(_id)
+
         return res
 
     def get_image_by_specimen(self, _id: int) -> str:
@@ -93,3 +122,50 @@ class SpecimenScraper():
 
         img = div.find("img")
         return img["src"]
+
[…]

### 2021-09-03 cuticle_analysis/datasets/__init__.py

@@ -1,5 +1,6 @@
 
 # exports
+from .helper import DatasetHelper
 from .dataset import Dataset
 from .full import FullDataset
 from .sub import SubDataset

### 2021-09-03 scripts/images_cr.py

@@ -1,8 +1,5 @@
 
-import re
 import logging
-
-from bs4 import BeautifulSoup
 
 # for debugging, env logging level must be set to debug/info
 from cuticle_analysis.core import init
@@ -10,11 +7,10 @@ init()
 
 from cuticle_analysis.antweb import SpecimenScraper  # noqa
 
-scraper = SpecimenScraper()
 logger = logging.getLogger(__name__)
 
-ids = [1]
-for _id in ids:
-    logger.info(f"Processing ID: {_id}")
-    res = scraper.get_meta_by_specimen(_id)
-    logger.info(f"{res}")
+
+if __name__ == "__main__":
+    scraper = SpecimenScraper()
+    fig = scraper.build_figure(1)
+    print("done")

### 2021-09-03 paper/paper.tex
auto generate antweb figures
add subimage dataset

@@
 into three main categories: texture classification, texture segmentation, and
-texture synthesis. The process of classifying a texture into a set of categories
-and relies on three different approaches: feature-based, model-based, or
-structural \cite{reed_review_1993}. In this paper, we focus on a
-\textit{model-based approach} which attempt to extract parameters to reveal
-common patterns and use those parameters to automatically distinguish between
-different textures \cite{maillard_texture_2003}.
+texture synthesis \cite{reed_review_1993}. The process of classifying a texture
+into a set of categories and relies on three different approaches. In this
+paper, we focus on a \textit{model-based approach} which attempts to extract
+parameters to reveal common patterns and use those parameters to automatically
+distinguish between different textures \cite{maillard_texture_2003}.
 
@@
 
+\subsection{Subimage Dataset}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\textwidth]{assets/images/segmented_1.png}
+  \caption{Hand segmented ant head image to separate head cuticle from other
+    sections of the image.
+    \href{https://www.antweb.org/bigPicture.do?name=casent0217419&shot=h&number=1}{CASENT0217419}
+    by \href{https://www.antweb.org/artist.do?id=92}{Estella Ortega}, from
+    \href{https://www.antweb.org}{AntWeb}}
+  \label{fig:segmented_1}
+\end{figure}
+
+When using the Sculpture Identification Protocol, the class of the overall image
+is based only on the texture of the head, not the eyes, antenna, or any other
+exposed cuticle. We use image segmentation to analyze the texture by splitting
+the ant head image into a set of regions: \textit{cuticle} and
+\textit{background}. A benefit of this approach allows us to pull patches or
+subimages of various sizes from the region to increase the number of samples
+available for training and at the same time, reduce the number of features to
+simplify the classification algorithm. With a given subimage size to extract
+from the segmented image, we label each subimage extracted based on the class of
+the input image. Overall, there are 73 images that were hand segmented in this
+manner and the class distribution for various sizes of subimages are shown in
+Table \ref{table:classdist_rs_sub}.
+
+\begin{table}[h]
+  \centering
+  \caption{Subimage Dataset Class Distribution}
+  \label{table:classdist_rs_sub}
+  \include{tables/rs_samples_sub.tex}
+\end{table}
+
 \section{Experimental Results}
+
+In the first experiment, we investigate the ability of a basic CNN to classify
+cuticle texture into the two main categories: \textit{rough} and
+\textit{smooth}. The goal of this experiment is to determine if more intricate
+data preprocessing is necessary to classify the texture of the ant. The dataset
+used is shown in Table \ref{table:classdist_rs_full}, and the results are shown
+in (TODO).
 

### 2021-09-03 cuticle_analysis/antweb/specimen.py
add cc attribution

@@ -33,6 +33,7 @@ def hyperlink(url: str, text: str) -> str:
 class SpecimenScraper():
     def __init__(self):
         self.dh = DatasetHelper()
+        self.cc_url = "https://creativecommons.org/licenses/by/4.0/"
         self.base_url = "https://www.antweb.org"
         self.images_url = f"{self.base_url}/images.do"
         self.pic_url = f"{self.base_url}/bigPicture.do"
@@ -159,8 +160,9 @@ class SpecimenScraper():
         specimen_str = hyperlink(info['specimen_url'], info['specimen'])
         author_str = hyperlink(info['author_url'], info['author'])
         antweb_str = hyperlink(self.base_url, "AntWeb")
+        cc_str = hyperlink(self.cc_url, "CC BY 4.0")
         fig.add_caption(NoEscape(
-            f"{specimen_str} by {author_str}, from {antweb_str}"))
+            f"{specimen_str} by {author_str}, from {antweb_str}, is licensed under {cc_str}"))
 
         # save to file
         filename = f"{info['specimen']}.tex"

### 2021-09-03 scripts/images_cr.py

@@ -13,4 +13,3 @@ logger = logging.getLogger(__name__)
 if __name__ == "__main__":
     scraper = SpecimenScraper()
     fig = scraper.build_figure(1)
-    print("done")

### 2021-09-03 paper/paper.tex
add cc attribution
add experiment 1 results

@@
     by \href{https://www.antweb.org/artist.do?id=92}{Estella Ortega}, from
-    \href{https://www.antweb.org}{AntWeb}}
+    \href{https://www.antweb.org}{AntWeb}, is licensed under
+    \href{https://creativecommons.org/licenses/by/4.0/}{CC BY 4.0}}
   \label{fig:segmented_1}
@@
 \textit{smooth}. The goal of this experiment is to determine if more intricate
-data preprocessing is necessary to classify the texture of the ant. The dataset
-used is shown in Table \ref{table:classdist_rs_full}, and the results are shown
-in (TODO).
+data preprocessing is necessary to classify the texture of the ant. Images were
+resized to different sizes for CNN input and different samples per class were
+also tested.The dataset used is shown in Table \ref{table:classdist_rs_full},
+and the results are shown in Table \ref{table:exp1_full}.
+
+\begin{table}[h]
+  \centering
+  \caption{Binary Classification of Full Size Images}
+  \label{table:exp1_full}
+  \include{tables/experiment_1.tex}
+\end{table}
 

### 2021-09-03 scripts/analyze_results.py

@@ -1,0 +1,56 @@
+
+import os
+import logging
+
+import numpy as np
+import pandas as pd
+
+# for debugging, env logging level must be set to debug/info
+from cuticle_analysis.core import init
+init()
+
+logger = logging.getLogger(__name__)
+
+size_col = "Size"
+samples_col = "Samples (per class)"
+acc_col = "Accuracy (avg)"
+loss_col = "Loss (avg)"
+
+
+def experiment_1() -> pd.DataFrame:
+    """
+        Analyze the results of experiment 1.
+    """
+    filepath = "./output/experiment_1.csv"
+    if not os.path.isfile(filepath):
+        logger.info(
+            "Error: experiment_1.csv does not exist. Skipping analysis.")
+        return
+
+    df = pd.read_csv(filepath)
+    sizes = np.sort(df["size"].unique())
+    samples_per_class = np.sort(df["samples_per_class"].unique())
+
+    res_cols = [size_col, samples_col, acc_col, loss_col]
+    res_df = pd.DataFrame(columns=res_cols)
+    for size in sizes:
+        for samples in samples_per_class:
+            temp = df[(df["size"] == size) & (
+                df["samples_per_class"] == samples)]
+            res_df = res_df.append(
+                {
+                    size_col: f"({int(size)}, {int(size)})",
+                    samples_col: samples,
+                    acc_col: round(temp["test_accuracy"].mean(), 3)*100,
+                    loss_col: round(temp["test_loss"].mean(), 3)
+                }, ignore_index=True)
+
+    return res_df
+
+
+if __name__ == "__main__":
+    e1_res = experiment_1()
+    logger.info(e1_res)
+
+    # save to latex file
+    e1_res.to_latex("./paper/tables/experiment_1.tex", index=False)

### 2021-09-03 scripts/experiments/experiment_1.py

@@ -6,31 +6,33 @@ import pandas as pd
 from cuticle_analysis.datasets import GaborRoughSmoothFull
 from cuticle_analysis.models import CNN
 
-logging.basicConfig(filename='./logs/experiment_1.log', level=logging.INFO)
 logger = logging.getLogger(__name__)
 
-# use first 500 images for dataset
-excludes = list(range(500, 2877))
-
-sizes = [64, 128, 256, 512]
-samples_per_class = [250, 500, 750]
-
-# results
 res_cols = ['size', 'samples_per_class', 'run', 'test_loss', 'test_accuracy']
-df = pd.DataFrame(columns=res_cols)
 
-for size in sizes:
-    for run in range(10):
-        # build dataset
-        data = GaborRoughSmoothFull(
-            size=(size, size), excludes=excludes, save=True)
-        for s in samples_per_class:
-            # train and evaluate model
-            model = CNN(data)
-            model.train(1, s)
-            loss, acc = model.evaluate()
 
-            # save results
-            run_df = pd.DataFrame([[size, s, loss, acc]], columns=res_cols)
-            df.append(run_df, ignore_index=True)
-            df.to_csv('./output/experiment_1.csv')
+if __name__ == "__main__":
+    # use first 500 images for dataset
+    excludes = list(range(500, 2877))
+
+    sizes = [64, 128, 256, 512]
+    samples_per_class = [250, 500, 750]
+
+    # results
+    df = pd.DataFrame(columns=res_cols)
+
+    for size in sizes:
+        for run in range(10):
+            # build dataset
+            data = GaborRoughSmoothFull(
+                size=(size, size), excludes=excludes, save=True)
+            for s in samples_per_class:
+                # train and evaluate model
+                model = CNN(data)
+                model.train(1, s)
+                loss, acc = model.evaluate()
+
+                # save results
+                run_df = pd.DataFrame([[size, s, loss, acc]], columns=res_cols)
+                df.append(run_df, ignore_index=True)
+                df.to_csv('./output/experiment_1.csv')

### 2021-09-03 paper/paper.tex
adjust experiment 1 table

@@
 
+\section{Related Work}
+
+\section{Proposed Approach}
+
 \section{Methods}
@@
 \textit{smooth}. The goal of this experiment is to determine if more intricate
-data preprocessing is necessary to classify the texture of the ant. Images were
-resized to different sizes for CNN input and different samples per class were
-also tested.The dataset used is shown in Table \ref{table:classdist_rs_full},
-and the results are shown in Table \ref{table:exp1_full}.
+data preprocessing or a more powerful model is necessary to classify the texture
+of the ant. Images were resized to different sizes for CNN input and different
+samples per class were also tested. The dataset used is shown in Table
+\ref{table:classdist_rs_full}, and the results are shown in Table
+\ref{table:exp1_full}.
 

### 2021-09-03 scripts/analyze_results.py

@@ -41,7 +41,7 @@ def experiment_1() -> pd.DataFrame:
                 {
                     size_col: f"({int(size)}, {int(size)})",
                     samples_col: samples,
-                    acc_col: round(temp["test_accuracy"].mean(), 3)*100,
+                    acc_col: f"{round(temp['test_accuracy'].mean()*100, 2)}%",
                     loss_col: round(temp["test_loss"].mean(), 3)
                 }, ignore_index=True)
 

### 2021-09-14 paper/paper.tex
jp: expand sculpturing sections

@@
 \section{Related Work}
+Taxonomists have developed extensive terminology describing ant cuticle
+sculpturing as it is often a useful diagnostic trait to distinguish between
+closely related ant species \cite{blaimer_taxonomy_2019,fisher_ants_2007}. The
+definitive text on ant cuticle terminology — \textit{The Glossary of Surface
+  Sculpturing} — contains over 100 terms to describe the cuticle sculpturing
+patterns of ants \cite{harris_glossary_1979}. There is often substantial overlap
+among terms, and closely related cuticle patterns are likely to be functionally
+similar. For example, the definition for “imbricate” is, “partly overlapping and
+appearing like shingles on a roof or scales on a fish,” which is difficult to
+distinguish from the definition of tesselate, “made up of squares like a chess
+board, either in sculpturing or in color.”
+
+Cuticle sculpturing in ants has been explored thoroughly from a taxonomic
+perspective; however, the function of nano and microstructures on insect
+exoskeletons is a developing topic in entomology. Watson et al. reviews the
+literature of cuticle nano and microstructure function and proposes 21 possible
+functions associated with these structures \cite{watson_diversity_2017}. Many of
+these functions are related to structures not found in ants such as scales and
+nanostructures. The review does include functions that may relate to ants such
+as friction control, enhanced surface area, and increased hardness.  Watson et
+al. also describe seven types of cuticle structures ranging from hairs and
+scales to nano and micro structures \cite{watson_diversity_2017}. The cuticle
+sculpturing of ants seems to fall within one type – complex microstructures.
+
+The five broad functional groupings developed that describes the complex
+microstructures found on ants were derived from reviewing the variation of
+cuticle sculpturing across ants. These functional groupings were also reflected
+in Harris as many terms could be grouped together based on similar definitions
+and comparing the SEM photographs provided in the publication
+\cite{harris_glossary_1979}.
 
@@
 \subsection{Sculpture Identification Protocol}
+A team of three assistants were in charge of the manual identification process.
+The team was trained to identify cuticle sculpturing through a process which
+consisted of one 45-minute introductory lesson explaining the project and
+texture categories and then given a training set of photos to identify from the
+genus \textit{Polyrhachis}, whose members display among the highest diversity of
+cuticle patterns all categories. The training set identifications were reviewed
+together as a group. Once training was complete, assistants were assigned the
+same genera of ants to identify independently each week. A weekly meeting was
+held to discuss identifications and assign new ones. These identifications were
+collected in a spreadsheet and the identifications were assigned to individual
+ant species on a majority basis.
 
@@
 
-\begin{table}[h]
+\begin{table}[ht]
   \centering
@@
 
-Images were collected and classified by undergraduate students according to the
-sculpturing identification protocol. Overall, there are a total of 2443 images
-in the dataset. The class distribution with all considered subclasses is shown
-in Table \ref{table:classdist_all_full}, and the class distribution with overall
+Images were collected and classified by according to the sculpturing
+identification protocol. Overall, there are a total of 2443 images in the
+dataset. The class distribution with all considered subclasses is shown in Table
+\ref{table:classdist_all_full}, and the class distribution with overall
 distribution is shown in Table \ref{table:classdist_rs_full}. The relatively
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
@@
 \section{Experimental Results}
+
+\subsection{Sculpturing Identification}
+In our initial trial, we examined the entire \textit{Polyrhachis} genus, which
+was 581 eligible specimens out of the 775 species. 71\% of the classifications
+were unanimous. We then expanded our categorizations to 2846 images of
+individual species from 13 genera. These genera were from the four major ant
+subfamilies Ponerinae, Myrmicinae, Formicinae, and Dolichoderinae along with
+Dorylinus and Ectatomminae. 87\% (2486) of classifications were unanimous. This
+percentage represents the accuracy for the entirety of identifications starting
+November 2020 and ending in March 2021. Assistants were also asked to complete
+an ant cuticle assessment near the end of the identification process. The
+assessment consisted of images from forty species of ants that spanned all
+categories. The purpose of the assessment was to measure the consistency of
+identifications while removing any biases from weekly meeting discussions and
+cross referencing of other identifications. Unanimous classification was reached
+90\% of the time during the individual assessment.
+
+Classifications became more consistent overtime. One likely reason for this
+increase was that students gained additional experience in cuticle
+classification over the course of these trials. However, definitions of cuticle
+texture categories were refined overtime. At the inception of the project,
+assistants based classifications off of reference photos and tentative
+definitions that were sometimes unclear. For example, striate was initially
+defined as “long linear ridges, corresponding in-between ridges vary in length.”
+Also, some tentative categories had low agreement amongst the team. For example,
+striate initially was a broad category containing two subclasses (Severe and
+Mild) and encompassed longitudinal and latitudinal ridges. Based on several
+rounds of feedback from students and identification of these problem categories
+where there was low agreement, the definitions developed into the current
+classification system.
+
+\subsection{Binary Classification}
 

### 2021-09-14 cuticle_analysis/core/__init__.py

@@ -42,5 +42,10 @@ def start_app():
         download_dataset()
         unzip_dataset()
 
+    elif args.cite:
+        from ..antweb import SpecimenScraper
+        scraper = SpecimenScraper()
+        scraper.build_figure(args.cite)
+
     else:
         dataset_setup()

### 2021-09-16 cuticle_analysis/models/cnn.py
make base classes for experiments and models

@@ -1,5 +1,5 @@
 
-
+import logging
 from typing import List, Tuple
 
 from tensorflow.keras.models import Sequential
@@ -8,21 +8,20 @@ from tensorflow import keras
 import numpy as np
 
 from ..datasets import Dataset
+from .model import Model
+
+logger = logging.getLogger(__name__)
 
 
-class CNN():
+class CNN(Model):
     def __init__(self, data: Dataset):
-        self.data = data
-        self.name = 'CNN'
-        self.num_classes = len(np.unique(data.labels))
-        self.subimage_size = data.size
-        self.path = f'./output/model_{data.size[0]}_{data.size[1]}'
+        super().__init__('CNN', data)
 
         self.model = Sequential([
             layers.experimental.preprocessing.Rescaling(
                 1./255,
                 input_shape=(
-                    self.subimage_size[0], self.subimage_size[1], 3)),
+                    self.size[0], self.size[1], 3)),
             layers.Conv2D(16, 3, padding='same', activation='relu'),
             layers.MaxPooling2D(),
             layers.Conv2D(32, 3, padding='same', activation='relu'),
@@ -34,13 +33,8 @@ class CNN():
             layers.Dense(self.num_classes)
         ])
 
-    def metadata(self) -> List[str]:
-        return [
-            f'Model Type: {self.name}'
-        ]
-
-    def train(self, epochs: int, n: int) -> None:
-        train_x, train_y = self.data.stratified_split(n)
+    def train(self, epochs: int, samples_per_class: int) -> None:
+        train_x, train_y = self.data.stratified_split(samples_per_class)
         val_x, val_y = self.data.build_validation_set()
 
         self.model.compile(
@@ -56,12 +50,6 @@ class CNN():
             validation_data=(val_x, val_y),
             epochs=epochs
         )
-
-    def save_weights(self) -> None:
-        self.model.save(self.path)
-
-    def load_weights(self) -> None:
-        self.model = keras.models.load_model(self.path)
 
     def predict(self, image: np.ndarray) -> np.ndarray:
         pred = self.model.predict(image)

### 2021-09-16 scripts/experiments/base.py

@@ -1,0 +1,38 @@
+
+import logging
+from typing import Type, List
+
+import pandas as pd
+from cuticle_analysis.datasets import Dataset
+from cuticle_analysis.models import Model
+
+logger = logging.getLogger(__name__)
+
+res_cols = ['size', 'samples_per_class', 'run', 'test_loss', 'test_accuracy']
+
+
+class ExperimentRunner:
+    def __init__(self, name: str, dataset_type: Type[Dataset], model_type: Type[Model]):
+        self.name = name
+        self.dataset_type = dataset_type
+        self.model_type = model_type
+
+    def run(self,
+            samples: int,
+            size: tuple,
+            runs: int = 10,
+            epochs: int = 10,
+            excludes: list = []):
+        dataset = self.dataset_type(size=size, excludes=excludes, save=True)
+        res_df = pd.DataFrame(columns=res_cols)
+
+        for run in range(runs):
+            model = self.model_type(dataset)
+            model.train(epochs, samples)
+            loss, acc = model.evaluate()
+
+            # save results
+            run_df = pd.DataFrame(
+                [[size, samples, run, loss, acc]], columns=res_cols)
+            res_df = res_df.append(run_df, ignore_index=True)
+            res_df.to_csv(f'./output/{self.name}')

### 2021-09-16 scripts/experiments/experiment_1.py

@@ -1,38 +1,17 @@
 
 import logging
 
-import pandas as pd
-
 from cuticle_analysis.datasets import GaborRoughSmoothFull
 from cuticle_analysis.models import CNN
 
+import base
+
 logger = logging.getLogger(__name__)
 
-res_cols = ['size', 'samples_per_class', 'run', 'test_loss', 'test_accuracy']
+sizes = [64, 128, 256, 512]
+samples_per_class = [250, 500, 750]
+exp = base.ExperimentRunner('experiment_1', GaborRoughSmoothFull, CNN)
 
-
-if __name__ == "__main__":
-    # use first 500 images for dataset
-    excludes = list(range(500, 2877))
-
-    sizes = [64, 128, 256, 512]
-    samples_per_class = [250, 500, 750]
-
-    # results
-    df = pd.DataFrame(columns=res_cols)
-
-    for size in sizes:
-        for run in range(10):
-            # build dataset
-            data = GaborRoughSmoothFull(
-                size=(size, size), excludes=excludes, save=True)
-            for s in samples_per_class:
-                # train and evaluate model
-                model = CNN(data)
-                model.train(1, s)
-                loss, acc = model.evaluate()
-
-                # save results
-                run_df = pd.DataFrame([[size, s, loss, acc]], columns=res_cols)
-                df.append(run_df, ignore_index=True)
-                df.to_csv('./output/experiment_1.csv')
+for size in sizes:
+    for samples in samples_per_class:
+        exp.run(samples, size, runs=10, epochs=20, excludes=[])

### 2021-09-16 cuticle_analysis/antweb/specimen.py
drop unused imports

@@ -2,7 +2,7 @@
 import os
 import re
 import logging
-from typing import Dict, List
+from typing import Dict
 
 import cv2
 import requests

### 2021-09-16 cuticle_analysis/core/__init__.py

@@ -1,7 +1,6 @@
 
 import logging
 import os
-import sys
 
 import colorama
 from colorama import Fore

### 2021-09-16 cuticle_analysis/datasets/dataset.py

@@ -1,8 +1,7 @@
 
 import logging
-from typing import Dict, List, Tuple
+from typing import Dict, Tuple
 
-import cv2
 import pandas as pd
 import numpy as np
 from numpy.random import default_rng

### 2021-09-16 cuticle_analysis/models/cnn.py

@@ -1,6 +1,6 @@
 
 import logging
-from typing import List, Tuple
+from typing import Tuple
 
 from tensorflow.keras.models import Sequential
 from tensorflow.keras import layers

### 2021-09-16 scripts/experiments/base.py

@@ -1,6 +1,6 @@
 
 import logging
-from typing import Type, List
+from typing import Type
 
 import pandas as pd
 from cuticle_analysis.datasets import Dataset

### 2021-09-16 setup.py

@@ -1,6 +1,6 @@
 # -*- coding: utf-8 -*-
 
-from setuptools import setup, find_packages
+from setuptools import setup
 
 with open('README.md') as f:
     readme = f.read()

### 2021-09-16 cuticle_analysis/antweb/specimen.py
Paper building (#13)
* add subimage dataset

* add cc attribution

* add experiment 1 results

* adjust experiment 1 table

* jp: expand sculpturing sections

* update workflow (#11) (#12)

* update workflow

* don't use dev libraries for build/test

* drop unstable, update readme

* remove redundant ignore

* proper antweb citation

* make base classes for experiments and models

* drop unused imports

@@ -2,7 +2,7 @@
 import os
 import re
 import logging
-from typing import Dict, List
+from typing import Dict
 
 import cv2
 import requests
@@ -33,6 +33,7 @@ def hyperlink(url: str, text: str) -> str:
 class SpecimenScraper():
     def __init__(self):
         self.dh = DatasetHelper()
+        self.cc_url = "https://creativecommons.org/licenses/by/4.0/"
         self.base_url = "https://www.antweb.org"
         self.images_url = f"{self.base_url}/images.do"
         self.pic_url = f"{self.base_url}/bigPicture.do"
@@ -159,8 +160,9 @@ class SpecimenScraper():
         specimen_str = hyperlink(info['specimen_url'], info['specimen'])
         author_str = hyperlink(info['author_url'], info['author'])
         antweb_str = hyperlink(self.base_url, "AntWeb")
+        cc_str = hyperlink(self.cc_url, "CC BY 4.0")
         fig.add_caption(NoEscape(
-            f"{specimen_str} by {author_str}, from {antweb_str}"))
+            f"{specimen_str} by {author_str}, from {antweb_str}, is licensed under {cc_str}"))
 
         # save to file
         filename = f"{info['specimen']}.tex"

### 2021-09-16 cuticle_analysis/const.py

@@ -36,8 +36,7 @@ ALL_LABEL_MAP = {
     'Rough Netted': 2,
     'Rough Ridged': 3,
     'Rough Tuberous': 4,
-    'Smooth Gritty': 5,
-    'Smooth': 6,
+    'Smooth': 5,
 }
 
 INT_ALL_LABEL_MAP = {v: k for k, v in ALL_LABEL_MAP.items()}

### 2021-09-16 cuticle_analysis/core/arguments.py

@@ -44,7 +44,9 @@ class Args:
 _builder = Args(
     group=[
         Args('--download-dataset', action='store_true',
-             help='download and unzip dataset')
+             help='download and unzip dataset'),
+        # int of id to store
+        Args('--cite', type=int, default=0,)
     ],
 )
 

### 2021-09-16 cuticle_analysis/datasets/utils.py

@@ -82,8 +82,9 @@ def convert_labels(label: pd.Series) -> pd.Series:
         to_replace=r'^[r][r].*', value=2, regex=True)
     label = label.replace(
         to_replace=r'^[r][t].*', value=3, regex=True)
+    # smooth gritty to rough tuberous
     label = label.replace(
-        to_replace=r'^[s][g].*', value=4, regex=True)
+        to_replace=r'^[s][g].*', value=3, regex=True)
     label = label.replace(
         to_replace=r'^[s][s].*', value=5, regex=True)
 

### 2021-09-16 cuticle_analysis/models/model.py

@@ -1,0 +1,36 @@
+
+from typing import List, Tuple
+
+from tensorflow import keras
+import numpy as np
+
+from ..datasets import Dataset
+
+
+class Model:
+    def __init__(self, name: str, data: Dataset):
+        self.name = name
+        self.data = data
+        self.size = data.size
+        self.num_classes = len(np.unique(data.labels))
+        self.path = f'./output/model_{name}_{data.size[0]}_{data.size[1]}'
+
+    def metadata(self) -> List[str]:
+        return [
+            f'Model Type: {self.name}'
+        ]
+
+    def save_weights(self) -> None:
+        self.model.save(self.path)
+
+    def load_weights(self) -> None:
+        self.model = keras.models.load_model(self.path)
+
+    def train(self, epochs: int, samples_per_class: int) -> None:
+        raise NotImplementedError()
+
+    def predict(self, image: np.ndarray) -> np.ndarray:
+        raise NotImplementedError()
+
+    def evaluate(self) -> Tuple[float, float]:
+        raise NotImplementedError()

### 2021-09-16 scripts/analyze_results.py

@@ -1,0 +1,56 @@
+
+import os
+import logging
+
+import numpy as np
+import pandas as pd
+
+# for debugging, env logging level must be set to debug/info
+from cuticle_analysis.core import init
+init()
+
+logger = logging.getLogger(__name__)
+
+size_col = "Size"
+samples_col = "Samples (per class)"
+acc_col = "Accuracy (avg)"
+loss_col = "Loss (avg)"
+
+
+def experiment_1() -> pd.DataFrame:
+    """
+        Analyze the results of experiment 1.
+    """
+    filepath = "./output/experiment_1.csv"
+    if not os.path.isfile(filepath):
+        logger.info(
+            "Error: experiment_1.csv does not exist. Skipping analysis.")
+        return
+
+    df = pd.read_csv(filepath)
+    sizes = np.sort(df["size"].unique())
+    samples_per_class = np.sort(df["samples_per_class"].unique())
+
+    res_cols = [size_col, samples_col, acc_col, loss_col]
+    res_df = pd.DataFrame(columns=res_cols)
+    for size in sizes:
+        for samples in samples_per_class:
+            temp = df[(df["size"] == size) & (
+                df["samples_per_class"] == samples)]
+            res_df = res_df.append(
+                {
+                    size_col: f"({int(size)}, {int(size)})",
+                    samples_col: samples,
+                    acc_col: f"{round(temp['test_accuracy'].mean()*100, 2)}%",
+                    loss_col: round(temp["test_loss"].mean(), 3)
+                }, ignore_index=True)
+
+    return res_df
+
+
+if __name__ == "__main__":
+    e1_res = experiment_1()
+    logger.info(e1_res)
+
+    # save to latex file
+    e1_res.to_latex("./paper/tables/experiment_1.tex", index=False)

### 2021-09-16 scripts/data_spread.py

@@ -1,6 +1,5 @@
 
 from typing import Dict, List
-import gc
 import logging
 
 import pandas as pd
@@ -10,6 +9,7 @@ from cuticle_analysis.core import init
 init()
 
 from cuticle_analysis import const  # noqa
+from cuticle_analysis.datasets import Dataset  # noqa
 from cuticle_analysis.datasets import AllFull  # noqa
 from cuticle_analysis.datasets import RoughSmoothFull  # noqa
 from cuticle_analysis.datasets import RoughSmoothSub  # noqa
@@ -25,13 +25,16 @@ rs_labels.remove('_background_')
 _TOTAL = 'Total'
 _TYPE = 'Type'
 
-all_cols = [_TYPE] + all_labels + [_TOTAL]
-rs_cols = [_TYPE] + rs_labels + [_TOTAL]
-
 
 def total_samples(class_data: Dict) -> int:
     """
-    Total number of samples in the dataset.
+        Total number of samples in the dataset.
+
+        Args:
+            class_data: Dictionary of class labels and number of samples.
+
+        Returns:
+            total: Total number of samples per class label.
     """
     res = 0
     for _, data in class_data.items():
@@ -39,49 +42,48 @@ def total_samples(class_data: Dict) -> int:
     return res
 
 
-def all_row(type: str, class_data: Dict) -> pd.DataFrame:
-    data = [class_data[label] for label in all_labels]
-    cols = [type] + data + [total_samples(class_data)]
-    row = pd.DataFrame(
-        [cols], columns=all_cols)
-    return row
+def append_all(df: pd.DataFrame, data: Dataset, cols: List, _type: str) -> pd.DataFrame:
+    """
+        Build a all class distribution row from an input dataset and append the
+        row to a dataframe.
+    """
+    def all_row(type: str, class_data: Dict, _cols: List) -> pd.DataFrame:
+        """
+            Build a row for the all class distribution table.
+        """
+        data = [class_data[label] for label in all_labels]
+        cols = [type] + data + [total_samples(class_data)]
+        row = pd.DataFrame(
+            [cols], columns=_cols)
+        return row
+
+    row = all_row(_type, data.class_data(), cols)
+    df = df.append(row, ignore_index=True)
+
+    return df
 
 
[…]

### 2021-09-16 scripts/experiments/base.py

@@ -1,0 +1,38 @@
+
+import logging
+from typing import Type
+
+import pandas as pd
+from cuticle_analysis.datasets import Dataset
+from cuticle_analysis.models import Model
+
+logger = logging.getLogger(__name__)
+
+res_cols = ['size', 'samples_per_class', 'run', 'test_loss', 'test_accuracy']
+
+
+class ExperimentRunner:
+    def __init__(self, name: str, dataset_type: Type[Dataset], model_type: Type[Model]):
+        self.name = name
+        self.dataset_type = dataset_type
+        self.model_type = model_type
+
+    def run(self,
+            samples: int,
+            size: tuple,
+            runs: int = 10,
+            epochs: int = 10,
+            excludes: list = []):
+        dataset = self.dataset_type(size=size, excludes=excludes, save=True)
+        res_df = pd.DataFrame(columns=res_cols)
+
+        for run in range(runs):
+            model = self.model_type(dataset)
+            model.train(epochs, samples)
+            loss, acc = model.evaluate()
+
+            # save results
+            run_df = pd.DataFrame(
+                [[size, samples, run, loss, acc]], columns=res_cols)
+            res_df = res_df.append(run_df, ignore_index=True)
+            res_df.to_csv(f'./output/{self.name}')

### 2021-09-16 scripts/experiments/experiment_1.py

@@ -1,36 +1,17 @@
 
 import logging
 
-import pandas as pd
-
 from cuticle_analysis.datasets import GaborRoughSmoothFull
 from cuticle_analysis.models import CNN
 
-logging.basicConfig(filename='./logs/experiment_1.log', level=logging.INFO)
-logger = logging.getLogger(__name__)
+import base
 
-# use first 500 images for dataset
-excludes = list(range(500, 2877))
+logger = logging.getLogger(__name__)
 
 sizes = [64, 128, 256, 512]
 samples_per_class = [250, 500, 750]
-
-# results
-res_cols = ['size', 'samples_per_class', 'run', 'test_loss', 'test_accuracy']
-df = pd.DataFrame(columns=res_cols)
+exp = base.ExperimentRunner('experiment_1', GaborRoughSmoothFull, CNN)
 
 for size in sizes:
-    for run in range(10):
-        # build dataset
-        data = GaborRoughSmoothFull(
-            size=(size, size), excludes=excludes, save=True)
-        for s in samples_per_class:
-            # train and evaluate model
-            model = CNN(data)
-            model.train(1, s)
-            loss, acc = model.evaluate()
-
-            # save results
-            run_df = pd.DataFrame([[size, s, loss, acc]], columns=res_cols)
-            df.append(run_df, ignore_index=True)
-            df.to_csv('./output/experiment_1.csv')
+    for samples in samples_per_class:
+        exp.run(samples, size, runs=10, epochs=20, excludes=[])

### 2021-09-16 scripts/gen_assests.sh

@@ -2,5 +2,6 @@
 
 nix-shell --run "
 cuticle_analysis --download-dataset \
-&& python scripts/data_spread.py
+&& python scripts/data_spread.py \
+&& python scripts/analyze_results.py
 "

### 2021-09-16 cuticle_analysis/antweb/specimen.py
fix base classes, experiment_1 rerun in progress

@@ -2,7 +2,7 @@
 import os
 import re
 import logging
-from typing import Dict
+from typing import Dict, Tuple
 
 import cv2
 import requests
@@ -13,6 +13,9 @@ from bs4 import BeautifulSoup
 from ..datasets.helper import DatasetHelper
 
 logger = logging.getLogger(__name__)
+
+ANT_WEB_URL = "https://antweb.org/"
+CC_URL = "https://creativecommons.org/licenses/by/4.0/"
 
 
 def hyperlink(url: str, text: str) -> str:
@@ -33,12 +36,13 @@ def hyperlink(url: str, text: str) -> str:
 class SpecimenScraper():
     def __init__(self):
         self.dh = DatasetHelper()
-        self.cc_url = "https://creativecommons.org/licenses/by/4.0/"
-        self.base_url = "https://www.antweb.org"
         self.images_url = f"{self.base_url}/images.do"
         self.pic_url = f"{self.base_url}/bigPicture.do"
         self.assets_path = "./paper/assets"
         self.images_path = "./paper/assets/images"
+
+        self.antweb_str = hyperlink(ANT_WEB_URL, "AntWeb")
+        self.cc_str = hyperlink(CC_URL, "CC BY 4.0")
 
     def get_specimen(self, info: dict) -> str:
         """
@@ -142,7 +146,7 @@ class SpecimenScraper():
         # remove tif
         os.remove(f"{path}.tif")
 
-    def build_figure(self, _id: int):
+    def build_figure(self, _id: int) -> Tuple[Dict, pylatex.Figure]:
         """
             Build latex figure with included image and CC attribution from a
             list of given ids
@@ -159,11 +163,14 @@ class SpecimenScraper():
                       width=NoEscape(r'0.2\textwidth'))
         specimen_str = hyperlink(info['specimen_url'], info['specimen'])
         author_str = hyperlink(info['author_url'], info['author'])
-        antweb_str = hyperlink(self.base_url, "AntWeb")
-        cc_str = hyperlink(self.cc_url, "CC BY 4.0")
-        fig.add_caption(NoEscape(
-            f"{specimen_str} by {author_str}, from {antweb_str}, is licensed under {cc_str}"))
+        fig.add_caption(
+            NoEscape(str(
+                specimen_str
+                + " by " + author_str
+                + ", from " + self.antweb_str
+                + ", is licensed under " + self.cc_str)))
 
+    def save_figure(self, info: dict, fig: pylatex.Figure):
[…]

### 2021-09-16 cuticle_analysis/models/cnn.py

@@ -1,19 +1,16 @@
 
 import logging
-from typing import Tuple
 
 from tensorflow.keras.models import Sequential
 from tensorflow.keras import layers
-from tensorflow import keras
-import numpy as np
 
 from ..datasets import Dataset
-from .model import Model
+from .model import TFModel
 
 logger = logging.getLogger(__name__)
 
 
-class CNN(Model):
+class CNN(TFModel):
     def __init__(self, data: Dataset):
         super().__init__('CNN', data)
 
@@ -32,31 +29,3 @@ class CNN(Model):
             layers.Dense(128, activation='relu'),
             layers.Dense(self.num_classes)
         ])
-
-    def train(self, epochs: int, samples_per_class: int) -> None:
-        train_x, train_y = self.data.stratified_split(samples_per_class)
-        val_x, val_y = self.data.build_validation_set()
-
-        self.model.compile(
-            optimizer='adam',
-            loss=keras.losses.SparseCategoricalCrossentropy(
-                from_logits=True),
-            metrics=['accuracy'])
-        self.model.summary()
-        self.epochs = epochs
-
-        self.history = self.model.fit(
-            train_x, train_y,
-            validation_data=(val_x, val_y),
-            epochs=epochs
-        )
-
-    def predict(self, image: np.ndarray) -> np.ndarray:
-        pred = self.model.predict(image)
-        pred = np.argmax(pred, axis=1)
-        return pred
-
-    def evaluate(self) -> Tuple[float, float]:
-        test_x, test_y = self.data.test_set()
-        loss, acc = self.model.evaluate(test_x, test_y, batch_size=128)
-        return loss, acc

### 2021-09-16 cuticle_analysis/models/model.py

@@ -1,5 +1,5 @@
 
-from typing import List, Tuple
+from typing import List
 
 from tensorflow import keras
 import numpy as np
@@ -21,16 +21,55 @@ class Model:
         ]
 
     def save_weights(self) -> None:
+        raise NotImplementedError
+
+    def load_weights(self) -> None:
+        raise NotImplementedError
+
+    def train(self, epochs: int, samples_per_class: int) -> None:
+        raise NotImplementedError
+
+
+class TFModel(Model):
+    def save_weights(self) -> None:
         self.model.save(self.path)
 
     def load_weights(self) -> None:
         self.model = keras.models.load_model(self.path)
 
     def train(self, epochs: int, samples_per_class: int) -> None:
-        raise NotImplementedError()
+        train_x, train_y = self.data.stratified_split(samples_per_class)
+        val_x, val_y = self.data.build_validation_set()
+
+        # compile
+        self.model.compile(
+            optimizer='adam',
+            loss='categorical_crossentropy',
+            metrics=[
+                'accuracy',
+                'mean_squared_error',
+                'categorical_crossentropy',
+                'precision',
+                'recall',
+            ]
+        )
+        self.model.summary()
+
+        # fit
+        self.epochs = epochs
+        self.history = self.model.fit(
+            train_x,
+            train_y,
+            epochs=epochs,
+            validation_data=(val_x, val_y),
+            verbose=1
+        )
 
     def predict(self, image: np.ndarray) -> np.ndarray:
-        raise NotImplementedError()
+        pred = self.model.predict(image)
+        pred = np.argmax(pred, axis=1)
+        return pred
 
-    def evaluate(self) -> Tuple[float, float]:
-        raise NotImplementedError()
+    def evaluate(self):
+        test_x, test_y = self.data.test_set()
+        return self.model.evaluate(test_x, test_y, batch_size=128)

### 2021-09-16 scripts/experiments/experiment.py

@@ -3,12 +3,17 @@ import logging
 from typing import Type
 
 import pandas as pd
-from cuticle_analysis.datasets import Dataset
-from cuticle_analysis.models import Model
+
+# for debugging, env logging level must be set to debug/info
+from cuticle_analysis.core import init
+init()
+
+from cuticle_analysis.datasets import Dataset  # noqa
+from cuticle_analysis.models import Model  # noqa
 
 logger = logging.getLogger(__name__)
 
-res_cols = ['size', 'samples_per_class', 'run', 'test_loss', 'test_accuracy']
+RES_COLS = ['size', 'samples_per_class', 'run', 'test_loss', 'test_accuracy']
 
 
 class ExperimentRunner:
@@ -24,7 +29,7 @@ class ExperimentRunner:
             epochs: int = 10,
             excludes: list = []):
         dataset = self.dataset_type(size=size, excludes=excludes, save=True)
-        res_df = pd.DataFrame(columns=res_cols)
+        res_df = pd.DataFrame(columns=RES_COLS)
 
         for run in range(runs):
             model = self.model_type(dataset)
@@ -33,6 +38,6 @@ class ExperimentRunner:
 
             # save results
             run_df = pd.DataFrame(
-                [[size, samples, run, loss, acc]], columns=res_cols)
+                [[size, samples, run, loss, acc]], columns=RES_COLS)
             res_df = res_df.append(run_df, ignore_index=True)
             res_df.to_csv(f'./output/{self.name}')

### 2021-09-16 cuticle_analysis/datasets/dataset.py
experiment 1 running

@@ -257,7 +257,7 @@ class Dataset(DatasetHelper):
     def build_validation_set(self, split: float = 0.5) -> Tuple[np.ndarray, np.ndarray]:
         try:
             # see if test_y exists
-            self.train_y
+            self.test_y
         except Exception as e:
             msg = "Test samples have yet to be made."
             logger.error(msg)
@@ -272,10 +272,10 @@ class Dataset(DatasetHelper):
         test_idxs = np.array([idx for idx in range(
             len(self.test_y)) if idx not in val_idxs])
 
-        self.test_x = self.test_x[(test_idxs)]
-        self.test_y = self.test_y[(test_idxs)]
         self.val_x = self.test_x[(val_idxs)]
         self.val_y = self.test_y[(val_idxs)]
+        self.test_x = self.test_x[(test_idxs)]
+        self.test_y = self.test_y[(test_idxs)]
 
         return self.val_x, self.val_y
 

### 2021-09-16 cuticle_analysis/datasets/helper.py

@@ -19,7 +19,7 @@ class DatasetHelper():
     def __init__(self):
         self.ant_data = pd.read_excel(f'./dataset/labels.xlsx', header=0)
 
-    def get_image(self, _id: int) -> np.ndarray:
+    def get_image(self, _id: int, grayscale: bool = True) -> np.ndarray:
         """Get image by ID.
 
         Args:
@@ -35,6 +35,9 @@ class DatasetHelper():
             msg = f'Failed to open image {path}'
             logger.error(msg)
             raise ValueError(msg)
+
+        if grayscale:
+            img = cv2.cvtColor(img, cv2.COLOR_BGR2GRAY)
 
         return img
 

### 2021-09-16 cuticle_analysis/models/model.py

@@ -1,8 +1,9 @@
 
 from typing import List
 
-from tensorflow import keras
 import numpy as np
+import tensorflow as tf
+from tensorflow import keras
 
 from ..datasets import Dataset
 
@@ -44,13 +45,10 @@ class TFModel(Model):
         # compile
         self.model.compile(
             optimizer='adam',
-            loss='categorical_crossentropy',
+            loss='sparse_categorical_crossentropy',
             metrics=[
                 'accuracy',
-                'mean_squared_error',
-                'categorical_crossentropy',
-                'precision',
-                'recall',
+                tf.keras.metrics.MeanSquaredError(),
             ]
         )
         self.model.summary()

### 2021-09-16 scripts/experiments/experiment.py

@@ -13,7 +13,8 @@ from cuticle_analysis.models import Model  # noqa
 
 logger = logging.getLogger(__name__)
 
-RES_COLS = ['size', 'samples_per_class', 'run', 'test_loss', 'test_accuracy']
+RES_COLS = ['size', 'samples_per_class', 'run',
+            'loss', 'accuracy', 'error (mse)']
 
 
 class ExperimentRunner:
@@ -21,6 +22,13 @@ class ExperimentRunner:
         self.name = name
         self.dataset_type = dataset_type
         self.model_type = model_type
+        self.res_df = pd.DataFrame(columns=RES_COLS)
+
+    def append_df(self, df: pd.DataFrame):
+        self.res_df = self.res_df.append(df, ignore_index=True)
+
+    def save_df(self):
+        self.res_df.to_csv(f'./output/{self.name}.csv')
 
     def run(self,
             samples: int,
@@ -29,15 +37,21 @@ class ExperimentRunner:
             epochs: int = 10,
             excludes: list = []):
         dataset = self.dataset_type(size=size, excludes=excludes, save=True)
-        res_df = pd.DataFrame(columns=RES_COLS)
 
         for run in range(runs):
             model = self.model_type(dataset)
             model.train(epochs, samples)
-            loss, acc = model.evaluate()
+            loss, acc, err = model.evaluate()
 
             # save results
             run_df = pd.DataFrame(
-                [[size, samples, run, loss, acc]], columns=RES_COLS)
-            res_df = res_df.append(run_df, ignore_index=True)
-            res_df.to_csv(f'./output/{self.name}')
+                [[
+                    size,
+                    samples,
+                    run,
+                    loss,
+                    acc,
+                    err
+                ]], columns=RES_COLS)
+            self.append_df(run_df)
+            self.save_df()

### 2021-09-19 GUI/objects/image_viewer.py
Removed image_id variable on GUI_Main to reduce class coupling between imageviewer and GUI main class. Modified functions in imageviewer class to remove dependency on GUI_Main class to change images.

@@ -22,8 +22,8 @@ class ImageViewer():
         self.bkg_color = (50, 50, 50)
         self.bkg_obj = pygame.Rect(position[0], position[1], size[0], size[1])
         self.size = size
-        self.increment_id = False
-        self.decrement_id = False
+        self.increment_flag = False
+        self.decrement_flag = False
 
     def __scale_image__(self):
         """
@@ -62,32 +62,40 @@ class ImageViewer():
 
     def __get_increment_flag__(self):
         "Returns the increment id flag for the main thread to increment image id."
-        return self.increment_id
+        return self.increment_flag
 
     def __get_decrement_flag__(self):
         "Returns the increment id flag for the main thread to decrement image id."
-        return self.decrement_id
+        return self.decrement_flag
 
-    def __set_increment_flag__(self, new_flag):
+    def __increment_image__(self):
         "Sets the increment id flag for the main thread to increment image id."
-        self.increment_id = new_flag
+        if self.id < 2876:
+            self.id = self.id + 1
 
-    def __set_decrement_flag__(self, new_flag):
+    def __decrement_image__(self):
         "Sets the decrement id flag for the main thread to decrement image id."
-        self.decrement_id = new_flag
+        if self.id > 1:
+            self.id -= 1
 
-    def __update_image__(self, id):
+    def __correct_color__(self):
+        """Corrects the color of the image by swapping blue and red color values."""
+        print(str(self.image_arr[0][0]))
+
+    def __update_image__(self, id=None):
         """
             Updates the image viewer to current image, which is an image cache.
             Post-condition: Ant image is shown in image viewer. If file not found,
             The cache image is deleted and the image viewer shows error.
         """
         #IMPLEMENT FUNCTION THAT OBTAINS THE MAXIMUM IMAGE ID.
-        if id >= 1 and id <= 2876:
-            self.id = id
+        if (id >= 1 and id <= 2876) or id is None:
+            if id is not None:
+                self.id = id
             try:
                 self.image_arr = im.fromarray(
                     self.image_list.get_image(self.id))
+                self.__correct_color__()
                 self.image_arr.save(self.cache_path)
                 return 1
             except Exception as e:
@@ -104,3 +112,8 @@ class ImageViewer():
         except Exception as e:
             print("Cache deleted.")
         self.image = None
+
+    
+    def get_image_id(self):
+        """Returns the image id."""
+        return self.id

### 2021-09-19 GUI_main.py

@@ -46,8 +46,6 @@ def start():
         (const.WINDOW_SIZE[0], const.WINDOW_SIZE[1]), HIDDEN)
     main = Gui(window)
     white = (255, 255, 255)
-    __image_id__ = 1
-    main.set_caption(str(__image_id__) + ".jpg")
     main.get_surface().fill(white)
     pygame.mouse.set_visible(1)
     body_font = pygame.font.SysFont('Arial', const.BODY_FONT_SIZE)
@@ -70,11 +68,12 @@ def start():
                               "width": 50, "height": 50}, [prev_bttn_pos[0], prev_bttn_pos[1], 50])
     next_button = Buttons(main.get_surface(), "rectangle", ">", const.BUTTON_COLOR, {
                           "width": 50, "height": 50}, [next_bttn_pos[0], next_bttn_pos[1], 50])
-    id_text = body_font.render(str(__image_id__), True, (0, 0, 0))
     ant_iv = ImageViewer(main.get_surface(), data,
                          'rough_smooth', (225, 100), (350, 350))
-    next_button.on_click(lambda: ant_iv.__set_increment_flag__(True))
-    previous_button.on_click(lambda: ant_iv.__set_decrement_flag__(True))
+    id_text = body_font.render(str(ant_iv.get_image_id()), True, (0, 0, 0))
+    main.set_caption(str(ant_iv.get_image_id()) + ".jpg")
+    next_button.on_click(lambda: ant_iv.__increment_image__())
+    previous_button.on_click(lambda: ant_iv.__decrement_image__())
     ant_iv.__show__()
     previous_button.show()
     next_button.show()
@@ -84,14 +83,14 @@ def start():
     #Launch event listener
     is_running = True
     while is_running == True:
-        main.set_caption(str(__image_id__) + ".jpg")
+        main.set_caption(str(ant_iv.get_image_id()) + ".jpg")
         main.get_surface().fill(white)
         previous_button = Buttons(main.get_surface(), "rectangle", "<", (200, 200, 200), {
                                   "width": 50, "height": 50}, [prev_bttn_pos[0], prev_bttn_pos[1], 50])
         next_button = Buttons(main.get_surface(), "rectangle", ">", (200, 200, 200), {
                               "width": 50, "height": 50}, [next_bttn_pos[0], next_bttn_pos[1], 50])
         id_text = body_font.render(
-            "Image ID: " + str(__image_id__), True, (0, 0, 0))
+            "Image ID: " + str(ant_iv.get_image_id()), True, (0, 0, 0))
[…]

### 2022-06-01 paper/main.tex
re-add paper with thesis writing, use new journal template

\documentclass{aci}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{lipsum}
\usepackage{hyperref}
\hypersetup{colorlinks=true}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{tikz}
\usepackage{makecell}
\usetikzlibrary{automata,positioning,arrows}
\usepackage{caption}
\usepackage{changepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\ep}{\varepsilon}
\newcommand{\eps}[1]{{#1}_{\varepsilon}}

\def\typeofarticle{Research Article} 
\def\currentvolume{3} 
\def\currentissue{1}
\def\currentyear{2022} 
\def\currentmonth{} 
\def\ppages{xx--xx} 
\def\DOI{xxxx} 
\def\Received{xx xx 2022} 
\def\Revised{xx xx 2022}
\def\Accepted{xx xx 2022} 
\def\Published{xx 2022}

\numberwithin{equation}{section}
\DeclareMathOperator*{\essinf}{ess\,inf}

\usepackage{array}
\newcolumntype{L}{>{\arraybackslash}m{8cm}}

\hyphenpenalty=10000
\usepackage{cite}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Ant Cuticle Texture Classification Algorithm for Ecological Anaylsis}

\author{%
    Noah Gardner\affil{1},
    John Paul Hellenbrand\affil{2},
    Anthony Phan\affil{1},
    Evan Zhu\affil{1},
    Zhiling Long\affil{3},
    and Chih-Cheng Hung\affil{1}\corrauth
}%

% \shortauthors is used in copyright information in the end of the paper
\shortauthors{the Author(s)}

\address{%
    \addr{\affilnum{1}}{Laboratory of Machine Vision and Security Research, %
        College of Computing and Software Engineering, Kennesaw State %
        University, Marietta GA, USA}%
    \addr{\affilnum{2}}{College of Science and Mathematics, Kennesaw State
        University, Kennesaw, GA, USA}%
    \addr{\affilnum{3}}{Department of Mathematics, Science, and Informatics,
        Mercer University, Macon, GA, USA}%
}
\corraddr{chung1@kennesaw.edu}

\editor{First-name Last-name}

\begin{abstract}
    Definition modeling, the task of generating a definition for a given term,
    is a relatively new area of research applied in evaluating word embeddings.
    Automatic generation of dictionary quality definitions has many applications
    in natural language processing, such as sentiment analysis, machine
    translation, and word sense disambiguation. Additionally, definition
    modeling is also helpful for evaluating the quality of word embeddings. As
    more research is done in this field, the need for a summary of different
    applications, approaches, and obstacles grows apparent. This review provides
    an overview of the current research in definition modeling and a list of
    future directions and trends.
\end{abstract}
\keywords{texture analysis; image processing; clustering; machine learning; ecology}
\maketitle

\section{Introduction}
Insects compose half of biodiversity and rank among the most dominant organisms
in terrestrial ecosystems \cite{sheikh_diverse_2017}. A key factor for the
ecological success of insects is their exoskeleton, also known as cuticle. The
cuticle protects insects from predation, provides structural support, prevents
desiccation, and serves as a canvas for advertising visual and chemical signals
\cite{gullan_insects_2009}. Research has heavily focused on the macrostructures
and internal chemical components that make the exoskeleton functional and more
recent work is being done to understand the functional aspects of external
cuticle micro sculpturing \cite{muthukrishnan_insect_2020,
    gunderson_insect_1989, watson_diversity_2017}.

Due to the extensive number of insect species, manual exploration of
insect-based information is difficult and often requires specialized expertise.
Therefore, automated entomology is gaining attraction by both biologists and
computer scientists and is expected to be a major contribution to the future of
insect-based research \cite{martineau_survey_2017}. One of the most commonly
used data types for insect analysis is image data. To develop an image-based
system for insect analysis, we can take advantage of existing work in general
image processing and texture analysis methods.

We examine ants (\textit{Formicidae}) as they display an extreme diversity of
cuticle micro sculpturing across all subfamilies. Sculpturing ranges from
parallel longitudinal ridges to deep oval impressions to erratic protuberances.
The sculpturing has arisen convergently and independently throughout ant's
evolutionary history, which suggests some inherent function. Cuticle sculpturing
on ants may help increase strength and rigidness, resist abrasion, increase
internal and external surface area, resist microbial growth, and rear beneficial
anti-biotic producing bacteria \cite{johnson_effect_2011,
    bruckner_relationship_2017, currie_coevolved_2006}. These specific functions may
be associated with certain sculpturing types and the purpose of classification
is to group similar textures based on proposed function.

In many texture analysis methods, the general goal is to automatically
categorize an object into a set of objects with similar texture-based features.
The approach of texture analysis to categorize similarly texture-based objects
corresponds well to the demands of ant identification. Texture analysis has
shown promising results in related fields, such as plant identification
\cite{boudra_plant_2018}. With modern texture analysis methods, the
categorization of ants can be automated and the results can be used to study the
influence of cuticle texture on ant ecology.

\section{Related Work}
\subsection{Sculpturing Identification}
Taxonomists have developed extensive terminology describing ant cuticle
sculpturing as it is often a useful diagnostic trait to distinguish between
closely related ant species \cite{blaimer_taxonomy_2019,fisher_ants_2007}. The
definitive text on ant cuticle terminology — \textit{The Glossary of Surface
    Sculpturing} — contains over 100 terms to describe the cuticle sculpturing
patterns of ants \cite{harris_glossary_1979}. There is often substantial overlap
among terms, and closely related cuticle patterns are likely to be functionally
similar. For example, the definition for “imbricate” is, “partly overlapping and
appearing like shingles on a roof or scales on a fish,” which is difficult to
distinguish from the definition of tesselate, “made up of squares like a chess
board, either in sculpturing or in color.”

Cuticle sculpturing in ants has been explored thoroughly from a taxonomic
perspective; however, the function of nano and microstructures on insect
exoskeletons is a developing topic in entomology. Watson et al. reviews the
literature of cuticle nano and microstructure function and proposes 21 possible
functions associated with these structures \cite{watson_diversity_2017}. Many of
these functions are related to structures not found in ants such as scales and
nanostructures. The review does include functions that may relate to ants such
as friction control, enhanced surface area, and increased hardness.  Watson et
al. also describe seven types of cuticle structures ranging from hairs and
scales to nano and micro structures \cite{watson_diversity_2017}. The cuticle
sculpturing of ants seems to fall within one type - complex microstructures.

The five broad functional groupings developed that describes the complex
microstructures found on ants were derived from reviewing the variation of
cuticle sculpturing across ants. These functional groupings were also reflected
in Harris as many terms could be grouped together based on similar definitions
and comparing the SEM photographs provided in the publication
\cite{harris_glossary_1979}.

\subsection{Insect Classification}

In this section, we provide an overview of some insect classification methods.
Proposed insect classification methods seek to classify insects at different
hierarchical levels, such as species, genus, family, and order. Additionally,
some methods may classify insects at a combination of different hierarchical
levels. Insect classification methods can be applied to a variety of fields. In
agriculture, insect classification methods can be used to identify the presence
of pest insects in crops, which can inform crop managers in their choice of
pesticides and help prevent crop loss \cite{liu_pestnet_2019,
    kasinathan_machine_2021}.

Feng et al. \cite{feng_automated_2013} apply an automated system to classify
moth images based on semantic related visual attributes, which are defined as a
pattern on the moth wings. Feng et al. \cite{feng_automated_2013} use a custom
texture descriptor based on the combination of Grey Level Co-occurence Matrtix
and \textit{Scale-Invariant Feature Transform} (SIFT) features
\cite{gotlieb_texture_1990, lowe_distinctive_2004}. The method proposed by Feng
et al. \cite{feng_automated_2013} is used to classify 50 different moth species
across 8 families and is based on standard texture analysis features
\cite{feng_automated_2013}. The results from Feng et al.
\cite{feng_automated_2013} suggest that traditional feature extraction
techniques for the semantic visual attributes of the moth wings are sufficient
for training a classifier to classify an image between 10 randomly selected moth
species.

Urteaga et al. \cite{urteaga_scorpions_2016} use machine learning methods in
order to classify images between two different scorpion species:
\textit{Centruroides limpidus} and \textit{Centruroides noxius}. After applying
background distinction based on dynamic color threshold, Urteaga et al.
\cite{urteaga_scorpions_2016} apply feature extraction to extract features from
the separated scorpion image such as aspect ratio, rectangularity, and
compactness. Urteaga et al. \cite{urteaga_scorpions_2016} apply three different
classification models to classify the image as one of the species: Artificial
Neural Network, Regression Tree, and Random Forest classifiers
\cite{urteaga_scorpions_2016}. The results from Urteaga et al.
\cite{urteaga_scorpions_2016} show that after background removal,
characteristics from the entire body of the scorpion can be used to create a
binary classifier that can classify the image as one of the two species.
Therefore, we may consider an experiment on the presence of the background from
the ant head image in order to improve classification accuracy. However, the
features extracted from \cite{urteaga_scorpions_2016} do not make use of any
deep learning methods and are not compatible with our texture-based dataset.

Lim et al. \cite{lim_performance_2017} apply a CNN-based algorithm for insect
classification. Lim et al. \cite{lim_performance_2017} classify a subset of
insect species and families based on the classes available in the ImageNet
dataset. ImageNet is a widely used dataset of images labeled by experts with
millions of images and thousands of categories \cite{deng_imagenet_2009}. In the
ImageNet dataset, there are some categories that specify the class of the insect
on a species level, \textit{e.g.} \textit{monarch butterfly} and \textit{ringlet
    butterfly} as well as some categories that specify the class of the insect on a
family level, \textit{e.g.} \textit{ant}, \textit{fly}, and \textit{bee}
\cite{imagenet_labels}. Lim et al. \cite{lim_performance_2017} use a modified
AlexNet architecture and experiment with different numbers of kernels and their
effect the performance of the model. Glick et al. \cite{glick_insect_2016}
employ a similar approach by classifying 277 insect classes from ImageNet using
a hierarchical contains. The results from Lim et al. \cite{lim_performance_2017}
and Glick et al. \cite{glick_insect_2016} suggest that a CNN is capable of
differentiating between different hierarchical classes of insects. In our
research, we are interested in the classification of ant images of a
hierarchical level between species and family.

\section{Methodology}

\subsection{Dataset}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.7\textwidth]{thesis_assets/images/rough_collage.png}
    \caption{Examples of rough cuticle texture ant images in the dataset after
        center cropping.}
    \label{fig:rough-cuticle-texture}
\end{figure}

In this section, we describe the creation of the custom dataset used in this
research. In our dataset, we classify ant head images from AntWeb
\cite{perrichot_antweb_2012} into two categories: \textit{rough} and
\textit{smooth}. Some randomly selected images from each category are shown in
Figures \ref{fig:rough-cuticle-texture} and \ref{fig:smooth-cuticle-texture}.

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{thesis_assets/images/smooth_collage.png}
    \caption{Examples of smooth cuticle texture ant images in the dataset after
        center cropping.}
    \label{fig:smooth-cuticle-texture}
\end{figure}

\subsubsection{Sculpture Identification Protocol}

To begin, a master spreadsheet was created with the 2,499 different ant species
to be identified for the primary dataset. A team of three assistants were in
charge of the manual identification process. The team was trained to identify
cuticle sculpturing through a process which consisted of one 45-minute
introductory lesson explaining the project and texture categories and then given
a training set of photos to identify from the genus \textit{Polyrhachis}, whose
members display among the highest diversity of cuticle patterns all categories.
The sculpture identification protocol describes the two primary categories:
\textit{rough} and \textit{smooth}.

Initially, the sculpture identification protocol had 8 subcategories of cuticle
texture, including dimpled, ridged, and differing levels of smooth texture. For
simplicity, we work only with the two main categories. The training set
identifications were reviewed together as a group by the assistants. Once
training was complete, assistants were assigned the same genera of ants to
identify independently each week. A weekly meeting was held to discuss
identifications and assign new ones. These identifications were collected in the
master spreadsheet and the identifications were assigned to individual ant
species on a majority basis.

\subsubsection{Data Collection}

To collect the images, the assistants followed the taxonomy information
available in the master spreadsheet to the appropriate AntWeb page. In many
cases, there are multiple ant head images of the same species, and occasionally
there are multiple image resolution available from a single image. To simplify
the data collection process, the assistants were instructed to download the
first ant head image of the species being identified in the highest resolution
possible. Each image was named with an identifier that corresponds with the row
number in the master spreadsheet. The same ant head images that were downloaded
in the data collection phase were the same ones used in the sculpture
identification protocol. Ant species which did not have any images of the head
were excluded from the dataset. Additionally, ant species which only had a head
image of a queen ant were excluded from the dataset.

Ant specimen images taken from AntWeb \cite{perrichot_antweb_2012} are created
by different photographers and therefore have different attributes, such as
environment, resolution, and lighting. In the ant head images, the ant head is
in the center of the image and the body is pointing away from the camera. The
focus of the ant head image is centered on the head, with the background and
image artifacts from the ant body typically blurred. In most ant head images,
there is a bar which indicates the scale of the image due to the variety in the
sizes of different ant species. In a few ant head images, there exists some text
denoting the specimen identifier and other info. In terms of texture, some ant
specimens are very old, so their head images have other abnormalities such as
cracks in the cuticle and the presence of dust.

\subsubsection{Data Preprocessing}

Due to the variety of the ant head image attributes, we apply simple
preprocessing before the images are used in our model. We want the images to
have a uniform size for simplicity in our classification process. Since the ant
head images are typically centered in the image, we apply a center crop to each
image to create a square image of the same size. Once the image is square, we
resize each image to a fixed size of 256x256 pixels. We leave other
discrepancies in the images untouched.

Our custom dataset of ant head images contains 2,499 images. 1072 samples of
rough textured ant cuticle textures comprise 43\% of the dataset. The remaining
1427 samples of smooth textured ant cuticle textures comprise 57\% of the
dataset. To handle the imbalance of the dataset, we apply undersampling for each
class for the training dataset. By using random stratified sampling, we
construct a training set with 800 images per class. The remaining images are
randomly split between test and validation, which turns out to roughly a
60\%/20\%/20\% train, test, and validation data split. With 272 rough samples
and 627 smooth samples left over after the stratified split, the test dataset
has roughly 136 rough samples and 313 smooth samples. Since these leftover
samples are split with code by 50\% there will be some rounding variance and
therefore the test dataset built at run-time will not always have exactly the
same number of samples.

\subsection{Models}

Our first model is \textit{visual geometry group} (VGG), a convolutional neural
network that takes advantage of very small convolutional filters in a deep
network architecture \cite{simonyan_very_2015}. We compare four architectures of
VGG: VGG11, VGG13, VGG16, and VGG19. The primary difference between the
architectures is the number of layers in each model. Our second model is
\textit{residual network} (ResNet), a deep network architecture that includes
shortcut connections between layers (residual connections) \cite{he_deep_2015}.
We compare three architectures of ResNet: ResNet18, ResNet50, and ResNet101.
Again, the primary difference between the architectures is the number of layers
in each model.

For our ResNet models, we have two versions: randomized and pretrained. The
randomized version is the same architecture, but the weights are randomly
initialized. The pretrained version has weights from training on the CIFAR
dataset, an image dataset with 1000 classes. In this case, we are fine-tuning
the pretrained model. For VGG, we are only using the randomized version. The
base VGG architecture also has an output layer of size 1000. Since we are
working with a binary classification problem, we modify the architecture for all
models to have an output layer of size 2. Each model is trained over 100 epochs,
using stochastic gradient descent with momentum. The batch size is set to 16
images. We apply a learning rate of 0.001 and momentum parameter of 0.9.

\subsubsection{Evaluation}

We evaluate the performance of the models according to standard evaluation
methods. Since we are working with a binary classification problem, we use a
standard confusion matrix to evaluate the accuracy, precision, and F1 score. We
also apply Grad-CAM with manual inspection to visualize the activation weights
for incorrectly classified images to determine which features are interfering
with the classification. Finally, we apply t-SNE to visualize the separation
learned for the model to further analyze the classifications made by the model.

\section{Results}
\subsection{Environment}

Experiments are run on an Ubuntu 18.04 LTS Lambda Labs GPU server. The server
contains 8 NVIDIA GeForce RTX 2080 Ti graphics cards with 12GB of memory each.
The server uses an Intel Xeon Silver 4116 with 48 total threads and maximum
frequency of 3.000 GHz, and has 256GB of RAM.

% vgg results
\subsection{VGG Models}

To begin, we share the results on the VGG model architectures on our custom
dataset. The classification results are summarized in a confusion matrix for
each model. Then, the statistics for each VGG architecture are shared in Table
\ref{tab:vgg_results} and Figure \ref{fig:vgg_results}. The results are
collected on 8 iterations of training and averaged. The results are
rounded to 2 decimal places where appropriate.

\vspace{0.5in}

\begin{table}[h]
    \centering
    \caption{VGG11 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 101.75     & 53          \\
        Predicted Smooth & 37.375     & 257.875     \\
    \end{tabular}
    \label{tab:vgg11_confusion_matrix}
\end{table}

\vspace{0.5in}

\begin{table}[h]
    \centering
    \caption{VGG13 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 95.25      & 39.375      \\
        Predicted Smooth & 43.25      & 272.25      \\
    \end{tabular}
    \label{tab:vgg13_confusion_matrix}
\end{table}

\vspace{0.5in}

\begin{table}[h]
    \centering
    \caption{VGG16 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 87.5       & 42.625      \\
        Predicted Smooth & 47.375     & 272.5       \\
    \end{tabular}
    \label{tab:vgg16_confusion_matrix}
\end{table}

\vspace{0.5in}

\begin{table}[h]
    \centering
    \caption{VGG19 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 91.25      & 42.125      \\
        Predicted Smooth & 43.25      & 273.375     \\
    \end{tabular}
    \label{tab:vgg19_confusion_matrix}
\end{table}

\vspace{0.5in}

\begin{table}[h]
    \centering
    \caption{Average results for VGG architectures on ant head image dataset.}
    \begin{tabular}{r|llll}
              & Recall & Precision & F1 Score & Accuracy \\
        \hline
        VGG11 & 0.83   & 0.88      & 0.85     & 0.80     \\
        VGG13 & 0.87   & 0.87      & 0.87     & 0.82     \\
        VGG16 & 0.86   & 0.86      & 0.86     & 0.80     \\
        VGG19 & 0.87   & 0.87      & 0.87     & 0.81     \\
    \end{tabular}
    \label{tab:vgg_results}
\end{table}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{thesis_assets/plots/vgg_average.png}
    \caption{Average accuracy for VGG architectures on ant head image dataset.}
    \label{fig:vgg_results}
\end{figure}

\subsection{ResNet Models}

Next, we share the results on the ResNet model architectures on our custom
dataset with random weight initialization. The classification results are
summarized in a confusion matrix for each model. Then, the statistics for each
ResNet architecture are shared in Table \ref{tab:resnet_results} and Figure
\ref{fig:resnet_results}. The results are collected on 8 iterations of training and
averaged. The results are rounded to 2 decimal places where appropriate.

\begin{table}[h]
    \centering
    \caption{ResNet18 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 98.75      & 48.75       \\
        Predicted Smooth & 39.25      & 262         \\
    \end{tabular}
    \label{tab:resnet18_confusion_matrix}
\end{table}

\vspace{0.5in}

\begin{table}[h]
    \centering
    \caption{ResNet50 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 81.375     & 43.625      \\
        Predicted Smooth & 55.875     & 269.125     \\
    \end{tabular}
    \label{tab:resnet50_confusion_matrix}
\end{table}

\begin{table}[h]
    \centering
    \caption{ResNet101 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 85.75      & 53.875      \\
        Predicted Smooth & 50.125     & 260.25      \\
    \end{tabular}
    \label{tab:resnet101_confusion_matrix}
\end{table}

\begin{table}[h]
    \centering
    \caption{Average results for ResNet architectures on ant head image dataset.}
    \begin{tabular}{r|llll}
                  & Recall & Precision & F1 Score & Accuracy \\
        \hline
        ResNet18  & 0.84   & 0.87      & 0.85     & 0.80     \\
        ResNet50  & 0.86   & 0.83      & 0.84     & 0.77     \\
        ResNet101 & 0.83   & 0.84      & 0.83     & 0.76     \\
    \end{tabular}
    \label{tab:resnet_results}
\end{table}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{thesis_assets/plots/resnet_average.png}
    \caption{Average accuracy for ResNet architectures on ant head image dataset.}
    \label{fig:resnet_results}
\end{figure}

\clearpage
\subsection{Fine-tuned ResNet Models}

Next, we share the results on the fine-tuned ResNet model architectures on our
custom dataset with pretrained weights. The classification results are
summarized in a confusion matrix for each model. Then, the statistics for each
ResNet architecture are shared in Table \ref{tab:fresnet_results} and Figure
\ref{fig:fresnet_results}. The results are collected on 8 iterations of training
and averaged. The results are rounded to 2 decimal places where appropriate.

\begin{table}[h]
    \centering
    \caption{Fine-tuned ResNet18 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 117.375    & 35.125      \\
        Predicted Smooth & 17.875     & 279.625     \\
    \end{tabular}
    \label{tab:fresnet18_confusion_matrix}
\end{table}

\vspace{0.5in}

\begin{table}[h]
    \centering
    \caption{Fine-tuned ResNet50 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 114.5      & 33.25       \\
        Predicted Smooth & 24.125     & 278.125     \\
    \end{tabular}
    \label{tab:fresnet50_confusion_matrix}
\end{table}

\vspace{0.5in}

\begin{table}[h]
    \centering
    \caption{Fine-tuned ResNet101 confusion matrix on ant head image dataset.}
    \begin{tabular}{r|ll}
                         & True Rough & True Smooth \\
        \hline
        Predicted Rough  & 117.25     & 33.25       \\
        Predicted Smooth & 18.375     & 281.125     \\
    \end{tabular}
    \label{tab:fresnet101_confusion_matrix}
\end{table}

\vspace{0.5in}

\begin{table}[h]
    \centering
    \caption{Average results for Fine-tuned ResNet architectures on ant head image dataset.}
    \begin{tabular}{r|llll}
                  & Recall & Precision & F1 Score & Accuracy \\
        \hline
        ResNet18  & 0.89   & 0.94      & 0.91     & 0.88     \\
        ResNet50  & 0.89   & 0.92      & 0.91     & 0.87     \\
        ResNet101 & 0.89   & 0.94      & 0.92     & 0.88     \\
    \end{tabular}
    \label{tab:fresnet_results}
\end{table}

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{thesis_assets/plots/resnet_finetuned_average.png}
    \caption{Average accuracy for fine-tuned ResNet architectures on ant head image dataset.}
    \label{fig:fresnet_results}
\end{figure}

\clearpage

\section{Analysis}

The results in the previous sections show that the fine-tuned ResNet models
outperform the VGG and randomly initialized ResNet models on the task of ant
head image classification. It should be noted that due to the class imbalance in
the dataset, the F1 score is the preferable metric to the accuracy. On average,
the fine-tuned ResNet101 model performed the best with an average F1 score of
0.92. We further analyze the separation learned by both ResNet101 models in the
following section.

\subsection{t-SNE Visualization}

\begin{figure}[h]
    \centering
    \begin{subfigure}{.45\textwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/plots/resnet101_gt_tsne.png}
        % \caption{Ground truth labels}
    \end{subfigure}
    \begin{subfigure}{.45\textwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/plots/resnet101_pred_tsne.png}
        % \caption{Predicted labels}
    \end{subfigure}
    \caption{t-SNE visualization of the embeddings of the second to last layer
        of the randomly initialized ResNet101 model trained on ant head image
        dataset.}
    \label{fig:resnet101_tsne}
\end{figure}

In this section, we provide visualization of the fine-tuned ResNet101 model and
the randomly initialized ResNet101 model using t-SNE dimensionality reduction.
First, we run the dataset preprocessing method and initialize both models. Then,
both models are trained according to the training parameters and the state of
each model is saved. To visualize the deep extracted features, we modify each
model to obtain the embeddings of the second to last layer. Then, we use the
t-SNE algorithm to reduce the dimensionality of the embeddings to 2 dimensions.
We plot side-by-side the ground truth and predicted labels for each model.
Figure \ref{fig:resnet101_tsne} shows the results of the trained randomly
initialized model and Figure \ref{fig:fresnet101_tsne} shows the results of the
fine-tuned model.

\begin{figure}[h]
    \centering
    \begin{subfigure}{.45\textwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/plots/fresnet101_gt_tsne.png}
        % \caption{Ground truth labels}
    \end{subfigure}
    \begin{subfigure}{.45\textwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/plots/fresnet101_pred_tsne.png}
        % \caption{Predicted labels}
    \end{subfigure}
    \caption{t-SNE visualization of the embeddings of the second to last layer
        of the fine-tuned ResNet101 model trained on ant head image dataset.}
    \label{fig:fresnet101_tsne}
\end{figure}

Based on the visual results of the t-SNE visualization, we can see that the
fine-tuned model learned a stronger separation of the two classes, which
reinforces the results that the fine-tuned model received a higher average
accuracy.

\subsection{GradCAM Visualization}

In this section, we provide some visual analysis of some correctly and
incorrectly classified images using GradCAM. We provide two categories and two
subcategories in our analysis. The two categories are correct and incorrect
classification. Regardles of the feature activation map, the correctly
classified images have the same predicted label as the ground truth, and
incorrectly classified images have different predicted labels. The two
subcategories are ideal and non-ideal feature activation. In the ideal case, the
features that are used to compute the classification are the same as the
features used by the assistants in the sculpture identification process. In
general, the features used by the assistants are the textures of the cuticle on
the ant head. In the non-ideal case, the features used to compute the
classification are not from the head, for example, from the background,
extraneous text, or the body of the ant. We used randomly selected images from
the dataset and the fine-tuned ResNet101 model to perform the analysis. We show
the GradCAM results in Figures \ref{fig:correct_ideal},
\ref{fig:correct_nonideal}, \ref{fig:incorrect_ideal}, and
\ref{fig:incorrect_nonideal}. The left image shows the preprocessed image input
to the model. The right image shows the GradCAM output based on the
classification. Four specimens were selected randomly from each category and
subcategory.

Correctly classified images which use ideal features show the ideal performance
of the model. Incorrectly classified images which use ideal features should be
further analyzed. In essence, the model in this situation knows \textit{where}
to look, but not \textit{what} to look for. In Figure
\ref{fig:incorrect_ideal_84}, the features activated are mostly in the correct
location on the ant head, and the rough texture is clearly visible, yet the
model predicts the incorrect class \textit{smooth}. Similarly in Figure
\ref{fig:incorrect_ideal_177}, the features activated are also mostly in the
correct location, yet the model predicts the incorrect class \textit{smooth}. In
this case, it may be due to the pose of the ant being slightly different from
the average pose. In the incorrectly classified images with non-ideal features,
analysis shows that the model is unable to find \textit{where} to look, and
obtains feature information from other parts of the ant or the background. Cases
where the image was correctly classified using the non-ideal features can
basically be seen as noise. In order to further analyze this class, we should
introduce some parameter such as model confidence to examine further.

\newcommand{\subwidth}{0.35\textwidth}
\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
        \caption{Ground truth: smooth}
        \label{fig:correct_ideal_518}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
        \caption{Ground truth: smooth}
        \label{fig:correct_ideal_808}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
        \caption{Ground truth: rough}
        \label{fig:correct_ideal_842}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
        \caption{Ground truth: rough}
        \label{fig:correct_ideal_1091}
    \end{subfigure}
    \caption{Correctly classified images using ideal features.}
    \label{fig:correct_ideal}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
        \caption{Ground truth: rough}
        \label{fig:correct_nonideal_346}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
        \caption{Ground truth: rough}
        \label{fig:correct_nonideal_1554}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
        \caption{Ground truth: rough}
        \label{fig:correct_nonideal_1694}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
        \caption{Ground truth: smooth}
        \label{fig:correct_nonideal_388}
    \end{subfigure}
    \caption{Correctly classified images using non-ideal features.}
    \label{fig:correct_nonideal}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/84.png}
        \caption{Ground truth: rough}
        \label{fig:incorrect_ideal_84}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/86.png}
        \caption{Ground truth: smooth}
        \label{fig:incorrect_ideal_86}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/177.png}
        \caption{Ground truth: smooth}
        \label{fig:incorrect_ideal_177}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/197.png}
        \caption{Ground truth: smooth}
        \label{fig:incorrect_ideal_197}
    \end{subfigure}
    \caption{Incorrectly classified images using ideal features.}
    \label{fig:incorrect_ideal}
\end{figure}

\begin{figure}
    \centering
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/1.png}
        \caption{Ground truth: rough}
        \label{fig:incorrect_nonideal_1}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/22.png}
        \caption{Ground truth: rough}
        \label{fig:incorrect_nonideal_22}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/61.png}
        \caption{Ground truth: rough}
        \label{fig:incorrect_nonideal_61}
    \end{subfigure}
    \begin{subfigure}{\subwidth}
        \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/204.png}
        \caption{Ground truth: rough}
        \label{fig:incorrect_nonideal_204}
    \end{subfigure}
    \caption{Incorrectly classified images using non-ideal features.}
    \label{fig:incorrect_nonideal}
\end{figure}

\clearpage
\section{Conclusion}
Ant cuticle texture presumably has some function, but without the proper tools,
evaluating the function based on thousands of species is infeasible. We have
shown in this work that a deep learning approach can be used to automatically
categorize ants based on their cuticle texture, therefore supporting research on
the evaluation of the function in future work. Our categorization system is
novel in the field of automated insect indentification due to the broad number
of species captured by it. Additionally, a model that is pre-trained on a
diverse image task such as ResNet can be transferred to our domain of texture
analysis. All code is publicly available on GitHub
(https://github.com/ngngardner/cuticulus).

\bibliographystyle{AIMS}
\bibliography{main.bib}
\end{document}

### 2022-06-01 paper/main.tex
update abstract

@@
 \begin{abstract}
-    Definition modeling, the task of generating a definition for a given term,
-    is a relatively new area of research applied in evaluating word embeddings.
-    Automatic generation of dictionary quality definitions has many applications
-    in natural language processing, such as sentiment analysis, machine
-    translation, and word sense disambiguation. Additionally, definition
-    modeling is also helpful for evaluating the quality of word embeddings. As
-    more research is done in this field, the need for a summary of different
-    applications, approaches, and obstacles grows apparent. This review provides
-    an overview of the current research in definition modeling and a list of
-    future directions and trends.
+    There is a large variety of ant species, and most species are diverse in
+    terms of size, shape, behaviors, and especially cuticle textures. However,
+    the significance of ant cuticle texture is not widely researched. Ant
+    cuticle texture presumably provides some type of function, and therefore is
+    useful to research for ecological applications and bioinspired designs. This
+    research employs modern machine learning methods such as texture analysis
+    and deep machine learning to automatically group similar ant species based
+    on morphological traits. We provide a comparative study of the performance
+    of modern texture analysis methods on ant head images. We evaluate the
+    results of the classification methods with modern visualization techniques.
 \end{abstract}

### 2022-06-08 cuticulus/__main__.py
gabor filter feature extraction

@@ -1,17 +1,24 @@
 """Main entrypoint."""
 
-from cuticulus.datasets import BackgroundSub
-
-from rich import inspect
+from cuticulus.core.collages import build_collage
+from cuticulus.datasets import RoughSmoothSub
 
 
 def main():
     """Run main function."""
-    patch_size = (16, 16)
-    ds = BackgroundSub(
+    patch_size = (32, 32)
+    ds = RoughSmoothSub(
         size=patch_size,
+        rebuild=True,
     )
-    ds.stratified_split(n_samples=5000, clamp=True)
+
+    build_collage(
+        ds.labels,
+        ds.images,
+        name='bg',
+        rows=10,
+        cols=10,
+    )
 
 
 if __name__ == '__main__':

### 2022-06-08 cuticulus/core/collages.py

@@ -1,0 +1,57 @@
+"""Build collage of images from dataset for visualization."""
+
+from pathlib import Path
+
+import numpy as np
+from PIL import Image
+
+OUT_PATH = Path('./output/images')
+OUT_PATH.mkdir(exist_ok=True, parents=True)
+
+
+def build_collage(
+    labels,
+    images,
+    name,
+    out_path=OUT_PATH,
+    rows=3,
+    cols=5,
+):
+    """Build collage of images from dataset for visualization.
+
+    Args:
+        labels (np.ndarray): Labels of images.
+        images (np.ndarray): Images.
+        name (str): Name of collage.
+        out_path (Path): Path to save collage.
+        rows (int): Number of rows in collage.
+        cols (int): Number of columns in collage.
+
+    Returns:
+        None
+    """
+    label_map = {
+        0: 'rough',
+        1: 'smooth',
+    }
+    size = images[0].shape[:2]
+    for label in [0, 1]:
+        idx = np.random.choice(
+            np.where(labels == label)[0],
+            size=rows * cols,
+            replace=False,
+        )
+
+        # create a collage of images
+        collage = Image.new('RGB', (size[1] * cols, size[0] * rows))
+        for i, idx in enumerate(idx):
+            image = Image.fromarray(images[idx])
+            collage.paste(image, (size[1] * (i % cols), size[0] * (i // cols)))
+
+        # scale down the collage
+        collage = collage.resize((size[1] * cols // 2, size[0] * rows // 2))
+        collage.save('{0}/{1}_{2}_collage.png'.format(
+            out_path,
+            label_map[label],
+            name,
+        ))

### 2022-06-08 cuticulus/core/datasets/imutils.py

@@ -27,11 +27,22 @@ def autocrop(image: np.ndarray) -> np.ndarray:
     ]
 
 
+@beartype
 def shape_to_mask(
-    img_shape,
-    points,
-    shape_type=None,
-):
+    img_shape: tuple,
+    points: list,
+    shape_type: str = None,
+) -> np.ndarray:
+    """Create a mask from a set of points.
+
+    Args:
+        img_shape (tuple): The shape of the image.
+        points (list): The points to create the mask from.
+        shape_type (str): The type of shape to create (polygon).
+
+    Returns:
+        np.ndarray: The mask.
+    """
     mask = np.zeros(img_shape[:2], dtype=np.uint8)
     mask = Image.fromarray(mask)
     draw = ImageDraw.Draw(mask)
@@ -47,7 +58,22 @@ def shape_to_mask(
     return np.array(mask, dtype=bool)
 
 
-def shapes_to_label(img_shape, shapes, label_map):
+@beartype
+def shapes_to_label(
+    img_shape: tuple,
+    shapes: list,
+    label_map: dict,
+) -> tuple:
+    """Create a label mask array from a list of shapes.
+
+    Args:
+        img_shape (tuple): The shape of the image.
+        shapes (list): The shapes to create the label mask from.
+        label_map (dict): The label map to use.
+
+    Returns:
+        tuple: The label mask array and the label map.
+    """
     arr = np.zeros(img_shape[:2], dtype=np.int32)
     ins = np.zeros_like(arr)
     instances = []

### 2022-06-08 cuticulus/core/datasets/sub.py

@@ -110,7 +110,8 @@ class SubDataset(DatasetSplitter):
         img = Image.fromarray(arr)
 
         if self.source_size != const.NO_RESIZE:
-            img = img.resize(self.source_size, Image.ANTIALIAS)
+            raise NotImplementedError
+            # img = img.resize(self.source_size, Image.ANTIALIAS)
 
         return super().preprocess(np.array(img))
 
@@ -163,9 +164,11 @@ class SubDataset(DatasetSplitter):
         labels = []
         ids = []
 
-        rs_files = glob(str(self.base_path / 'rs' / '*.json'))
-        bg_files = glob(str(self.base_path / 'bg' / '*.json'))
-        files = rs_files + bg_files
+        files = None
+        if self.ds_type == 'rs':
+            files = glob(str(self.base_path / 'rs' / '*.json'))
+        elif self.ds_type == 'bg':
+            files = glob(str(self.base_path / 'bg' / '*.json'))
 
         for pfile in files:
             # load image data based on avaiable jsons
@@ -209,13 +212,14 @@ class SubDataset(DatasetSplitter):
 
                     # rough-smooth dataset only considers cuticle segments
                     if self.ds_type == 'rs':
-                        if label_names[sub_label] == 'cuticle':
-                            images.append(sub_image)
-                            labels.append(label)
-                            ids.append(iid)
+                        if label_names[sub_label] == '_background_':
+                            continue
+                        images.append(sub_image)
+                        labels.append(label)
+                        ids.append(iid)
 
                     # background dataset
-                    else:
+                    elif self.ds_type == 'bg':
                         if label_names[sub_label] == 'cuticle':
                             # not background
                             labels.append(1)

### 2022-06-10 paper/main.tex
paper update

@@
 \usepackage{hyperref}
+\usepackage{xcolor}
 \hypersetup{colorlinks=true}
@@
 
-\title{Ant Cuticle Texture Classification Algorithm for Ecological Anaylsis}
+\title{Ant cuticle image classification using texture analysis: a comparative study}
 
@@
     Zhiling Long\affil{3},
+    Min Wang\affil{2},
+    Clint Penick\affil{2},
     and Chih-Cheng Hung\affil{1}\corrauth
@@
     \addr{\affilnum{3}}{Department of Mathematics, Science, and Informatics,
-        Mercer University, Macon, GA, USA}%
+        Mercer University, Atlanta, GA, USA}%
 }
@@
 
-\editor{First-name Last-name}
+\editor{Pasi Fr\"{a}nti}
 
@@
     useful to research for ecological applications and bioinspired designs. This
-    research employs modern machine learning methods such as texture analysis
-    and deep machine learning to automatically group similar ant species based
-    on morphological traits. We provide a comparative study of the performance
-    of modern texture analysis methods on ant head images. We evaluate the
-    results of the classification methods with modern visualization techniques.
+    research employs image texture analysis and deep machine learning to
+    automatically group similar ant species based on morphological traits. We
+    provide a comparative study of the performance of texture analysis methods
+    on ant images. We evaluate the results of the classification methods with
+    modern visualization techniques.
+    \colorbox{green}{TODO: what is the result? experiments? data?}
 \end{abstract}
-\keywords{texture analysis; image processing; clustering; machine learning; ecology}
+\keywords{texture analysis; image processing; classification; machine learning;
+    ant cuticle images; ecology}
 \maketitle
@@
 \section{Introduction}
+\colorbox{green}{TODO: refer to ant head images in a consistent way}
+
+\colorbox{green}{TODO: refer to traditional/classical in a consistent way}
+
+\colorbox{green}{TODO: change ideal and non-ideal to features identified by
+    the biologist and to invalid features (background, etc.)}
+
 Insects compose half of biodiversity and rank among the most dominant organisms
@@
 system for insect analysis, we can take advantage of existing work in general
-image processing and texture analysis methods.
+image analysis methods.
 
@@
     bruckner_relationship_2017, currie_coevolved_2006}. These specific functions may
-be associated with certain sculpturing types and the purpose of classification
-is to group similar textures based on proposed function.
+be associated with certain sculpturing types. In order to analyze those
+functions, it is necessary to segment the image and group similar textures for
+further analysis.
 
-In many texture analysis methods, the general goal is to automatically
+In image texture analysis methods, the general goal is to automatically
 categorize an object into a set of objects with similar texture-based features.
-The approach of texture analysis to categorize similarly texture-based objects
-corresponds well to the demands of ant identification. Texture analysis has
-shown promising results in related fields, such as plant identification
-\cite{boudra_plant_2018}. With modern texture analysis methods, the
-categorization of ants can be automated and the results can be used to study the
-influence of cuticle texture on ant ecology.
+Texture analysis has shown promising results in related fields, such as plant
+identification \cite{boudra_plant_2018} \colorbox{green}{TODO: more references
+    and examples}. With modern texture analysis methods, the categorization of ants
+can be automated and the results can be used to study the influence of cuticle
+texture on ant ecology.
+
+Image texture analysis heavily depends on the spatial relationships among gray
+levels of pixels \cite{hung_image_2019}. Hence, feature extraction method is
+usually applied to the spatial relationship. The gray-level co-occurrence matrix
+(GLCM) and local binary patterns (LBP) are two of methods which are based on the
+spatial relationships to extracting features for this purpose
+\cite{haralick_textural_1973, goos_gray_2000}. Traditional image texture
+analysis can be divided into four categories: statistical, structural,
+model-based, and transform-based methods \colorbox{green}{[references]}. In the
+past decade, the deep learning network becomes the mainstream in image
+classification and segmentation \cite{liu_bow_2019}. Compared with the
+traditional method in which a kernel must be designed for extracting features,
+the deep learning network can automatically extract the features by itself.
+
+In this study, we used some of both approaches for our work on the ant images.
+
+\colorbox{green}{TODO: Specifically, what approaches used? Why do we use image
+    texture for ant images? What are the contributions?}
+
+\colorbox{green}{TODO: briefly discuss results from the research}
 
 \section{Related Work}
+
+\colorbox{green}{TODO: introduce the section}
+
 \subsection{Sculpturing Identification}
@@
 perspective; however, the function of nano and microstructures on insect
-exoskeletons is a developing topic in entomology. Watson et al. reviews the
-literature of cuticle nano and microstructure function and proposes 21 possible
+exoskeletons is a developing topic in entomology. Watson et al. review the
+literature of cuticle nano and microstructure function and propose 21 possible
 functions associated with these structures \cite{watson_diversity_2017}. Many of
@@
 
-The five broad functional groupings developed that describes the complex
+The five broad functional groupings developed that describe the complex
 microstructures found on ants were derived from reviewing the variation of
@@
 in Harris as many terms could be grouped together based on similar definitions
-and comparing the SEM photographs provided in the publication
-\cite{harris_glossary_1979}.
+and comparing the scanning electron microscope photographs provided in the
+publication \cite{harris_glossary_1979}.
+
+\colorbox{green}{TODO: how do we connect this section to image texture analysis?
+    (why do we research this and why image texture) }
+
+\subsection{Texture Analysis}
+\colorbox{green}{TODO: A brief survey on image texture and cite Liu's
+    IJCV2019 survey paper}
 
 \subsection{Insect Classification}
-
 In this section, we provide an overview of some insect classification methods.
@@
 pattern on the moth wings. Feng et al. \cite{feng_automated_2013} use a custom
-texture descriptor based on the combination of Grey Level Co-occurence Matrtix
-and \textit{Scale-Invariant Feature Transform} (SIFT) features
-\cite{gotlieb_texture_1990, lowe_distinctive_2004}. The method proposed by Feng
-et al. \cite{feng_automated_2013} is used to classify 50 different moth species
-across 8 families and is based on standard texture analysis features
-\cite{feng_automated_2013}. The results from Feng et al.
+texture descriptor based on the combination of GLCM and \textit{scale-invariant
+    feature transform} (SIFT) features \cite{gotlieb_texture_1990,
+    lowe_distinctive_2004}. The method proposed by Feng et al.
+\cite{feng_automated_2013} is used to classify 50 different moth species across
+8 families \cite{feng_automated_2013}. The results from Feng et al.
 \cite{feng_automated_2013} suggest that traditional feature extraction
@@
 binary classifier that can classify the image as one of the two species.
-Therefore, we may consider an experiment on the presence of the background from
-the ant head image in order to improve classification accuracy. However, the
-features extracted from \cite{urteaga_scorpions_2016} do not make use of any
-deep learning methods and are not compatible with our texture-based dataset.
 
@@
 employ a similar approach by classifying 277 insect classes from ImageNet using
-a hierarchical contains. The results from Lim et al. \cite{lim_performance_2017}
-and Glick et al. \cite{glick_insect_2016} suggest that a CNN is capable of
-differentiating between different hierarchical classes of insects. In our
-research, we are interested in the classification of ant images of a
-hierarchical level between species and family.
+a hierarchical convolutional neural network. The results from Lim et al.
+\cite{lim_performance_2017} and Glick et al. \cite{glick_insect_2016} suggest
+that a CNN is capable of differentiating between different hierarchical classes
+of insects.
 
@@
 
-\subsection{Dataset}
+\colorbox{green}{TODO: do not jump into subsection directly without any context}
+
+\subsection{Dataset preparation} % preparation generation
 
@@
     \caption{Examples of rough cuticle texture ant images in the dataset after
-        center cropping.}
+        center cropping, from AntWeb \cite{perrichot_antweb_2012}.}
     \label{fig:rough-cuticle-texture}
@@
 In this section, we describe the creation of the custom dataset used in this
-research. In our dataset, we classify ant head images from AntWeb
-\cite{perrichot_antweb_2012} into two categories: \textit{rough} and
-\textit{smooth}. Some randomly selected images from each category are shown in
-Figures \ref{fig:rough-cuticle-texture} and \ref{fig:smooth-cuticle-texture}.
+research. In our dataset, we use ant head images from AntWeb
+\cite{perrichot_antweb_2012} and define two categories for them based on the
+appearance of the cuticle texture: \textit{rough} and \textit{smooth}. Some
+randomly selected images from each category are shown in Figures
+\ref{fig:rough-cuticle-texture} and \ref{fig:smooth-cuticle-texture}.
 
@@
     \caption{Examples of smooth cuticle texture ant images in the dataset after
-        center cropping.}
+        center cropping, from AntWeb \cite{perrichot_antweb_2012}.}
     \label{fig:smooth-cuticle-texture}
@@
 
-\subsubsection{Sculpture Identification Protocol}
+% \subsubsection{Sculpture identification protocol}
 
 To begin, a master spreadsheet was created with the 2,499 different ant species
-to be identified for the primary dataset. A team of three assistants were in
-charge of the manual identification process. The team was trained to identify
+to be identified for the primary dataset. The team was trained to identify
 cuticle sculpturing through a process which consisted of one 45-minute
-introductory lesson explaining the project and texture categories and then given
-a training set of photos to identify from the genus \textit{Polyrhachis}, whose
-members display among the highest diversity of cuticle patterns all categories.
-The sculpture identification protocol describes the two primary categories:
-\textit{rough} and \textit{smooth}.
+introductory lesson explaining the project and texture categories. Then, the
+team was given a training set of photos to identify from the genus
+\textit{Polyrhachis}. The sculpture identification protocol describes the two
+primary categories: \textit{rough} and \textit{smooth}.
 
@@
 
-\subsubsection{Data Collection}
+% \subsubsection{Data collection}
 
 To collect the images, the assistants followed the taxonomy information
-available in the master spreadsheet to the appropriate AntWeb page. In many
-cases, there are multiple ant head images of the same species, and occasionally
-there are multiple image resolution available from a single image. To simplify
-the data collection process, the assistants were instructed to download the
-first ant head image of the species being identified in the highest resolution
-possible. Each image was named with an identifier that corresponds with the row
-number in the master spreadsheet. The same ant head images that were downloaded
-in the data collection phase were the same ones used in the sculpture
-identification protocol. Ant species which did not have any images of the head
-were excluded from the dataset. Additionally, ant species which only had a head
-image of a queen ant were excluded from the dataset.
+available in the master spreadsheet to the appropriate AntWeb page
+\cite{perrichot_antweb_2012}. In many cases, there are multiple ant head images
+of the same species, and occasionally there are multiple image resolution
+available from a single image. To simplify the data collection process, the
+assistants were instructed to download the first ant head image of the species
+being identified in the highest resolution possible. Each image was named with
+an identifier that corresponds with the row number in the master spreadsheet.
+The same ant head images that were downloaded in the data collection phase were
+the same ones used in the sculpture identification protocol. Ant species which
+did not have any images of the head were excluded from the dataset.
+Additionally, ant species which only had a head image of a queen ant were
+excluded from the dataset.
 
@@
 sizes of different ant species. In a few ant head images, there exists some text
-denoting the specimen identifier and other info. In terms of texture, some ant
-specimens are very old, so their head images have other abnormalities such as
-cracks in the cuticle and the presence of dust.
+denoting the specimen identifier and other information. In terms of texture,
+some ant specimens are very old, so their head images have other abnormalities
+such as cracks in the cuticle and the presence of dust.
 
-\subsubsection{Data Preprocessing}
+% \subsubsection{Data preprocessing}
 
@@
 
-\subsection{Models}
+\subsection{Classical texture analysis methods}
+
+\colorbox{green}{TODO: kviews, gabor, etc. will go here}
+
+\subsection{Deep learning models used for the experiments}
 
@@
 
-\subsubsection{Evaluation}
+\colorbox{green}{\parbox{\dimexpr\textwidth-2\fboxsep}{TODO: I suggest that
+        at least one deep learning based texture classification algorithm be applied
+        to the dataset for the comparative study. When this is the case, one more
+        subsection should be included here to describe that/those algorithm(s).}}
+
+\subsection{Evaluation}
 
@@
 methods. Since we are working with a binary classification problem, we use a
-standard confusion matrix to evaluate the accuracy, precision, and F1 score. We
-also apply Grad-CAM with manual inspection to visualize the activation weights
-for incorrectly classified images to determine which features are interfering
-with the classification. Finally, we apply t-SNE to visualize the separation
+standard confusion matrix to evaluate the accuracy, precision, and F1 score.
+\colorbox{green}{(TODO: AUC and ROC can also be used as additional evaluation
+    methods.)} We also apply Grad-CAM with manual inspection to visualize the
+activation weights for classified images to visualize which features lead to the
+classification result. Finally, we apply t-SNE to visualize the separation
 learned for the model to further analyze the classifications made by the model.
@@
 each model. Then, the statistics for each VGG architecture are shared in Table
-\ref{tab:vgg_results} and Figure \ref{fig:vgg_results}. The results are
-collected on 8 iterations of training and averaged. The results are
-rounded to 2 decimal places where appropriate.
+\ref{tab:vgg_results}. The results are collected on 8 runs of training and
+averaged. The results are rounded to 2 decimal places where appropriate.
 
-\vspace{0.5in}
+% \begin{table}[h]
+%     \centering
+%     \caption{VGG11 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 101.75     & 53          \\
+%         Predicted Smooth & 37.375     & 257.875     \\
+%     \end{tabular}
+%     \label{tab:vgg11_confusion_matrix}
+% \end{table}
 
-\begin{table}[h]
-    \centering
-    \caption{VGG11 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 101.75     & 53          \\
-        Predicted Smooth & 37.375     & 257.875     \\
-    \end{tabular}
-    \label{tab:vgg11_confusion_matrix}
-\end{table}
+% \begin{table}[h]
+%     \centering
+%     \caption{VGG13 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 95.25      & 39.375      \\
+%         Predicted Smooth & 43.25      & 272.25      \\
+%     \end{tabular}
+%     \label{tab:vgg13_confusion_matrix}
+% \end{table}
 
-\vspace{0.5in}
+% \begin{table}[h]
+%     \centering
+%     \caption{VGG16 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 87.5       & 42.625      \\
+%         Predicted Smooth & 47.375     & 272.5       \\
+%     \end{tabular}
+%     \label{tab:vgg16_confusion_matrix}
+% \end{table}
 
-\begin{table}[h]
-    \centering
-    \caption{VGG13 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 95.25      & 39.375      \\
-        Predicted Smooth & 43.25      & 272.25      \\
-    \end{tabular}
-    \label{tab:vgg13_confusion_matrix}
-\end{table}
-
-\vspace{0.5in}
-
-\begin{table}[h]
-    \centering
-    \caption{VGG16 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 87.5       & 42.625      \\
-        Predicted Smooth & 47.375     & 272.5       \\
-    \end{tabular}
-    \label{tab:vgg16_confusion_matrix}
-\end{table}
-
-\vspace{0.5in}
-
-\begin{table}[h]
-    \centering
-    \caption{VGG19 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 91.25      & 42.125      \\
-        Predicted Smooth & 43.25      & 273.375     \\
-    \end{tabular}
-    \label{tab:vgg19_confusion_matrix}
-\end{table}
-
-\vspace{0.5in}
+% \begin{table}[h]
+%     \centering
+%     \caption{VGG19 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 91.25      & 42.125      \\
+%         Predicted Smooth & 43.25      & 273.375     \\
+%     \end{tabular}
+%     \label{tab:vgg19_confusion_matrix}
+% \end{table}
 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.6\textwidth]{thesis_assets/plots/vgg_average.png}
-    \caption{Average accuracy for VGG architectures on ant head image dataset.}
-    \label{fig:vgg_results}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.6\textwidth]{thesis_assets/plots/vgg_average.png}
+%     \caption{Average accuracy for VGG architectures on ant head image dataset.}
+%     \label{fig:vgg_results}
+% \end{figure}
 
@@
 summarized in a confusion matrix for each model. Then, the statistics for each
-ResNet architecture are shared in Table \ref{tab:resnet_results} and Figure
-\ref{fig:resnet_results}. The results are collected on 8 iterations of training and
-averaged. The results are rounded to 2 decimal places where appropriate.
+ResNet architecture are shared in Table \ref{tab:resnet_results}. The results
+are collected on 8 runs of training and averaged. The results are rounded
+to 2 decimal places where appropriate.
 
-\begin{table}[h]
-    \centering
-    \caption{ResNet18 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 98.75      & 48.75       \\
-        Predicted Smooth & 39.25      & 262         \\
-    \end{tabular}
-    \label{tab:resnet18_confusion_matrix}
-\end{table}
+% \begin{table}[h]
+%     \centering
+%     \caption{ResNet18 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 98.75      & 48.75       \\
+%         Predicted Smooth & 39.25      & 262         \\
+%     \end{tabular}
+%     \label{tab:resnet18_confusion_matrix}
+% \end{table}
 
-\vspace{0.5in}
+% \begin{table}[h]
+%     \centering
+%     \caption{ResNet50 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 81.375     & 43.625      \\
+%         Predicted Smooth & 55.875     & 269.125     \\
+%     \end{tabular}
+%     \label{tab:resnet50_confusion_matrix}
+% \end{table}
 
-\begin{table}[h]
-    \centering
-    \caption{ResNet50 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 81.375     & 43.625      \\
-        Predicted Smooth & 55.875     & 269.125     \\
-    \end{tabular}
-    \label{tab:resnet50_confusion_matrix}
-\end{table}
-
-\begin{table}[h]
-    \centering
-    \caption{ResNet101 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 85.75      & 53.875      \\
-        Predicted Smooth & 50.125     & 260.25      \\
-    \end{tabular}
-    \label{tab:resnet101_confusion_matrix}
-\end{table}
+% \begin{table}[h]
+%     \centering
+%     \caption{ResNet101 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 85.75      & 53.875      \\
+%         Predicted Smooth & 50.125     & 260.25      \\
+%     \end{tabular}
+%     \label{tab:resnet101_confusion_matrix}
+% \end{table}
 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.6\textwidth]{thesis_assets/plots/resnet_average.png}
-    \caption{Average accuracy for ResNet architectures on ant head image dataset.}
-    \label{fig:resnet_results}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.6\textwidth]{thesis_assets/plots/resnet_average.png}
+%     \caption{Average accuracy for ResNet architectures on ant head image dataset.}
+%     \label{fig:resnet_results}
+% \end{figure}
 
@@
 summarized in a confusion matrix for each model. Then, the statistics for each
-ResNet architecture are shared in Table \ref{tab:fresnet_results} and Figure
-\ref{fig:fresnet_results}. The results are collected on 8 iterations of training
-and averaged. The results are rounded to 2 decimal places where appropriate.
+ResNet architecture are shared in Table \ref{tab:fresnet_results}. The results
+are collected on 8 runs of training and averaged. The results are rounded
+to 2 decimal places where appropriate.
 
-\begin{table}[h]
-    \centering
-    \caption{Fine-tuned ResNet18 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 117.375    & 35.125      \\
-        Predicted Smooth & 17.875     & 279.625     \\
-    \end{tabular}
-    \label{tab:fresnet18_confusion_matrix}
-\end{table}
+% \begin{table}[h]
+%     \centering
+%     \caption{Fine-tuned ResNet18 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 117.375    & 35.125      \\
+%         Predicted Smooth & 17.875     & 279.625     \\
+%     \end{tabular}
+%     \label{tab:fresnet18_confusion_matrix}
+% \end{table}
 
-\vspace{0.5in}
+% \begin{table}[h]
+%     \centering
+%     \caption{Fine-tuned ResNet50 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 114.5      & 33.25       \\
+%         Predicted Smooth & 24.125     & 278.125     \\
+%     \end{tabular}
+%     \label{tab:fresnet50_confusion_matrix}
+% \end{table}
 
-\begin{table}[h]
-    \centering
-    \caption{Fine-tuned ResNet50 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 114.5      & 33.25       \\
-        Predicted Smooth & 24.125     & 278.125     \\
-    \end{tabular}
-    \label{tab:fresnet50_confusion_matrix}
-\end{table}
-
-\vspace{0.5in}
-
-\begin{table}[h]
-    \centering
-    \caption{Fine-tuned ResNet101 confusion matrix on ant head image dataset.}
-    \begin{tabular}{r|ll}
-                         & True Rough & True Smooth \\
-        \hline
-        Predicted Rough  & 117.25     & 33.25       \\
-        Predicted Smooth & 18.375     & 281.125     \\
-    \end{tabular}
-    \label{tab:fresnet101_confusion_matrix}
-\end{table}
-
-\vspace{0.5in}
+% \begin{table}[h]
+%     \centering
+%     \caption{Fine-tuned ResNet101 confusion matrix on ant head image dataset.}
+%     \begin{tabular}{r|ll}
+%                          & True Rough & True Smooth \\
+%         \hline
+%         Predicted Rough  & 117.25     & 33.25       \\
+%         Predicted Smooth & 18.375     & 281.125     \\
+%     \end{tabular}
+%     \label{tab:fresnet101_confusion_matrix}
+% \end{table}
 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.6\textwidth]{thesis_assets/plots/resnet_finetuned_average.png}
-    \caption{Average accuracy for fine-tuned ResNet architectures on ant head image dataset.}
-    \label{fig:fresnet_results}
-\end{figure}
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.6\textwidth]{thesis_assets/plots/resnet_finetuned_average.png}
+%     \caption{Average accuracy for fine-tuned ResNet architectures on ant head image dataset.}
+%     \label{fig:fresnet_results}
+% \end{figure}
 
-\clearpage
-
-\section{Analysis}
+\section{The analysis with visualization}
 
@@
 subcategories in our analysis. The two categories are correct and incorrect
-classification. Regardles of the feature activation map, the correctly
+classification. Regardless of the feature activation map, the correctly
 classified images have the same predicted label as the ground truth, and
@@
 
-\clearpage
+\FloatBarrier
+% \clearpage
 \section{Conclusion}
@@
 evaluating the function based on thousands of species is infeasible. We have
-shown in this work that a deep learning approach can be used to automatically
-categorize ants based on their cuticle texture, therefore supporting research on
-the evaluation of the function in future work. Our categorization system is
-novel in the field of automated insect indentification due to the broad number
-of species captured by it. Additionally, a model that is pre-trained on a
-diverse image task such as ResNet can be transferred to our domain of texture
-analysis. All code is publicly available on GitHub
-(https://github.com/ngngardner/cuticulus).
+shown in this work that a deep learning approach and classical image texture
+analysis methods can be used to automatically categorize ants based on their
+cuticle texture, therefore supporting research on the evaluation of the function
+in future work. Our categorization system is novel in the field of automated
+insect identification due to the broad number of species captured by it.
+Additionally, a model that is pre-trained on a diverse image task such as ResNet
+can be transferred to our domain of texture analysis. All code is publicly
+available on GitHub (https://github.com/ngngardner/cuticulus).
 
 \bibliographystyle{AIMS}
-\bibliography{main.bib}
+\bibliography{main.bib,article.bib}
 \end{document}

### 2022-06-10 cuticulus/__main__.py

@@ -15,7 +15,7 @@ def main():
     build_collage(
         ds.labels,
         ds.images,
-        name='bg',
+        name='rs',
         rows=10,
         cols=10,
     )

### 2022-06-10 cuticulus/core/datasets/sub.py

@@ -137,17 +137,16 @@ class SubDataset(DatasetSplitter):
         length = min(rows, cols) // 2 - 1
 
         for point in points:
-            # clip points outside of the new squared image
-            if point[0] > length:
-                point[0] = length
-            if point[1] > length:
-                point[1] = length
-
             if self.source_size != const.NO_RESIZE:
                 # scale the points to the new scale
                 point[0] = point[0] * self.source_size[0] / length
                 point[1] = point[1] * self.source_size[1] / length
 
+            # clip points outside of the new squared image
+            if point[0] > length:
+                point[0] = length
+            if point[1] > length:
+                point[1] = length
         return points
 
     @beartype

### 2022-06-17 paper/main.tex
gaussian filter test
add deep texture section

@@
     Evan Zhu\affil{1},
-    Zhiling Long\affil{3},
-    Min Wang\affil{2},
+    Zhiling Long\affil{4},
+    Min Wang\affil{3},
     Clint Penick\affil{2},
@@
         University, Kennesaw, GA, USA}%
-    \addr{\affilnum{3}}{Department of Mathematics, Science, and Informatics,
+    \addr{\affilnum{3}}{Department of Mathematics, Kennesaw State University,
+        Kennesaw, GA, USA}%
+    \addr{\affilnum{4}}{Department of Mathematics, Science, and Informatics,
         Mercer University, Atlanta, GA, USA}%
@@
 
-\colorbox{green}{TODO: Specifically, what approaches used? Why do we use image
-    texture for ant images? What are the contributions?}
+\colorbox{green}{\parbox{\dimexpr\textwidth-2\fboxsep}{TODO:Specifically, what
+        approaches used? Why do we use image texture for ant images? What are
+        the contributions?}}
 
@@
     IJCV2019 survey paper}
+
+\subsection{Deep Learning for Texture Analysis}
+
+In the past decade, deep learning using convolutional neural networks (CNNs)
+\cite{krizhevsky_imagenet_2017} has emerged as the state-of-the-art technology
+for image analysis. Following this trend, various CNN-based network
+architectures have been designed to specifically characterize texture images.
+Among these, the Fisher-vector CNN descriptor (FV-CNN) proposed by Cimpoi et al.
+\cite{cimpoi_describing_2014} is widely accepted as one of the most important
+pioneering works. It applies FV pooling to deep features obtained via a CNN
+pre-trained using the ImageNet \cite{krizhevsky_imagenet_2017} to obtain encoded
+features for texture classification. Although it is capable of achieving much
+improved classification accuracy comparing to traditional hand-crafted texture
+features, FV-CNN does not support an end-to-end learning, where feature
+extraction, encoding, and classification are separated from each other. To
+achieve an end-to-end learning, Zhang et al. \cite{zhang_visual_2018} proposed
+the deep texture encoding network (DeepTEN), in which a novel texture encoding
+layer is added to a standard CNN architecture. Then, Xue et al.
+\cite{xue_deep_2018} constructed the deep encoding pooling network (DEP), which
+improves over DeepTEN by integrating local spatial characteristics into the
+texture representation. Based on DeepTEN and DEP, Hu et al. \cite{hu_multi_2019}
+further developed the multi-level texture encoding and representation (MuLTER)
+network, which embeds a learnable encoding module at each convolutional layer so
+that encoding is performed for both low-level and high-level features, yielding
+a multi-level texture representation.
+
+Other network architectures for end-to-end texture learning are also available.
+For example, in the deep multiple-attribute-perceived network (MAP-Net)
+\cite{zhai_deep_2019}, multiple perceptual attributes are progressively learned
+in a mutually reinforced manner through multiple branches. In the deep
+structure-revealed network (DSR-Net)\cite{zhai_deep_2020}, inherent structural
+representation for a texture pattern is obtained by employing a primitive
+capturing module to learn spatial primitives and a dependency learning module to
+capture the dependency among the primitives. In \cite{mao_deep_2021}, a residual
+pooling layer consisting of a residual encoding module and an aggregation module
+is used to generate discriminative features of low dimensions. In
+\cite{peeples_histogram_2021}, a histogram layer is designed to compute local
+spatial distribution of CNN features. In \cite{chen_deep_2021}, an innovative
+aggregation module is presented to exploit statistical self-similarity across
+layers. All these architectures customize the standard CNN structure to
+accomplish the characterization of certain spatial, visual, or statistical
+nature unique to texture images.
 
@@
 
-% \begin{table}[h]
-%     \centering
-%     \caption{VGG11 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 101.75     & 53          \\
-%         Predicted Smooth & 37.375     & 257.875     \\
-%     \end{tabular}
-%     \label{tab:vgg11_confusion_matrix}
-% \end{table}
-
-% \begin{table}[h]
-%     \centering
-%     \caption{VGG13 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 95.25      & 39.375      \\
-%         Predicted Smooth & 43.25      & 272.25      \\
-%     \end{tabular}
-%     \label{tab:vgg13_confusion_matrix}
-% \end{table}
-
-% \begin{table}[h]
-%     \centering
-%     \caption{VGG16 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 87.5       & 42.625      \\
-%         Predicted Smooth & 47.375     & 272.5       \\
-%     \end{tabular}
-%     \label{tab:vgg16_confusion_matrix}
-% \end{table}
-
-% \begin{table}[h]
-%     \centering
-%     \caption{VGG19 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 91.25      & 42.125      \\
-%         Predicted Smooth & 43.25      & 273.375     \\
-%     \end{tabular}
-%     \label{tab:vgg19_confusion_matrix}
-% \end{table}
-
 \begin{table}[h]
@@
 
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.6\textwidth]{thesis_assets/plots/vgg_average.png}
-%     \caption{Average accuracy for VGG architectures on ant head image dataset.}
-%     \label{fig:vgg_results}
-% \end{figure}
-
 \subsection{ResNet Models}
@@
 to 2 decimal places where appropriate.
-
-% \begin{table}[h]
-%     \centering
-%     \caption{ResNet18 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 98.75      & 48.75       \\
-%         Predicted Smooth & 39.25      & 262         \\
-%     \end{tabular}
-%     \label{tab:resnet18_confusion_matrix}
-% \end{table}
-
-% \begin{table}[h]
-%     \centering
-%     \caption{ResNet50 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 81.375     & 43.625      \\
-%         Predicted Smooth & 55.875     & 269.125     \\
-%     \end{tabular}
-%     \label{tab:resnet50_confusion_matrix}
-% \end{table}
-
-% \begin{table}[h]
-%     \centering
-%     \caption{ResNet101 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 85.75      & 53.875      \\
-%         Predicted Smooth & 50.125     & 260.25      \\
-%     \end{tabular}
-%     \label{tab:resnet101_confusion_matrix}
-% \end{table}
 
@@
 
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.6\textwidth]{thesis_assets/plots/resnet_average.png}
-%     \caption{Average accuracy for ResNet architectures on ant head image dataset.}
-%     \label{fig:resnet_results}
-% \end{figure}
-
 \clearpage
@@
 to 2 decimal places where appropriate.
-
-% \begin{table}[h]
-%     \centering
-%     \caption{Fine-tuned ResNet18 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 117.375    & 35.125      \\
-%         Predicted Smooth & 17.875     & 279.625     \\
-%     \end{tabular}
-%     \label{tab:fresnet18_confusion_matrix}
-% \end{table}
-
-% \begin{table}[h]
-%     \centering
-%     \caption{Fine-tuned ResNet50 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 114.5      & 33.25       \\
-%         Predicted Smooth & 24.125     & 278.125     \\
-%     \end{tabular}
-%     \label{tab:fresnet50_confusion_matrix}
-% \end{table}
-
-% \begin{table}[h]
-%     \centering
-%     \caption{Fine-tuned ResNet101 confusion matrix on ant head image dataset.}
-%     \begin{tabular}{r|ll}
-%                          & True Rough & True Smooth \\
-%         \hline
-%         Predicted Rough  & 117.25     & 33.25       \\
-%         Predicted Smooth & 18.375     & 281.125     \\
-%     \end{tabular}
-%     \label{tab:fresnet101_confusion_matrix}
-% \end{table}
 
@@
 \end{table}
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.6\textwidth]{thesis_assets/plots/resnet_finetuned_average.png}
-%     \caption{Average accuracy for fine-tuned ResNet architectures on ant head image dataset.}
-%     \label{fig:fresnet_results}
-% \end{figure}
 

### 2022-06-17 scripts/experiments/__init__.py

@@ -1,1 +1,0 @@
-"""Experimental results for the dataset."""

### 2022-06-17 scripts/experiments/torch_utils.py

@@ -1,219 +1,0 @@
-"""Utility functions for torch experiments."""
-
-import copy
-import time
-
-import numpy as np
-import torch
-
-from cuticulus.console import console
-
-
-class TorchDS(torch.utils.data.Dataset):
-    """Torch dataset class for ant image dataset."""
-
-    def __init__(self, imgs: np.ndarray, labels: np.ndarray):
-        """Initialize dataset.
-
-        Args:
-            imgs (np.ndarray): List of data.
-            labels (np.ndarray): List of labels.
-        """
-        self.imgs = imgs
-        self.labels = labels
-
-    def __len__(self) -> int:
-        """Return length of dataset.
-
-        Returns:
-            int: Length of dataset.
-        """
-        return len(self.imgs)
-
-    def __getitem__(self, idx) -> tuple:
-        """Return item at index idx.
-
-        Args:
-            idx (int): Index of item.
-
-        Returns:
-            tuple: Tuple of image and label.
-        """
-        return self.imgs[idx], self.labels[idx]
-
-
-def inception_loss(
-    model: torch.nn.Module,
-    inputs: torch.Tensor,
-    labels: torch.Tensor,
-    criterion,
-) -> tuple[torch.Tensor, float]:
-    """Calculate inception loss.
-
-    Args:
-        model (torch.nn.Module): Model to calculate inception loss for.
-        inputs (torch.Tensor): Inputs to calculate inception loss for.
-        labels (torch.Tensor): Labels for inputs.
-        criterion (torch.nn.modules.loss._Loss): Loss function.
-
-    Returns:
-        tuple[torch.Tensor, float]: Tuple of outputs and loss.
-    """
-    outputs, aux_outputs = model(inputs)
-    loss1 = criterion(outputs, labels)
-    loss2 = criterion(aux_outputs, labels)
-    inception_param = 0.4
-    loss = loss1 + inception_param * loss2
-    return outputs, loss
-
-
-def train_step(
-    model: torch.nn.Module,
-    phase: str,
-    inputs: torch.Tensor,
-    labels: torch.Tensor,
-    criterion: torch.nn.modules.loss._Loss,
-    optimizer: torch.optim.Optimizer,
-    device: torch.device,
-    is_inception: bool = False,
-) -> tuple[float, int]:
-    """One iteration of training.
-
-    Args:
-        model (torch.nn.Module): Model to train.
-        phase (str): Phase to train for ('train' or 'val').
[…]

### 2022-06-17 scripts/notebook.jl

@@ -1,1077 +1,0 @@
-### A Pluto.jl notebook ###
-# v0.18.4
-
-using Markdown
-using InteractiveUtils
-
-# ╔═╡ 1a8051c6-8534-11ec-1183-91c02c8e3bbf
-begin
-	using DocStringExtensions
-	using FileIO
-	using Images
-	using NPZ
-	using Random
-	using MultivariateStats
-end
-
-# ╔═╡ 3e8a30a7-19e7-46ee-8a6f-b86af478303b
-Threads.nthreads()
-
-# ╔═╡ 899e59b5-52ac-4e35-b77a-3ba0535f495d
-begin
-	BASE_PATH = "../dataset"
-	DATA_PATH = "$BASE_PATH/data"
-	METADATA = npzread(string(BASE_PATH, "/meta.npy"))
-
-	IMAGE_SIZE = (224, 224)
-	BATCH_SIZE = 4
-	VAL_SPLIT = 0.1
-	LR = 3e-4
-	EPOCHS = 1
-end
-
-# ╔═╡ eb05fa5d-e127-4f9b-b15a-ef854eb830b8
-begin
-	re = r"[\d]+\.jpg"
-	image_paths = readdir(DATA_PATH)
-	image_paths = filter(x -> occursin(re, x), image_paths)
-end
-
-# ╔═╡ ee1820d2-f516-48cd-8f47-81c0c1b6f0ce
-"""
-Given an image id, return the label.
-"""
-function get_label(id::Int, meta::Array{Int64, 2} = METADATA)::Int64
-	col = findfirst(x -> x == id, METADATA[1, :])
-	if col == nothing
-		return 0
-	end
-	return meta[2, col]
-end
-
-# ╔═╡ e08a6400-fd47-4a9c-85c0-2558c2e0f1cd
-"""
-Get list of image ids.
-
-From a vector of image filenames of format "x.y", where x is the id, and y is any file format, a vector of ids is returned.
-"""
-function get_ids(paths::Vector{String})::Vector{String}
-	l = size(paths)[1]
-	res = Vector{String}(undef, l)
-	for idx in 1:l
-		res[idx] = split(paths[idx], ".")[1]
-	end
-	return res
-end
-
-# ╔═╡ 8473c673-ff11-49d3-ac5c-bf19eded885e
-ids = get_ids(image_paths)
-
[…]

### 2022-06-17 paper/main.tex
post meeting updates

@@
 
-\colorbox{green}{TODO: how do we connect this section to image texture analysis?
-    (why do we research this and why image texture) }
+\colorbox{green}{\parbox{\dimexpr\textwidth-2\fboxsep}{TODO: how do we connect
+    this section to image texture analysis? (why do we research this and why
+    image texture) }}
 
 \subsection{Texture Analysis}
-\colorbox{green}{TODO: A brief survey on image texture and cite Liu's
-    IJCV2019 survey paper}
+Image texture analysis is an important part of image interpretation as many
+objects appearing in the textural patterns either partially or completely inside
+an image. This phenomenon appears frequently in many images we encountered in
+our environment as almost everybody is using the iPhone to capture the photos.
+Scientists are also studying their subjects depending on taking the images
+remotely such as remotely sensed satellite images. With an abundant number of
+textural images surrounding us, it is essential to have a reliable automatic
+tool for the textural image interpretation.
 
-\subsection{Deep Learning for Texture Analysis}
+Classical computer methods are developed for textural images based on the
+spatial relationship of gray levels of pixels encoded in the texture
+\colorbox{green}{[Haralick and Shapiro, Hung]}. Those methods by de-facto follow
+the steps used in the statistical pattern recognition
+\colorbox{green}{[Fukunaga]}. Therefore, textural features are extracted, and
+the classification and segmentation methods are used for the interpretation
+either using the supervised or unsupervised approaches. The interpretation
+method is usually taken and modified from the traditional pattern recognition
+methods. For example, the gray-level co-occurrence matrix (GLCM) and local
+binary patterns (LBP) are two of methods for extracting features
+\colorbox{green}{[references]}. Some other approaches also have been developed
+for the textural feature extraction \colorbox{green}{[references]}.
 
-In the past decade, deep learning using convolutional neural networks (CNNs)
+While most feature vectors are established based on a single pixel, Liu et al.
+gave a comprehensive survey on textural characterization in which they call this
+feature extraction as texture representation \colorbox{green}{[Liu]}. Their
+survey on Bag of Words and Convolutional Neural Networks are the spatial
+relationship concepts. The K-views was developed for taking the spatial
+information in the clustering approach \colorbox{green}{[Hung]}.  In the past
+decade, the deep learning network has become the mainstream in image
+classification and segmentation \colorbox{green}{[references]}. Compared with
+the traditional method in which a kernel must be designed by an engineer for
+extracting features, the deep learning network can automatically extract the
+features through the training. In addition, the deep learning networks achieve
+the higher accuracy than that of the classical approaches although the deep
+networks require much more data for training.
+
+Similar to the traditional pattern recognition, classical image texture analysis
+methods can be grouped into four categories: statistical, structural,
+model-based, and transform-based methods \colorbox{green}{[references]}. Among
+all of the categories, the statistical method is often used by employing the
+method from the pattern recognition principle.  In doing so, an image texture is
+represented as feature vectors and then fed into the algorithm. Markov random
+field models are also studied for textural image interpretation
+\colorbox{green}{[references]}. The transform-based methods use some functions
+to decompose an image texture into a set of basic feature images. Gabor filters
+and Wavelet expansions are two of the widely used approaches
+\colorbox{green}{[references]}.
+
+\colorbox{green}{TODO: add reference for sun phd thesis (Robust texture classification)}
+
+\colorbox{green}{TODO: cleanup transition}
+
+Recently, the deep learning network becomes the main stream in the
+interpretation of textural images [Sun and liu]. In the past decade, deep
+learning using convolutional neural networks (CNNs)
 \cite{krizhevsky_imagenet_2017} has emerged as the state-of-the-art technology
@@
 
-\colorbox{green}{TODO: kviews, gabor, etc. will go here}
+\colorbox{green}{TODO: kviews, kmeans, etc. will go here}
 
@@
     \label{tab:fresnet_results}
+\end{table}
+
+\colorbox{green}{TODO: Fix tables, decide on how to group models}
+
+\begin{table}[h]
+    \centering
+    \caption{Results}
+    \begin{tabular}{lllllll}
+        \toprule
+        {} & model                     & accuracy & precision & recall & f1   & roc\_auc \\
+        \midrule
+        0  & drp\_multi\_not\_pt       & 0.87     & 0.89      & 0.92   & 0.91 & 0.86     \\
+        1  & drp\_multi\_pt            & 0.88     & 0.89      & 0.93   & 0.91 & 0.87     \\
+        2  & drp\_single\_not\_pt      & 0.88     & 0.9       & 0.94   & 0.92 & 0.88     \\
+        3  & drp\_single\_pt           & 0.88     & 0.9       & 0.93   & 0.91 & 0.87     \\
+        4  & drp\_single\_aux\_not\_pt & 0.87     & 0.89      & 0.92   & 0.9  & 0.86     \\
+        5  & drp\_single\_aux\_pt      & 0.89     & 0.9       & 0.94   & 0.92 & 0.89     \\
+        6  & kviews\_17                & 0.62     & 0.79      & 0.59   & 0.68 & 0.62     \\
+        7  & kviews\_19                & 0.62     & 0.8       & 0.59   & 0.68 & 0.62     \\
+        8  & kviews\_25                & 0.52     & 0.71      & 0.51   & 0.59 & 0.52     \\
+        9  & kmeans                    & 0.47     & 0.07      & 0.35   & 0.11 & 0.47     \\
+        \bottomrule
+    \end{tabular}
+\end{table}
+
+\begin{table}[h]
+    \centering
+    \caption{Clustering Results}
+    \begin{tabular}{llll}
+        \toprule
+        {} & model      & rand\_score & adjusted\_rand\_score \\
+        \midrule
+        0  & kviews\_17 & 0.53        & 0.06                  \\
+        1  & kviews\_19 & 0.53        & 0.06                  \\
+        2  & kviews\_25 & 0.5         & -0.0                  \\
+        3  & kmeans     & 0.5         & 0.0                   \\
+        \bottomrule
+    \end{tabular}
 \end{table}
@@
 
+\colorbox{green}{TODO: mention how future work may include extra classes}
+
 \bibliographystyle{AIMS}

### 2022-06-18 paper/main.tex
draft

--- 2022-06-17 paper/main.tex
@@
     Anthony Phan\affil{1},
-    Evan Zhu\affil{1},
+    Haige Zhu\affil{1},
     Zhiling Long\affil{4},
@@
     automatically group similar ant species based on morphological traits. We
-    provide a comparative study of the performance of texture analysis methods
-    on ant images. We evaluate the results of the classification methods with
-    modern visualization techniques.
-    \colorbox{green}{TODO: what is the result? experiments? data?}
+    provide a comparative study of the performance of general image
+    classification and texture analysis methods on ant head images. We evaluate
+    the results of the classification methods with modern visualization
+    techniques. Our results show that deep learning methods are applicable for
+    classifying ant cuticle textures. We also show that the performance of the
+    deep texture analysis methods performs better than the deep classification
+    methods.
 \end{abstract}
@@
 \section{Introduction}
-\colorbox{green}{TODO: refer to ant head images in a consistent way}
-
-\colorbox{green}{TODO: refer to traditional/classical in a consistent way}
-
-\colorbox{green}{TODO: change ideal and non-ideal to features identified by
-    the biologist and to invalid features (background, etc.)}
 
@@
 
-In image texture analysis methods, the general goal is to automatically
-categorize an object into a set of objects with similar texture-based features.
-Texture analysis has shown promising results in related fields, such as plant
-identification \cite{boudra_plant_2018} \colorbox{green}{TODO: more references
-    and examples}. With modern texture analysis methods, the categorization of ants
-can be automated and the results can be used to study the influence of cuticle
-texture on ant ecology.
+Our primary contribution is the custom dataset used to explore the relationship
+between ant cuticle texture and automated texture analysis methods. We compare
+classical and modern image classification methods and texture analysis methods
+by evaluating their performance on our dataset.
 
-Image texture analysis heavily depends on the spatial relationships among gray
-levels of pixels \cite{hung_image_2019}. Hence, feature extraction method is
-usually applied to the spatial relationship. The gray-level co-occurrence matrix
-(GLCM) and local binary patterns (LBP) are two of methods which are based on the
-spatial relationships to extracting features for this purpose
-\cite{haralick_textural_1973, goos_gray_2000}. Traditional image texture
-analysis can be divided into four categories: statistical, structural,
-model-based, and transform-based methods \colorbox{green}{[references]}. In the
-past decade, the deep learning network becomes the mainstream in image
-classification and segmentation \cite{liu_bow_2019}. Compared with the
-traditional method in which a kernel must be designed for extracting features,
-the deep learning network can automatically extract the features by itself.
+\section{Related work and literature review}
 
-In this study, we used some of both approaches for our work on the ant images.
+In this section we will review the literature for ant sculpturing identification
+and texture analysis. Additionally, we discuss the work related to automated
+insect identification.
 
-\colorbox{green}{\parbox{\dimexpr\textwidth-2\fboxsep}{TODO:Specifically, what
-        approaches used? Why do we use image texture for ant images? What are
-        the contributions?}}
-
-\colorbox{green}{TODO: briefly discuss results from the research}
-
-\section{Related Work}
-
-\colorbox{green}{TODO: introduce the section}
-
-\subsection{Sculpturing Identification}
+\subsection{Sculpturing identification}
 Taxonomists have developed extensive terminology describing ant cuticle
@@
     Sculpturing} — contains over 100 terms to describe the cuticle sculpturing
-patterns of ants \cite{harris_glossary_1979}. There is often substantial overlap
-among terms, and closely related cuticle patterns are likely to be functionally
-similar. For example, the definition for “imbricate” is, “partly overlapping and
-appearing like shingles on a roof or scales on a fish,” which is difficult to
-distinguish from the definition of tesselate, “made up of squares like a chess
-board, either in sculpturing or in color.”
+patterns of ants \cite{harris_glossary_1979}.
 
@@
 
-\colorbox{green}{TODO: how do we connect this section to image texture analysis?
-    (why do we research this and why image texture) }
+There is often substantial overlap among terms, and closely related cuticle
+patterns are likely to be functionally similar. For example, the definition for
+“imbricate” is, “partly overlapping and appearing like shingles on a roof or
+scales on a fish,” which is difficult to distinguish from the definition of
+tesselate, “made up of squares like a chess board, either in sculpturing or in
+color.” However, we note that the terms used to describe the cuticle patterns
+are similar to the terms used to describe textural patterns in texture analysis.
+In the next subsection, we review some literature of image texture analysis.
 
-\subsection{Texture Analysis}
-\colorbox{green}{TODO: A brief survey on image texture and cite Liu's
-    IJCV2019 survey paper}
+\subsection{Texture analysis}
 
-\subsection{Deep Learning for Texture Analysis}
+Image texture analysis is an important part of image interpretation as many
+objects appearing in the textural patterns either partially or completely inside
+an image. This phenomenon appears frequently in many images we encountered in
+our environment as almost everybody is using the iPhone to capture the photos.
+Scientists are also studying their subjects depending on taking the images
+remotely such as remotely sensed satellite images. With an abundant number of
+textural images surrounding us, it is essential to have a reliable automatic
+tool for the textural image interpretation.
 
-In the past decade, deep learning using convolutional neural networks (CNNs)
-\cite{krizhevsky_imagenet_2017} has emerged as the state-of-the-art technology
-for image analysis. Following this trend, various CNN-based network
-architectures have been designed to specifically characterize texture images.
-Among these, the Fisher-vector CNN descriptor (FV-CNN) proposed by Cimpoi et al.
-\cite{cimpoi_describing_2014} is widely accepted as one of the most important
+Classical computer methods are developed for textural images based on the
+spatial relationship of gray levels of pixels encoded in the texture
+\cite{haralick_image_1985,hung_image_2019}. Those methods by de-facto follow the
+steps used in the statistical pattern recognition
+\cite{fukunaga_introduction_2013}. Therefore, textural features are extracted,
+and the classification and segmentation methods are used for the interpretation
+either using the supervised or unsupervised approaches. The interpretation
+method is usually taken and modified from the traditional pattern recognition
+methods. For example, the gray-level co-occurrence matrix (GLCM) and local
+binary patterns (LBP) are two of methods for extracting features
+\cite{haralick_textural_1973, goos_gray_2000}.
+
+Similar to the traditional pattern recognition, classical image texture analysis
+methods can be grouped into four categories: statistical, structural,
+model-based, and transform-based methods \cite{bharati_image_2004}. Among all of
+the categories, the statistical method is often used by employing the method
+from the pattern recognition principle.  In doing so, an image texture is
+represented as feature vectors and then fed into the algorithm. Markov random
+field models are also studied for textural image interpretation
+\cite{hassner_use_1981, cross_markov_1983}. The transform-based methods use some functions to
+decompose an image texture into a set of basic feature images. Gabor filters and
+Wavelet expansions are two of the widely used approaches
+\cite{bovik_multichannel_1990}.
+
+While most feature vectors are established based on a single pixel, Liu et al.
+gave a comprehensive survey on textural characterization in which they call this
+feature extraction as texture representation \cite{liu_bow_2019}. Their survey
+on Bag of Words and Convolutional Neural Networks (CNNs)
+\cite{krizhevsky_imagenet_2017} are the spatial relationship concepts. The
+K-views was developed for taking the spatial information in the clustering
+approach \cite{hung_image_2019}.
+
+In the past decade, deep learning using CNNs has emerged as the mainstream
+technology for image analysis. Following this trend, various CNN-based network
+architectures have been designed to specifically characterize textural images.
+Among these, the Fisher-vector CNN descriptor (FV-CNN) proposed by Cimpoi et
+al. \cite{cimpoi_deep_2015} is widely accepted as one of the most important
 pioneering works. It applies FV pooling to deep features obtained via a CNN
-pre-trained using the ImageNet \cite{krizhevsky_imagenet_2017} to obtain encoded
-features for texture classification. Although it is capable of achieving much
-improved classification accuracy comparing to traditional hand-crafted texture
-features, FV-CNN does not support an end-to-end learning, where feature
-extraction, encoding, and classification are separated from each other. To
-achieve an end-to-end learning, Zhang et al. \cite{zhang_visual_2018} proposed
-the deep texture encoding network (DeepTEN), in which a novel texture encoding
-layer is added to a standard CNN architecture. Then, Xue et al.
-\cite{xue_deep_2018} constructed the deep encoding pooling network (DEP), which
-improves over DeepTEN by integrating local spatial characteristics into the
-texture representation. Based on DeepTEN and DEP, Hu et al. \cite{hu_multi_2019}
-further developed the multi-level texture encoding and representation (MuLTER)
-network, which embeds a learnable encoding module at each convolutional layer so
-that encoding is performed for both low-level and high-level features, yielding
-a multi-level texture representation.
+pre-trained using the ImageNet \cite{krizhevsky_imagenet_2017} to obtain encoded features
+for texture classification. Although it is capable of achieving much improved
+classification accuracy comparing to traditional hand-crafted texture features,
+FV-CNN does not support an end-to-end learning, where feature extraction,
+encoding, and classification are separated from each other. To achieve an
+end-to-end learning, Zhang et al. \cite{zhang_deep_2017} proposed the deep texture
+encoding network (DeepTEN), in which a novel texture encoding layer is added to
+a standard CNN architecture. Then, Xue et al. \cite{xue_deep_2018} constructed the
+deep encoding pooling network (DEP), which improves over DeepTEN by integrating
+local spatial characteristics into the texture representation. Based on DeepTEN
+and DEP, Hu et al. \cite{hu_multi_2019} further developed the multi-level texture
+encoding and representation (MuLTER) network, which embeds a learnable encoding
+module at each convolutional layer so that encoding is performed for both
+low-level and high-level features, yielding a multi-level texture
+representation.
 
@@
 in a mutually reinforced manner through multiple branches. In the deep
-structure-revealed network (DSR-Net)\cite{zhai_deep_2020}, inherent structural
+structure-revealed network (DSR-Net) \cite{zhai_deep_2020}, inherent structural
 representation for a texture pattern is obtained by employing a primitive
@@
 accomplish the characterization of certain spatial, visual, or statistical
-nature unique to texture images.
+nature unique to textural images.
 
-\subsection{Insect Classification}
+Compared with the traditional method in which a kernel must be designed by an
+engineer for extracting features, the deep learning networks can automatically
+extract the features through the training. In addition, the deep learning
+networks achieve the higher accuracy than that of the classical approaches,
+although the deep networks require much more data for training.
+
+\subsection{Insect classification}
 In this section, we provide an overview of some insect classification methods.
@@
 family level, \textit{e.g.} \textit{ant}, \textit{fly}, and \textit{bee}
-\cite{imagenet_labels}. Lim et al. \cite{lim_performance_2017} use a modified
+\cite{deng_imagenet_2009}. Lim et al. \cite{lim_performance_2017} use a modified
 AlexNet architecture and experiment with different numbers of kernels and their
@@
 
-\colorbox{green}{TODO: do not jump into subsection directly without any context}
+In this section, we provide an overview of the methodology used in this paper.
+We describe the steps used to prepare the custom dataset. Finally, we describe
+the algorithms applied and experimental setup to obtain the results.
 
@@
 \cite{perrichot_antweb_2012} and define two categories for them based on the
-appearance of the cuticle texture: \textit{rough} and \textit{smooth}. Some
+appearance of the cuticle texture: \textit{rough} and \textit{smooth}. We also
+refer to these ant head images, which describe the pose of the ant, as ant
+cuticule images, due to the clear cuticle texture present on the head. Some
 randomly selected images from each category are shown in Figures
@@
 \end{figure}
-
-% \subsubsection{Sculpture identification protocol}
 
@@
 species on a majority basis.
-
-% \subsubsection{Data collection}
 
@@
 
-% \subsubsection{Data preprocessing}
-
 Due to the variety of the ant head image attributes, we apply simple
@@
 
-\colorbox{green}{TODO: kviews, gabor, etc. will go here}
+The K-means algorithm is one of widely used clustering algorithms in the pattern
+recognition \cite{lloyd_least_1982}. It is a single pixel based classification
+algorithm for images. The K-views is an algorithm that uses several
+characteristic views for the classification of images \cite{hung_use_2002}. The
+K-views algorithm is suitable for classifying image textures that have basic
+local patterns repeated in a periodic manner. In contrast to the K-means,
+K-views looks at the neighboring pixels for providing the spatial relationship
+for classification.  Several variations of the basic K-views model have been
+proposed \cite{yang_image_2003, lan_improved_2010}.
+
+For the feasibility study in this ant image datasets, we use both K-means and
+K-views in our experiments as both are statistical clustering methods. The
+experiments may indicate how well the statistical approach will do in this type
+of textural images. In pre-processing the dataset, we apply the Gabor filter to
+transform the gray-scale ant image into four Gabor-filtered channels, subtract
+the gray-scale image from each channel, take the absolute value of the resulting
+images, and then normalize them. This pre-processing will highlight some of
+characteristics features such as edge-like in the image. This step is very
+useful in feeding the datasets to the K-views algorithms. A 10-channel Gabor
+filter is applied in the pre-processing step for the K-means algorithm.
 
@@
 
-\colorbox{green}{\parbox{\dimexpr\textwidth-2\fboxsep}{TODO: I suggest that
-        at least one deep learning based texture classification algorithm be applied
-        to the dataset for the comparative study. When this is the case, one more
-        subsection should be included here to describe that/those algorithm(s).}}
+% \colorbox{green}{\parbox{\dimexpr\textwidth-2\fboxsep}{TODO: I suggest that at
+%         least one deep learning based texture classification algorithm be
+%         applied to the dataset for the comparative study. When this is the
+%         case, one more subsection should be included here to describe
+%         that/those algorithm(s).}}
 
@@
 methods. Since we are working with a binary classification problem, we use a
-standard confusion matrix to evaluate the accuracy, precision, and F1 score.
-\colorbox{green}{(TODO: AUC and ROC can also be used as additional evaluation
-    methods.)} We also apply Grad-CAM with manual inspection to visualize the
-activation weights for classified images to visualize which features lead to the
-classification result. Finally, we apply t-SNE to visualize the separation
-learned for the model to further analyze the classifications made by the model.
+standard confusion matrix to evaluate the accuracy, precision, and F1 score. We
+also apply Grad-CAM with manual inspection to visualize the activation weights
+for classified images to visualize which features lead to the classification
+result. Finally, we apply t-SNE to visualize the separation learned for the
+model to further analyze the classifications made by the model.
 
@@
 
-% vgg results
-\subsection{VGG Models}
+\subsection{Evaluation results}
 
-To begin, we share the results on the VGG model architectures on our custom
-dataset. The classification results are summarized in a confusion matrix for
-each model. Then, the statistics for each VGG architecture are shared in Table
-\ref{tab:vgg_results}. The results are collected on 8 runs of training and
-averaged. The results are rounded to 2 decimal places where appropriate.
+To begin, we share the results on the each algorithm specified in the
+methodology section using the evaluation metrics described. The results are
+shown in Table \ref{tab:results}. It should be noted that due to the class
+imbalance in the dataset, the F1 score is the preferable metric to the accuracy.
+Therefore, the results have been sorted by F1 score in descending order.
 
@@
     \centering
-    \caption{Average results for VGG architectures on ant head image dataset.}
-    \begin{tabular}{r|llll}
-              & Recall & Precision & F1 Score & Accuracy \\
-        \hline
-        VGG11 & 0.83   & 0.88      & 0.85     & 0.80     \\
-        VGG13 & 0.87   & 0.87      & 0.87     & 0.82     \\
-        VGG16 & 0.86   & 0.86      & 0.86     & 0.80     \\
-        VGG19 & 0.87   & 0.87      & 0.87     & 0.81     \\
-    \end{tabular}
-    \label{tab:vgg_results}
-\end{table}
-
-\subsection{ResNet Models}
-
-Next, we share the results on the ResNet model architectures on our custom
-dataset with random weight initialization. The classification results are
-summarized in a confusion matrix for each model. Then, the statistics for each
-ResNet architecture are shared in Table \ref{tab:resnet_results}. The results
-are collected on 8 runs of training and averaged. The results are rounded
-to 2 decimal places where appropriate.
-
-\begin{table}[h]
-    \centering
-    \caption{Average results for ResNet architectures on ant head image dataset.}
-    \begin{tabular}{r|llll}
-                  & Recall & Precision & F1 Score & Accuracy \\
-        \hline
-        ResNet18  & 0.84   & 0.87      & 0.85     & 0.80     \\
-        ResNet50  & 0.86   & 0.83      & 0.84     & 0.77     \\
-        ResNet101 & 0.83   & 0.84      & 0.83     & 0.76     \\
-    \end{tabular}
-    \label{tab:resnet_results}
-\end{table}
-
-\clearpage
-\subsection{Fine-tuned ResNet Models}
-
-Next, we share the results on the fine-tuned ResNet model architectures on our
-custom dataset with pretrained weights. The classification results are
-summarized in a confusion matrix for each model. Then, the statistics for each
-ResNet architecture are shared in Table \ref{tab:fresnet_results}. The results
-are collected on 8 runs of training and averaged. The results are rounded
-to 2 decimal places where appropriate.
-
-\begin{table}[h]
-    \centering
-    \caption{Average results for Fine-tuned ResNet architectures on ant head image dataset.}
-    \begin{tabular}{r|llll}
-                  & Recall & Precision & F1 Score & Accuracy \\
-        \hline
-        ResNet18  & 0.89   & 0.94      & 0.91     & 0.88     \\
-        ResNet50  & 0.89   & 0.92      & 0.91     & 0.87     \\
-        ResNet101 & 0.89   & 0.94      & 0.92     & 0.88     \\
-    \end{tabular}
-    \label{tab:fresnet_results}
+    \caption{Results}
+    \input{assets/results.tex}
+    \label{tab:results}
 \end{table}
@@
 outperform the VGG and randomly initialized ResNet models on the task of ant
-head image classification. It should be noted that due to the class imbalance in
-the dataset, the F1 score is the preferable metric to the accuracy. On average,
-the fine-tuned ResNet101 model performed the best with an average F1 score of
-0.92. We further analyze the separation learned by both ResNet101 models in the
-following section.
+head image classification. Additionally, the DRP models perform better than the
+the general deep learning models VGG and ResNet on the task of ant head image
+classification.The fine-tuned DRP single-layer model with auxiliary classifier
+performed the best with an average F1 score of 0.92. We further analyze the
+separation learned by both ResNet101 models in the following section.
+
+% \colorbox{green}{(TODO: DRP is best performing model, it should be included in
+%     visualization.)}
 
@@
 
-\subsection{GradCAM Visualization}
+\subsection{GradCAM visualization}
 
@@
 incorrectly classified images have different predicted labels. The two
-subcategories are ideal and non-ideal feature activation. In the ideal case, the
-features that are used to compute the classification are the same as the
-features used by the assistants in the sculpture identification process. In
+subcategories are expected and non-expected feature activation. In the expected
+case, the features that are used to compute the classification are the same as
+the features used by the assistants in the sculpture identification process. In
 general, the features used by the assistants are the textures of the cuticle on
-the ant head. In the non-ideal case, the features used to compute the
+the ant head. In the non-expected case, the features used to compute the
 classification are not from the head, for example, from the background,
@@
 
-Correctly classified images which use ideal features show the ideal performance
-of the model. Incorrectly classified images which use ideal features should be
-further analyzed. In essence, the model in this situation knows \textit{where}
-to look, but not \textit{what} to look for. In Figure
+Correctly classified images which use the expected features show the ideal
+performance of the model. Incorrectly classified images which use the expected
+features should be further analyzed. In essence, the model in this situation
+knows \textit{where} to look, but not \textit{what} to look for. In Figure
 \ref{fig:incorrect_ideal_84}, the features activated are mostly in the correct
@@
 this case, it may be due to the pose of the ant being slightly different from
-the average pose. In the incorrectly classified images with non-ideal features,
-analysis shows that the model is unable to find \textit{where} to look, and
-obtains feature information from other parts of the ant or the background. Cases
-where the image was correctly classified using the non-ideal features can
-basically be seen as noise. In order to further analyze this class, we should
-introduce some parameter such as model confidence to examine further.
+the average pose. In the incorrectly classified images with non-expected
+features, analysis shows that the model is unable to find \textit{where} to
+look, and obtains feature information from other parts of the ant or the
+background. Cases where the image was correctly classified using the
+non-expected features can basically be seen as noise. In order to further
+analyze this class, we should introduce some parameter such as model confidence
+to examine further.
 
@@
     \end{subfigure}
-    \caption{Correctly classified images using ideal features.}
+    \caption{Correctly classified images using expected features.}
     \label{fig:correct_ideal}
@@
     \end{subfigure}
-    \caption{Correctly classified images using non-ideal features.}
+    \caption{Correctly classified images using non-expected features.}
     \label{fig:correct_nonideal}
@@
     \end{subfigure}
-    \caption{Incorrectly classified images using ideal features.}
+    \caption{Incorrectly classified images using expected features.}
     \label{fig:incorrect_ideal}
@@
     \end{subfigure}
-    \caption{Incorrectly classified images using non-ideal features.}
+    \caption{Incorrectly classified images using non-expected features.}
     \label{fig:incorrect_nonideal}
@@
 \FloatBarrier
-% \clearpage
 \section{Conclusion}
@@
 Additionally, a model that is pre-trained on a diverse image task such as ResNet
-can be transferred to our domain of texture analysis. All code is publicly
-available on GitHub (https://github.com/ngngardner/cuticulus).
+can be transferred to our domain of texture analysis. However, a deep learning
+algorithm created specifically for the domain of texture analysis had the best
+results in our experiments. In future work, we will continue to explore texture
+classes present in the dataset which are not captured by the binary class
+system. All code is publicly available on GitHub
+(https://github.com/ngngardner/cuticulus).
 

### 2022-06-24 paper/main.tex
add drp section

@@
 
-Our custom dataset of ant head images contains 2,499 images. 1072 samples of
-rough textured ant cuticle textures comprise 43\% of the dataset. The remaining
-1427 samples of smooth textured ant cuticle textures comprise 57\% of the
-dataset. To handle the imbalance of the dataset, we apply undersampling for each
-class for the training dataset. By using random stratified sampling, we
-construct a training set with 800 images per class. The remaining images are
-randomly split between test and validation, which turns out to roughly a
-60\%/20\%/20\% train, test, and validation data split. With 272 rough samples
-and 627 smooth samples left over after the stratified split, the test dataset
-has roughly 136 rough samples and 313 smooth samples. Since these leftover
-samples are split with code by 50\% there will be some rounding variance and
-therefore the test dataset built at run-time will not always have exactly the
-same number of samples.
-
 \subsection{Classical texture analysis methods}
@@
 
+\subsection{Deep texture analysis method used for the experiments}
+
+In addition to the classical texture analysis methods and the deep learning
+models for general image analysis, we also examine deep learning analysis
+methods specifically designed for textural images. For this examination, we
+adopt the deep residual pooling (DRP) network developed by Mao et al.
+\cite{mao_deep_2021}. The DRP framework consists of a unique residual pooling
+layer, which is formed by a residual encoding module followed by an aggregation
+module. The residual encoding module extracts relevant spatial information,
+while the aggregation module performs averaging to obtain orderless
+low-dimension features for classification.
+
+According to \cite{mao_deep_2021}, the DRP network can be used either with a
+single residual pooling layer, applied right before the classifier layer, or
+with multiple residual pooling layers, applied to the output of multiple
+convolutional layers before feeding the combined pooling results into the
+classifier layer. Additionally, an auxiliary classifier layer may also be used.
+In our experiments, we follow these strategies. For each scenario, we conduct
+two experiments. One experiment employs weights that are randomly initialized,
+and the other uses weights that are fine tuned over those from a pretrained
+model.
+
+\section{Results and analysis}
+\subsection{Environment}
+
+Experiments are run on an Ubuntu 18.04 LTS Lambda Labs GPU server. The server
+contains 8 NVIDIA GeForce RTX 2080 Ti graphics cards with 12GB of memory each.
+The server uses an Intel Xeon Silver 4116 with 48 total threads and maximum
+frequency of 3.000 GHz, and has 256GB of RAM.
+
+\subsection{Evaluation}
+
+We evaluate the performance of the models according to standard evaluation
+methods. Since we are working with a binary classification problem, we use a
+standard confusion matrix to evaluate the accuracy, precision, and F1 score. We
+also apply Grad-CAM with manual inspection to visualize the activation weights
+for classified images to visualize which features lead to the classification
+result. Finally, we apply t-SNE to visualize the separation learned for the
+model to further analyze the classifications made by the model.
+
+\subsection{Experimental results}
+
 For our ResNet models, we have two versions: randomized and pretrained. The
@@
 
-% \colorbox{green}{\parbox{\dimexpr\textwidth-2\fboxsep}{TODO: I suggest that at
-%         least one deep learning based texture classification algorithm be
-%         applied to the dataset for the comparative study. When this is the
-%         case, one more subsection should be included here to describe
-%         that/those algorithm(s).}}
-
-\subsection{Evaluation}
-
-We evaluate the performance of the models according to standard evaluation
-methods. Since we are working with a binary classification problem, we use a
-standard confusion matrix to evaluate the accuracy, precision, and F1 score. We
-also apply Grad-CAM with manual inspection to visualize the activation weights
-for classified images to visualize which features lead to the classification
-result. Finally, we apply t-SNE to visualize the separation learned for the
-model to further analyze the classifications made by the model.
-
-\section{Results}
-\subsection{Environment}
-
-Experiments are run on an Ubuntu 18.04 LTS Lambda Labs GPU server. The server
-contains 8 NVIDIA GeForce RTX 2080 Ti graphics cards with 12GB of memory each.
-The server uses an Intel Xeon Silver 4116 with 48 total threads and maximum
-frequency of 3.000 GHz, and has 256GB of RAM.
-
-\subsection{Evaluation results}
+For the experiments using the deep learning algorithms, we use our custom
+dataset of ant head images which contains 2,499 images. 1072 samples of rough
+textured ant cuticle textures comprise 43\% of the dataset. The remaining 1427
+samples of smooth textured ant cuticle textures comprise 57\% of the dataset. To
+handle the imbalance of the dataset, we apply undersampling for each class for
+the training dataset. By using random stratified sampling, we construct a
+training set with 800 images per class. The remaining images are randomly split
+between test and validation, which turns out to roughly a 60\%/20\%/20\% train,
+test, and validation data split. With 272 rough samples and 627 smooth samples
+left over after the stratified split, the test dataset has roughly 136 rough
+samples and 313 smooth samples. Since these leftover samples are split with code
+by 50\% there will be some rounding variance and therefore the test dataset
+built at run-time will not always have exactly the same number of samples.
 
@@
 
-\section{The analysis with visualization}
-
+\FloatBarrier
 The results in the previous sections show that the fine-tuned ResNet models
@@
 
+\FloatBarrier
 Based on the visual results of the t-SNE visualization, we can see that the
@@
 
+\FloatBarrier
 \subsection{GradCAM visualization}

### 2022-07-08 paper/main.tex
add kviews parameters
address comments first

@@
 
-\title{Ant cuticle image classification using texture analysis: a comparative study}
+\title{Ant cuticle image classification using texture analysis: a comparative
+    study}
 
@@
 \begin{abstract}
-    There is a large variety of ant species, and most species are diverse in
-    terms of size, shape, behaviors, and especially cuticle textures. However,
-    the significance of ant cuticle texture is not widely researched. Ant
-    cuticle texture presumably provides some type of function, and therefore is
-    useful to research for ecological applications and bioinspired designs. This
-    research employs image texture analysis and deep machine learning to
-    automatically group similar ant species based on morphological traits. We
-    provide a comparative study of the performance of general image
-    classification and texture analysis methods on ant head images. We evaluate
-    the results of the classification methods with modern visualization
-    techniques. Our results show that deep learning methods are applicable for
-    classifying ant cuticle textures. We also show that the performance of the
-    deep texture analysis methods performs better than the deep classification
-    methods.
+    Ant cuticle texture presumably provides some type of function, and therefore
+    is useful to research for ecological applications and bioinspired designs.
+    In this study, we employ statistical image texture analysis and deep machine
+    learning methods to automatically group similar ant species based on
+    morphological features. We establish a public database of ant cuticle images
+    for research. We provide a comparative study of the performance of image
+    texture classification and deep machine learning methods on ant cuticle
+    images. Our results show that the deep learning methods are giving higher
+    accuracy in recognizing ant cuticle textures. Our experiments also reveal
+    that the deep learning networks designed for image texture performs better
+    than the general deep learning networks.
 \end{abstract}
@@
 desiccation, and serves as a canvas for advertising visual and chemical signals
-\cite{gullan_insects_2009}. Research has heavily focused on the macrostructures
-and internal chemical components that make the exoskeleton functional and more
-recent work is being done to understand the functional aspects of external
-cuticle micro sculpturing \cite{muthukrishnan_insect_2020,
-    gunderson_insect_1989, watson_diversity_2017}.
+\cite{gullan_insects_2009}. Research has focused on the macrostructures and
+internal chemical components that describe the functionality of the exoskeleton
+\cite{gunderson_insect_1989}. More recent work is being done to understand the
+functional aspects of external cuticle micro sculpturing
+\cite{muthukrishnan_insect_2020, watson_diversity_2017}.
 
 Due to the extensive number of insect species, manual exploration of
-insect-based information is difficult and often requires specialized expertise.
-Therefore, automated entomology is gaining attraction by both biologists and
-computer scientists and is expected to be a major contribution to the future of
+insect-based information is difficult and often requires expert training.
+Therefore, automated entomology has attracted biologists and computer
+scientists, and is expected to be a major contribution to the future of
 insect-based research \cite{martineau_survey_2017}. One of the most commonly
@@
 
-Our primary contribution is the custom dataset used to explore the relationship
-between ant cuticle texture and automated texture analysis methods. We compare
-classical and modern image classification methods and texture analysis methods
-by evaluating their performance on our dataset.
+There is a large variety of ant species and most species are diverse in terms of
+size, shape, behaviors, and cuticle textures \colorbox{green}{[references]}.
+Based on our initial observation of the cuticle sculpturing in ants, it appears
+that some type of textural structures can be identified in images. Therefore,
+this image dataset may be worth to be explored from the image texture analysis
+point of view for the further knowledge discovery. Image texture analysis has
+been used in image processing for many decades \colorbox{green}{[references]}. Due to its
+diversity of images on scales, smoothness, coarseness, microtexture, and
+macrotexture occurring in images, it is a difficult problem for image textural
+interpreters. With a rapid development of deep machine learning, deep neural
+networks have been frequently used in image texture analysis \colorbox{green}{[references]}.
 
-\section{Related work and literature review}
+Our contribution is to establish an ant image dataset which can be used to
+explore the categorization of different ant cuticle textures using image texture
+analysis methods. \colorbox{green}{(comment: we may need to say how the dataset
+    is established)} Our experiment shows that the deep learning networks with image
+texture embedded in the design give a better accuracy than the K-views
+clustering method.
 
-In this section we will review the literature for ant sculpturing identification
-and texture analysis. Additionally, we discuss the work related to automated
-insect identification.
+This paper is organized as below:  Section 2 reviews the sculpturing
+identification in ants, and image texture analysis and insect classification.
+Section 3 describes the dataset preparation and image texture analysis methods
+used in our experiments. Section 4 gives the experimental results, visualization
+of both raw and processed images and analysis. The conclusion then follows.
 
-\subsection{Sculpturing identification}
+\section{Related work on sculpturing identification in entomology, image texture
+  analysis, and insect classification}
+
+% \subsection{Sculpturing identification}
+
 Taxonomists have developed extensive terminology describing ant cuticle
@@
 are similar to the terms used to describe textural patterns in texture analysis.
-In the next subsection, we review some literature of image texture analysis.
 
-\subsection{Texture analysis}
+% \subsection{Texture analysis}
 
@@
 Scientists are also studying their subjects depending on taking the images
-remotely such as remotely sensed satellite images. With an abundant number of
-textural images surrounding us, it is essential to have a reliable automatic
-tool for the textural image interpretation.
+remotely such as satellite images. With an abundant number of textural images
+surrounding us, it is essential to have a reliable automatic tool for the
+textural image interpretation.
 
-Classical computer methods are developed for textural images based on the
-spatial relationship of gray levels of pixels encoded in the texture
+Computer methods are developed for textural images based on the spatial
+relationship of gray levels of pixels encoded in the texture
 \cite{haralick_image_1985,hung_image_2019}. Those methods by de-facto follow the
@@
 from the pattern recognition principle.  In doing so, an image texture is
-represented as feature vectors and then fed into the algorithm. Markov random
-field models are also studied for textural image interpretation
-\cite{hassner_use_1981, cross_markov_1983}. The transform-based methods use some functions to
-decompose an image texture into a set of basic feature images. Gabor filters and
-Wavelet expansions are two of the widely used approaches
-\cite{bovik_multichannel_1990}.
+represented as feature vector using feature extraction methods and then fed into
+the algorithm. Markov random field models are also studied for textural image
+interpretation \cite{hassner_use_1981, cross_markov_1983}. The transform-based
+methods use some functions to decompose an image texture into a set of basic
+feature images. Gabor filters and Wavelet expansions are two of the widely used
+approaches \cite{bovik_multichannel_1990}.
 
@@
 layers. All these architectures customize the standard CNN structure to
-accomplish the characterization of certain spatial, visual, or statistical
-nature unique to textural images.
+accomplish the characterization of certain spatial, visual, and statistical
+traits unique to textural images.
 
@@
 
-\subsection{Insect classification}
-In this section, we provide an overview of some insect classification methods.
+% \subsection{Insect classification}
+
 Proposed insect classification methods seek to classify insects at different
@@
 ImageNet dataset, there are some categories that specify the class of the insect
-on a species level, \textit{e.g.} \textit{monarch butterfly} and \textit{ringlet
+on a species level, e.g. \textit{monarch butterfly} and \textit{ringlet
     butterfly} as well as some categories that specify the class of the insect on a
-family level, \textit{e.g.} \textit{ant}, \textit{fly}, and \textit{bee}
+family level, e.g. \textit{ant}, \textit{fly}, and \textit{bee}
 \cite{deng_imagenet_2009}. Lim et al. \cite{lim_performance_2017} use a modified
@@
 
-\section{Methodology}
+\section{Ant cuticle image dataset}
 
-In this section, we provide an overview of the methodology used in this paper.
-We describe the steps used to prepare the custom dataset. Finally, we describe
-the algorithms applied and experimental setup to obtain the results.
-
-\subsection{Dataset preparation} % preparation generation
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.7\textwidth]{thesis_assets/images/rough_collage.png}
-    \caption{Examples of rough cuticle texture ant images in the dataset after
-        center cropping, from AntWeb \cite{perrichot_antweb_2012}.}
-    \label{fig:rough-cuticle-texture}
-\end{figure}
+% \subsection{Dataset preparation} % preparation generation
 
@@
 randomly selected images from each category are shown in Figures
-\ref{fig:rough-cuticle-texture} and \ref{fig:smooth-cuticle-texture}.
+\ref{fig:smooth-cuticle-texture} and \ref{fig:rough-cuticle-texture}.
+
+To begin, a master spreadsheet was created with the 2,499 different ant species
+to be identified for the primary dataset. The team was trained to identify
+cuticle sculpturing through a process which consisted of one 45-minute
+introductory lesson explaining the project and texture categories. Then, the
+team was given a training set of photos to identify from the genus
+\textit{Polyrhachis}. The sculpture identification protocol describes the two
+primary categories: rough and smooth.
 
@@
 
-To begin, a master spreadsheet was created with the 2,499 different ant species
-to be identified for the primary dataset. The team was trained to identify
-cuticle sculpturing through a process which consisted of one 45-minute
-introductory lesson explaining the project and texture categories. Then, the
-team was given a training set of photos to identify from the genus
-\textit{Polyrhachis}. The sculpture identification protocol describes the two
-primary categories: \textit{rough} and \textit{smooth}.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.7\textwidth]{thesis_assets/images/rough_collage.png}
+    \caption{Examples of rough cuticle texture ant images in the dataset after
+        center cropping, from AntWeb \cite{perrichot_antweb_2012}.}
+    \label{fig:rough-cuticle-texture}
+\end{figure}
 
@@
 To collect the images, the assistants followed the taxonomy information
-available in the master spreadsheet to the appropriate AntWeb page
-\cite{perrichot_antweb_2012}. In many cases, there are multiple ant head images
-of the same species, and occasionally there are multiple image resolution
-available from a single image. To simplify the data collection process, the
-assistants were instructed to download the first ant head image of the species
-being identified in the highest resolution possible. Each image was named with
-an identifier that corresponds with the row number in the master spreadsheet.
-The same ant head images that were downloaded in the data collection phase were
-the same ones used in the sculpture identification protocol. Ant species which
-did not have any images of the head were excluded from the dataset.
-Additionally, ant species which only had a head image of a queen ant were
-excluded from the dataset.
+available in the master spreadsheet to the appropriate
+AntWeb\footnote{\url{https://www.antweb.org/}} page. In many cases, there are
+multiple ant head images of the same species, and occasionally there are
+multiple image resolution available from a single image. To simplify the data
+collection process, the assistants were instructed to download the first ant
+head image of the species being identified in the highest resolution possible.
+Each image was named with an identifier that corresponds with the row number in
+the master spreadsheet. The same ant head images that were downloaded in the
+data collection phase were the same ones used in the sculpture identification
+protocol. Ant species which did not have any images of the head were excluded
+from the dataset. Additionally, ant species which only had a head image of a
+queen ant were excluded from the dataset.
 
@@
 
-Due to the variety of the ant head image attributes, we apply simple
-preprocessing before the images are used in our model. We want the images to
-have a uniform size for simplicity in our classification process. Since the ant
-head images are typically centered in the image, we apply a center crop to each
-image to create a square image of the same size. Once the image is square, we
-resize each image to a fixed size of 256x256 pixels. We leave other
-discrepancies in the images untouched.
+The completed dataset of contains 2,499 images. 1072 samples of rough textured
+ant cuticle textures comprise 43\% of the dataset. The remaining 1427 samples of
+smooth textured ant cuticle textures comprise 57\% of the dataset. Due to the
+variety of the ant head image attributes, we apply simple preprocessing before
+the images are used in our model. We want the images to have a uniform size for
+simplicity in our classification process. Since the ant head images are
+typically centered in the image, we apply a center crop to each image to create
+a square image of the same size. Once the image is square, we resize each image
+to a fixed size of 256x256 pixels. We leave other discrepancies in the images
+untouched.
 
-\subsection{Classical texture analysis methods}
+\section{Methods for comparative analysis and experimental setup}
+
+% \subsection{Classical texture analysis methods}
 
@@
 useful in feeding the datasets to the K-views algorithms. A 10-channel Gabor
-filter is applied in the pre-processing step for the K-means algorithm.
+filter is applied in the pre-processing step for the K-means algorithm. The
+theta $\theta$ parameters in the Gabor filters used are multiples of $\pi/10$
+radians starting from $0$ to $\pi$ radians. The Gabor filter lambda $\lambda$
+parameter is equal to $\pi$.
 
-\subsection{Deep learning models used for the experiments}
+% \subsection{Deep learning models used for the experiments}
 
-Our first model is \textit{visual geometry group} (VGG), a convolutional neural
-network that takes advantage of very small convolutional filters in a deep
-network architecture \cite{simonyan_very_2015}. We compare four architectures of
-VGG: VGG11, VGG13, VGG16, and VGG19. The primary difference between the
-architectures is the number of layers in each model. Our second model is
-\textit{residual network} (ResNet), a deep network architecture that includes
-shortcut connections between layers (residual connections) \cite{he_deep_2015}.
-We compare three architectures of ResNet: ResNet18, ResNet50, and ResNet101.
-Again, the primary difference between the architectures is the number of layers
-in each model.
+Our first deep learning model is \textit{visual geometry group} (VGG), a
+convolutional neural network that takes advantage of very small convolutional
+filters in a deep network architecture \cite{simonyan_very_2015}. We compare
+four architectures of VGG: VGG11, VGG13, VGG16, and VGG19. The primary
+difference between the architectures is the number of layers in each model. Our
+second model is \textit{residual network} (ResNet), a deep network architecture
+that includes shortcut connections between layers (residual connections)
+\cite{he_deep_2015}. We compare three architectures of ResNet: ResNet18,
+ResNet50, and ResNet101. Again, the primary difference between the architectures
+is the number of layers in each model.
 
-\subsection{Deep texture analysis method used for the experiments}
+For our ResNet models, we have two versions: randomized and pretrained. The
+randomized version is the same architecture, but the weights are randomly
+initialized. The pretrained version has weights from training on the CIFAR
+dataset, an image dataset with 1000 classes. In this case, we are fine-tuning
+the pretrained model. For VGG, we are only using the randomized version. The
+base VGG architecture also has an output layer of size 1000. Since we are
+working with a binary classification problem, we modify the architecture for all
+models to have an output layer of size 2. Each model is trained over 100 epochs,
+using stochastic gradient descent with momentum. The batch size is set to 16
+images. We apply a learning rate of 0.001 and momentum parameter of 0.9.
+
+% \subsection{Deep texture analysis method used for the experiments}
 
@@
 
-\section{Results and analysis}
-\subsection{Environment}
+To handle the imbalance of the dataset, we apply undersampling for each class for
+the training dataset. By using random stratified sampling, we construct a
+training set with 800 images per class. The remaining images are randomly split
+between test and validation, which turns out to roughly a 60\%/20\%/20\% train,
+test, and validation data split. With 272 rough samples and 627 smooth samples
+left over after the stratified split, the test dataset has roughly 136 rough
+samples and 313 smooth samples. Since these leftover samples are split with code
+by 50\% there will be some rounding variance and therefore the test dataset
+built at run-time will not always have exactly the same number of samples.
 
-Experiments are run on an Ubuntu 18.04 LTS Lambda Labs GPU server. The server
-contains 8 NVIDIA GeForce RTX 2080 Ti graphics cards with 12GB of memory each.
-The server uses an Intel Xeon Silver 4116 with 48 total threads and maximum
-frequency of 3.000 GHz, and has 256GB of RAM.
-
-\subsection{Evaluation}
+% \subsection{Evaluation}
 
@@
 
-\subsection{Experimental results}
+% \subsection{Environment}
 
-For our ResNet models, we have two versions: randomized and pretrained. The
-randomized version is the same architecture, but the weights are randomly
-initialized. The pretrained version has weights from training on the CIFAR
-dataset, an image dataset with 1000 classes. In this case, we are fine-tuning
-the pretrained model. For VGG, we are only using the randomized version. The
-base VGG architecture also has an output layer of size 1000. Since we are
-working with a binary classification problem, we modify the architecture for all
-models to have an output layer of size 2. Each model is trained over 100 epochs,
-using stochastic gradient descent with momentum. The batch size is set to 16
-images. We apply a learning rate of 0.001 and momentum parameter of 0.9.
+Experiments are run on an Ubuntu 18.04 LTS Lambda Labs GPU server. The server
+contains 8 NVIDIA GeForce RTX 2080 Ti graphics cards with 12GB of memory each.
+The server uses an Intel Xeon Silver 4116 with 48 total threads and maximum
+frequency of 3.000 GHz, and has 256GB of RAM.
 
-For the experiments using the deep learning algorithms, we use our custom
-dataset of ant head images which contains 2,499 images. 1072 samples of rough
-textured ant cuticle textures comprise 43\% of the dataset. The remaining 1427
-samples of smooth textured ant cuticle textures comprise 57\% of the dataset. To
-handle the imbalance of the dataset, we apply undersampling for each class for
-the training dataset. By using random stratified sampling, we construct a
-training set with 800 images per class. The remaining images are randomly split
-between test and validation, which turns out to roughly a 60\%/20\%/20\% train,
-test, and validation data split. With 272 rough samples and 627 smooth samples
-left over after the stratified split, the test dataset has roughly 136 rough
-samples and 313 smooth samples. Since these leftover samples are split with code
-by 50\% there will be some rounding variance and therefore the test dataset
-built at run-time will not always have exactly the same number of samples.
+\section{Results and analysis}
+
+% \subsection{Experimental results}
 
@@
 imbalance in the dataset, the F1 score is the preferable metric to the accuracy.
-Therefore, the results have been sorted by F1 score in descending order.
+% Therefore, the results have been sorted by F1 score in descending order.
 
@@
 
-\subsection{t-SNE Visualization}
+% \subsection{Clustering segmentation}
+
+% \subsection{t-SNE Visualization}
 
@@
 initialized model and Figure \ref{fig:fresnet101_tsne} shows the results of the
-fine-tuned model.
+fine-tuned model. Based on the visual results of the t-SNE visualization, we can see that the
+fine-tuned model learned a stronger separation of the two classes, which
+reinforces the results that the fine-tuned model received a higher average
+accuracy.
 
@@
 
-\FloatBarrier
-Based on the visual results of the t-SNE visualization, we can see that the
-fine-tuned model learned a stronger separation of the two classes, which
-reinforces the results that the fine-tuned model received a higher average
-accuracy.
+% \subsection{GradCAM visualization}
 
-\FloatBarrier
-\subsection{GradCAM visualization}
-
-In this section, we provide some visual analysis of some correctly and
+Next, we provide some visual analysis of some correctly and
 incorrectly classified images using GradCAM. We provide two categories and two
@@
 features should be further analyzed. In essence, the model in this situation
-knows \textit{where} to look, but not \textit{what} to look for. In Figure
+knows where to look, but not what to look for. In Figure
 \ref{fig:incorrect_ideal_84}, the features activated are mostly in the correct

### 2022-07-10 paper/main.tex
figures and tables update

--- 2022-06-24 paper/main.tex
@@
 We examine ants (\textit{Formicidae}) as they display an extreme diversity of
-cuticle micro sculpturing across all subfamilies. Sculpturing ranges from
+cuticle micro sculpturing across all subfamilies. Microsculpture ranges from
 parallel longitudinal ridges to deep oval impressions to erratic protuberances.
-The sculpturing has arisen convergently and independently throughout ant's
-evolutionary history, which suggests some inherent function. Cuticle sculpturing
-on ants may help increase strength and rigidness, resist abrasion, increase
-internal and external surface area, resist microbial growth, and rear beneficial
-anti-biotic producing bacteria \cite{johnson_effect_2011,
-    bruckner_relationship_2017, currie_coevolved_2006}. These specific functions may
-be associated with certain sculpturing types. In order to analyze those
-functions, it is necessary to segment the image and group similar textures for
-further analysis.
+Microsculpture has arisen convergently and independently throughout ant's
+evolutionary history, which suggests  that these are complex traits undergoing
+selection. Cuticle microsculpture on ants may help increase strength and
+rigidness, resist abrasion, increase internal and external surface area, resist
+microbial growth, and rear beneficial anti-biotic producing bacteria
+\cite{johnson_effect_2011, bruckner_relationship_2017, currie_coevolved_2006}.
+These specific functions may be associated with certain sculpturing types. In
+order to analyze those functions, it is necessary to segment the image and group
+similar textures for further analysis.
 
@@
 
-\subsection{Sculpturing identification}
+\subsection{Microsculpture identification}
 Taxonomists have developed extensive terminology describing ant cuticle
-sculpturing as it is often a useful diagnostic trait to distinguish between
+microsculpture as it is often a useful diagnostic trait to distinguish between
 closely related ant species \cite{blaimer_taxonomy_2019,fisher_ants_2007}. The
@@
     Sculpturing} — contains over 100 terms to describe the cuticle sculpturing
-patterns of ants \cite{harris_glossary_1979}.
+patterns of ants \cite{harris_glossary_1979}. Despite the numerous ways to
+describe microsculpture, in general, microsculpture diversity falls into two
+classes - smooth and rough. Smooth ants are free of any obvious microsculpture
+and may have small hairs that give the cuticle a dull appearance. Rough ants
+have a variety of raised lines or depressed textures on the cuticle. The
+entirety of ant microsculpture is represented within these two classes.
 
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{thesis_assets/images/rough_collage.png}
+    \includegraphics[width=0.7\textwidth]{assets/images/rough_full_collage.png}
     \caption{Examples of rough cuticle texture ant images in the dataset after
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{thesis_assets/images/smooth_collage.png}
+    \includegraphics[width=0.7\textwidth]{assets/images/smooth_full_collage.png}
     \caption{Examples of smooth cuticle texture ant images in the dataset after
@@
 
-Initially, the sculpture identification protocol had 8 subcategories of cuticle
-texture, including dimpled, ridged, and differing levels of smooth texture. For
-simplicity, we work only with the two main categories. The training set
-identifications were reviewed together as a group by the assistants. Once
-training was complete, assistants were assigned the same genera of ants to
-identify independently each week. A weekly meeting was held to discuss
-identifications and assign new ones. These identifications were collected in the
-master spreadsheet and the identifications were assigned to individual ant
-species on a majority basis.
+Initially, the sculpture identification protocol had 5 subcategories of cuticle
+texture: smooth, punctate, striate, reticulate, and tuberous. For simplicity, we
+work only with the two main categories. The training set identifications were
+reviewed together as a group by the assistants. Once training was complete,
+assistants were assigned the same genera of ants to identify independently each
+week. A weekly meeting was held to discuss identifications and assign new ones.
+These identifications were collected in the master spreadsheet and the
+identifications were assigned to individual ant species on a majority basis.
 
@@
 resize each image to a fixed size of 256x256 pixels. We leave other
-discrepancies in the images untouched.
+discrepancies in the images untouched. A summary of the resulting dataset is
+shared in Table \ref{tab:dataset-summary}.
+
+\begin{table}[h]
+    \centering
+    \caption{Summary of the dataset.}
+    \begin{tabular}{lc}
+        \toprule
+        Total images  & 2499    \\
+        Rough images  & 1072    \\
+        Smooth images & 1427    \\
+        Image size    & 256x256 \\
+    \end{tabular}
+    \label{tab:dataset-summary}
+\end{table}
 
@@
 useful in feeding the datasets to the K-views algorithms. A 10-channel Gabor
-filter is applied in the pre-processing step for the K-means algorithm.
+filter is applied in the pre-processing step for the K-means algorithm. The
+theta $\theta$ parameters in the Gabor filters used are multiples of $\pi/10$
+radians starting from $0$ to $\pi$ radians. The Gabor filter lambda $\lambda$
+parameter is equal to $\pi$.
 
@@
     \centering
-    \caption{Results}
+    \caption{Experimental results of different algorithms}
     \input{assets/results.tex}
@@
 %     visualization.)}
+
+% \subsection{Clustering segmentation}
 
@@
 the dataset and the fine-tuned ResNet101 model to perform the analysis. We show
-the GradCAM results in Figures \ref{fig:correct_ideal},
-\ref{fig:correct_nonideal}, \ref{fig:incorrect_ideal}, and
-\ref{fig:incorrect_nonideal}. The left image shows the preprocessed image input
+the GradCAM results in Figures \ref{fig:correct_images} and
+\ref{fig:incorrect_images}. The left image shows the preprocessed image input
 to the model. The right image shows the GradCAM output based on the
@@
 knows \textit{where} to look, but not \textit{what} to look for. In Figure
-\ref{fig:incorrect_ideal_84}, the features activated are mostly in the correct
-location on the ant head, and the rough texture is clearly visible, yet the
-model predicts the incorrect class \textit{smooth}. Similarly in Figure
-\ref{fig:incorrect_ideal_177}, the features activated are also mostly in the
-correct location, yet the model predicts the incorrect class \textit{smooth}. In
-this case, it may be due to the pose of the ant being slightly different from
-the average pose. In the incorrectly classified images with non-expected
-features, analysis shows that the model is unable to find \textit{where} to
-look, and obtains feature information from other parts of the ant or the
-background. Cases where the image was correctly classified using the
-non-expected features can basically be seen as noise. In order to further
-analyze this class, we should introduce some parameter such as model confidence
-to examine further.
+\ref{fig:incorrect_images}, there are results where the features activated are
+mostly in the correct location on the ant head, and the rough texture is clearly
+visible, yet the model predicts the incorrect class \textit{smooth}. Similarly
+in Figure \ref{fig:incorrect_images}, there are results where the features
+activated are also mostly in the correct location, yet the model predicts the
+incorrect class \textit{smooth}. In this case, it may be due to the pose of the
+ant being slightly different from the average pose. In the incorrectly
+classified images with non-expected features, analysis shows that the model is
+unable to find \textit{where} to look, and obtains feature information from
+other parts of the ant or the background. Cases where the image was correctly
+classified using the non-expected features can basically be seen as noise. In
+order to further analyze this class, we should introduce some parameter such as
+model confidence to examine further.
 
 \newcommand{\subwidth}{0.35\textwidth}
+% \begin{figure}
+%     \centering
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
+%         \caption*{Smooth texture}
+%         \label{fig:correct_ideal_518}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
+%         \caption*{Smooth texture}
+%         \label{fig:correct_ideal_808}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_ideal_842}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_ideal_1091}
+%     \end{subfigure}
+%     \caption{Correctly classified images using expected features.}
+%     \label{fig:correct_ideal}
+% \end{figure}
+
+% \begin{figure}
+%     \centering
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_nonideal_346}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_nonideal_1554}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_nonideal_1694}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
+%         \caption*{Smooth texture}
+%         \label{fig:correct_nonideal_388}
+%     \end{subfigure}
+%     \caption{Correctly classified images using non-expected features.}
+%     \label{fig:correct_nonideal}
+% \end{figure}
+
 \begin{figure}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
-        \caption{Ground truth: smooth}
+        \caption*{Smooth texture}
         \label{fig:correct_ideal_518}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
-        \caption{Ground truth: smooth}
+        \caption*{Smooth texture}
         \label{fig:correct_ideal_808}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:correct_ideal_842}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:correct_ideal_1091}
     \end{subfigure}
-    \caption{Correctly classified images using expected features.}
-    \label{fig:correct_ideal}
-\end{figure}
-
-\begin{figure}
-    \centering
     \begin{subfigure}{\subwidth}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:correct_nonideal_346}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:correct_nonideal_1554}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:correct_nonideal_1694}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
-        \caption{Ground truth: smooth}
+        \caption*{Smooth texture}
         \label{fig:correct_nonideal_388}
     \end{subfigure}
-    \caption{Correctly classified images using non-expected features.}
-    \label{fig:correct_nonideal}
+    \caption{Correctly classified images.}
+    \label{fig:correct_images}
 \end{figure}
+
+% \begin{figure}
+%     \centering
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/84.png}
+%         \caption*{Rough texture}
+%         \label{fig:incorrect_ideal_84}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/86.png}
+%         \caption*{Smooth texture}
+%         \label{fig:incorrect_ideal_86}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/177.png}
+%         \caption*{Smooth texture}
+%         \label{fig:incorrect_ideal_177}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/197.png}
+%         \caption*{Smooth texture}
+%         \label{fig:incorrect_ideal_197}
+%     \end{subfigure}
+%     \caption{Incorrectly classified images using expected features.}
+%     \label{fig:incorrect_ideal}
+% \end{figure}
+
+% \begin{figure}
+%     \centering
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/1.png}
+%         \caption*{Rough texture}
+%         \label{fig:incorrect_nonideal_1}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/22.png}
+%         \caption*{Rough texture}
+%         \label{fig:incorrect_nonideal_22}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/61.png}
+%         \caption*{Rough texture}
+%         \label{fig:incorrect_nonideal_61}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/204.png}
+%         \caption*{Rough texture}
+%         \label{fig:incorrect_nonideal_204}
+%     \end{subfigure}
+%     \caption{Incorrectly classified images using non-expected features.}
+%     \label{fig:incorrect_nonideal}
+% \end{figure}
 
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/84.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:incorrect_ideal_84}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/86.png}
-        \caption{Ground truth: smooth}
+        \caption*{Smooth texture}
         \label{fig:incorrect_ideal_86}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/177.png}
-        \caption{Ground truth: smooth}
+        \caption*{Smooth texture}
         \label{fig:incorrect_ideal_177}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/197.png}
-        \caption{Ground truth: smooth}
+        \caption*{Smooth texture}
         \label{fig:incorrect_ideal_197}
     \end{subfigure}
-    \caption{Incorrectly classified images using expected features.}
-    \label{fig:incorrect_ideal}
-\end{figure}
-
-\begin{figure}
-    \centering
     \begin{subfigure}{\subwidth}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/1.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:incorrect_nonideal_1}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/22.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:incorrect_nonideal_22}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/61.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:incorrect_nonideal_61}
@@
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/204.png}
-        \caption{Ground truth: rough}
+        \caption*{Rough texture}
         \label{fig:incorrect_nonideal_204}
     \end{subfigure}
-    \caption{Incorrectly classified images using non-expected features.}
-    \label{fig:incorrect_nonideal}
+    \caption{Incorrectly classified images.}
+    \label{fig:incorrect_images}
 \end{figure}

### 2022-07-10 cuticulus/core/collages.py

@@ -42,11 +42,22 @@ def build_collage(
             replace=False,
         )
 
-        # create a collage of images
-        collage = Image.new('RGB', (size[1] * cols, size[0] * rows))
+        # create a collage of images, with padding between images
+        padding = 32
+        col_size = (size[1] + padding) * cols - padding
+        row_size = (size[0] + padding) * rows - padding
+        collage = Image.new('RGB', (col_size, row_size))
+
+        # set background color to white
+        collage.paste((255, 255, 255), (0, 0, col_size, row_size))
+
+        # paste images into collage
         for i, idx in enumerate(idx):
             image = Image.fromarray(images[idx])
-            collage.paste(image, (size[1] * (i % cols), size[0] * (i // cols)))
+            collage.paste(image, (
+                (i % cols) * (size[1] + padding),
+                (i // cols) * (size[0] + padding),
+            ))
 
         # scale down the collage
         collage = collage.resize((size[1] * cols // 2, size[0] * rows // 2))

### 2022-07-10 paper/main.tex

--- 2022-07-08 paper/main.tex
@@
 Taxonomists have developed extensive terminology describing ant cuticle
-sculpturing as it is often a useful diagnostic trait to distinguish between
+microsculpture as it is often a useful diagnostic trait to distinguish between
 closely related ant species \cite{blaimer_taxonomy_2019,fisher_ants_2007}. The
@@
 to a fixed size of 256x256 pixels. We leave other discrepancies in the images
-untouched.
+untouched. A summary of the resulting dataset is shared in Table
+\ref{tab:dataset-summary}.
+
+\begin{table}[h]
+    \centering
+    \caption{Summary of the dataset.}
+    \begin{tabular}{lc}
+        \toprule
+        Total images  & 2499    \\
+        Rough images  & 1072    \\
+        Smooth images & 1427    \\
+        Image size    & 256x256 \\
+    \end{tabular}
+    \label{tab:dataset-summary}
+\end{table}
 
@@
 features should be further analyzed. In essence, the model in this situation
-knows where to look, but not what to look for. In Figure
-\ref{fig:incorrect_ideal_84}, the features activated are mostly in the correct
-location on the ant head, and the rough texture is clearly visible, yet the
-model predicts the incorrect class \textit{smooth}. Similarly in Figure
-\ref{fig:incorrect_ideal_177}, the features activated are also mostly in the
-correct location, yet the model predicts the incorrect class \textit{smooth}. In
-this case, it may be due to the pose of the ant being slightly different from
-the average pose. In the incorrectly classified images with non-expected
-features, analysis shows that the model is unable to find \textit{where} to
-look, and obtains feature information from other parts of the ant or the
-background. Cases where the image was correctly classified using the
-non-expected features can basically be seen as noise. In order to further
-analyze this class, we should introduce some parameter such as model confidence
-to examine further.
+knows \textit{where} to look, but not \textit{what} to look for. In Figure
+\ref{fig:incorrect_images}, there are results where the features activated are
+mostly in the correct location on the ant head, and the rough texture is clearly
+visible, yet the model predicts the incorrect class \textit{smooth}. Similarly
+in Figure \ref{fig:incorrect_images}, there are results where the features
+activated are also mostly in the correct location, yet the model predicts the
+incorrect class \textit{smooth}. In this case, it may be due to the pose of the
+ant being slightly different from the average pose. In the incorrectly
+classified images with non-expected features, analysis shows that the model is
+unable to find \textit{where} to look, and obtains feature information from
+other parts of the ant or the background. Cases where the image was correctly
+classified using the non-expected features can basically be seen as noise. In
+order to further analyze this class, we should introduce some parameter such as
+model confidence to examine further.
 
 \newcommand{\subwidth}{0.35\textwidth}
+% \begin{figure}
+%     \centering
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
+%         \caption*{Smooth texture}
+%         \label{fig:correct_ideal_518}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
+%         \caption*{Smooth texture}
+%         \label{fig:correct_ideal_808}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_ideal_842}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_ideal_1091}
+%     \end{subfigure}
+%     \caption{Correctly classified images using expected features.}
+%     \label{fig:correct_ideal}
+% \end{figure}
+
+% \begin{figure}
+%     \centering
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_nonideal_346}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_nonideal_1554}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
+%         \caption*{Rough texture}
+%         \label{fig:correct_nonideal_1694}
+%     \end{subfigure}
+%     \begin{subfigure}{\subwidth}
+%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
+%         \caption*{Smooth texture}
+%         \label{fig:correct_nonideal_388}
+%     \end{subfigure}
+%     \caption{Correctly classified images using non-expected features.}
+%     \label{fig:correct_nonideal}
+% \end{figure}
+
 \begin{figure}

### 2022-07-22 paper/main.tex
hung changes

@@
 
-\title{Ant cuticle image classification using texture analysis: a comparative
-    study}
+\title{Investigation of ant cuticle dataset using image texture
+    analysis methods}
 
@@
 There is a large variety of ant species and most species are diverse in terms of
-size, shape, behaviors, and cuticle textures \colorbox{green}{[references]}.
+size, shape, behaviors, and cuticle textures \cite{harris_glossary_1979}.
 Based on our initial observation of the cuticle sculpturing in ants, it appears
@@
 point of view for the further knowledge discovery. Image texture analysis has
-been used in image processing for many decades \colorbox{green}{[references]}. Due to its
+been used in image processing for many decades \cite{hung_image_2019}. Due to its
 diversity of images on scales, smoothness, coarseness, microtexture, and
@@
 interpreters. With a rapid development of deep machine learning, deep neural
-networks have been frequently used in image texture analysis \colorbox{green}{[references]}.
+networks have been frequently used in image texture analysis
+\cite{sun_robust_2014,liu_bow_2019,liu_pestnet_2019}.
 
@@
 explore the categorization of different ant cuticle textures using image texture
-analysis methods. \colorbox{green}{(comment: we may need to say how the dataset
-    is established)} Our experiment shows that the deep learning networks with image
-texture embedded in the design give a better accuracy than the K-views
-clustering method.
+analysis methods. We share how the dataset of ant cuticle images is established
+in Section 3. Our experiment shows that the deep learning networks designed for
+image texture analysis give a better accuracy than the K-views clustering
+method.
 
@@
 
-Taxonomists have developed extensive terminology describing ant cuticle
-microsculpture as it is often a useful diagnostic trait to distinguish between
-closely related ant species \cite{blaimer_taxonomy_2019,fisher_ants_2007}. The
-definitive text on ant cuticle terminology — \textit{The Glossary of Surface
-    Sculpturing} — contains over 100 terms to describe the cuticle sculpturing
-patterns of ants \cite{harris_glossary_1979}. Despite the numerous ways to
-describe microsculpture, in general, microsculpture diversity falls into two
-classes - smooth and rough. Smooth ants are free of any obvious microsculpture
-and may have small hairs that give the cuticle a dull appearance. Rough ants
-have a variety of raised lines or depressed textures on the cuticle. The
-entirety of ant microsculpture is represented within these two classes.
+Because ants lack bold coloration and complex wing venation patterns used to
+identify other insect species, microsculpturing has taken a prominent role in
+ant identification. To aid in species identification, ant taxonomists have
+developed nearly 100 terms to describe  fine differences in microsculpturing
+patterns \cite{harris_glossary_1979}. Despite the prominent role of
+microsculpturing in ant identification, however, almost nothing is known about
+how these patterns evolved or their function. Similar microsculpturing patterns
+in other organisms have been found to reduce abrasion, increase stiffness, and
+perform as structural antibiotics \cite{zhiwu_erosion_2012, han_efficient_2015,
+    rees_form_1975, chung_impact_2007,hasan_selective_2013}. Developing a better
+understanding of the function of microsculpturing in ants could provide insight
+into their evolution and biology as well as lead to the development of
+bio-inspired technologies \cite{penick_comparative_2022}.
 
-Cuticle sculpturing in ants has been explored thoroughly from a taxonomic
-perspective; however, the function of nano and microstructures on insect
-exoskeletons is a developing topic in entomology. Watson et al. review the
-literature of cuticle nano and microstructure function and propose 21 possible
-functions associated with these structures \cite{watson_diversity_2017}. Many of
-these functions are related to structures not found in ants such as scales and
-nanostructures. The review does include functions that may relate to ants such
-as friction control, enhanced surface area, and increased hardness. Watson et
-al. also describe seven types of cuticle structures ranging from hairs and
-scales to nano and micro structures \cite{watson_diversity_2017}. The cuticle
-sculpturing of ants seems to fall within one type - complex microstructures.
+One challenge for functional investigations of cuticle microsculpturing is the
+complexity of terms used to describe these patterns. Complex terminology may be
+useful for describing fine differences between closely related species, but
+broader comparisons require a simplified classification scheme. We recently
+developed a classification system for microsculpturing that consists of only
+five categories: smooth, striate, punctate, reticulate, and tuberous. Using
+these categories to classify 11,722 ant species, we found that the earliest ants
+were likely smooth and that cuticle microsculpturing independently evolved later
+in multiple lineages. An even simpler classification scheme is to divide ants
+into only two categories: smooth and rough. Because microsculptuing has evolved
+numerous times in ants, this binary classification scheme is sufficient to
+detect conditions that select for any sculpturing at all compared to a smooth
+state.
 
-The five broad functional groupings developed that describe the complex
-microstructures found on ants were derived from reviewing the variation of
-cuticle sculpturing across ants. These functional groupings were also reflected
-in Harris as many terms could be grouped together based on similar definitions
-and comparing the scanning electron microscope photographs provided in the
-publication \cite{harris_glossary_1979}.
-
-There is often substantial overlap among terms, and closely related cuticle
-patterns are likely to be functionally similar. For example, the definition for
-“imbricate” is, “partly overlapping and appearing like shingles on a roof or
-scales on a fish,” which is difficult to distinguish from the definition of
-tesselate, “made up of squares like a chess board, either in sculpturing or in
-color.” However, we note that the terms used to describe the cuticle patterns
-are similar to the terms used to describe textural patterns in texture analysis.
+Using either a simplified or binary classification scheme, the next challenge is
+figuring out how to apply either of these schemes to large numbers of species.
+We live in an era of big data, and insect researchers have access to millions of
+images available through digital museum collections \cite{beaman_mass_2012}.
+Going through these digital collections manually is not feasible for large
+projects. Focusing only on ants, for example, there are over 14,000 described
+species, and each species has multiple worker and reproductive castes that may
+vary in their microsculpturing patterns. Even within an individual ant,
+microsculpturing typically varies among and within body segments, which
+exponentially increases the number of images that must be assessed. Applied
+computing offers a solution to this problem through developing tools that
+automate trait classification. Automated approaches would allow researchers to
+leverage existing image repositories to extract large trait datasets that can
+then be used to detect evolutionary patterns.
 
@@
     \caption{Summary of the dataset.}
-    \begin{tabular}{lc}
+    \begin{tabular}{|l|c|}
         \toprule
@@
         Image size    & 256x256 \\
+        \bottomrule
     \end{tabular}
@@
     \end{subfigure}
-    \caption{Correctly classified images.}
+    \caption{Correctly classified images from fine-tuned ResNet101.}
     \label{fig:correct_images}
@@
     \end{subfigure}
-    \caption{Incorrectly classified images.}
+    \caption{Incorrectly classified images from fine-tuned ResNet101.}
     \label{fig:incorrect_images}
 \end{figure}
+
+\newcommand{\segmentedsubwidth}{0.24\textwidth}
+\begin{figure}
+    \centering
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r103.png}
+        \caption*{Rough texture}
+    \end{subfigure}
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r111.png}
+        \caption*{Smooth texture}
+    \end{subfigure}
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s105.png}
+        \caption*{Rough texture}
+    \end{subfigure}
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s107.png}
+        \caption*{Smooth texture}
+    \end{subfigure}
+
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/sr103.png}
+    \end{subfigure}
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/sr111.png}
+    \end{subfigure}
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/ss105.png}
+    \end{subfigure}
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/ss107.png}
+    \end{subfigure}
+
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat103.png}
+    \end{subfigure}
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat111.png}
+    \end{subfigure}
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat105.png}
+    \end{subfigure}
+    \begin{subfigure}{\segmentedsubwidth}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat107.png}
+    \end{subfigure}
+    \caption{Raw images are shown with corresponding segmented outputs from
+        K-views algorithm. The second row shows the results of using size of
+        15x15 patches. The third row shows the results of using size of 19x19
+        patches.}
+    \label{fig:kviews_segmented}
+\end{figure}
+
+\FloatBarrier Some classified results from the K-views are shown in Figure
+\ref{fig:kviews_segmented}. A training set of 500 patches for background, 500
+patches for rough, and 500 patches for smooth are applied with patch sizes of
+15x15 and 19x19. These patches are extracted from two rough and two smooth
+images. The K-views algorithm converges early although we set the maximum number
+of iterations to be 1000.
 

### 2022-07-22 paper/main.tex
pre submission

@@
 developed nearly 100 terms to describe  fine differences in microsculpturing
-patterns \cite{harris_glossary_1979}. Despite the prominent role of
-microsculpturing in ant identification, however, almost nothing is known about
-how these patterns evolved or their function. Similar microsculpturing patterns
-in other organisms have been found to reduce abrasion, increase stiffness, and
-perform as structural antibiotics \cite{zhiwu_erosion_2012, han_efficient_2015,
-    rees_form_1975, chung_impact_2007,hasan_selective_2013}. Developing a better
-understanding of the function of microsculpturing in ants could provide insight
-into their evolution and biology as well as lead to the development of
-bio-inspired technologies \cite{penick_comparative_2022}.
+patterns \cite{harris_glossary_1979, bolton_identification_1994}. Despite the
+prominent role of microsculpturing in ant identification, however, almost
+nothing is known about how these patterns evolved or their function. Similar
+microsculpturing patterns in other organisms have been found to reduce abrasion,
+increase stiffness, and perform as structural antibiotics
+\cite{zhiwu_erosion_2012, han_efficient_2015, rees_form_1975,
+    chung_impact_2007,hasan_selective_2013}. Developing a better understanding of
+the function of microsculpturing in ants could provide insight into their
+evolution and biology as well as lead to the development of bio-inspired
+technologies \cite{penick_comparative_2022}.
 
@@
 developed a classification system for microsculpturing that consists of only
-five categories: smooth, striate, punctate, reticulate, and tuberous. Using
-these categories to classify 11,722 ant species, we found that the earliest ants
-were likely smooth and that cuticle microsculpturing independently evolved later
-in multiple lineages. An even simpler classification scheme is to divide ants
-into only two categories: smooth and rough. Because microsculptuing has evolved
-numerous times in ants, this binary classification scheme is sufficient to
-detect conditions that select for any sculpturing at all compared to a smooth
+five categories: smooth, striate, punctate, reticulate, and tuberous
+\cite{hellenbrand_ant_2022}. Using these categories to classify 11,722 ant
+species, we found that the earliest ants were likely smooth and that cuticle
+microsculpturing independently evolved later in multiple lineages
+\cite{hellenbrand_ant_2022}. An even simpler classification scheme is to divide
+ants into only two categories: smooth and rough. Because microsculptuing has
+evolved numerous times in ants, this binary classification scheme is sufficient
+to detect conditions that select for any sculpturing at all compared to a smooth
 state.
@@
 characteristics features such as edge-like in the image. This step is very
-useful in feeding the datasets to the K-views algorithms. A 10-channel Gabor
+useful in feeding the datasets to the K-views algorithms. A 4-channel Gabor
 filter is applied in the pre-processing step for the K-means algorithm. The
-theta $\theta$ parameters in the Gabor filters used are multiples of $\pi/10$
+theta $\theta$ parameters in the Gabor filters used are multiples of $\pi/4$
 radians starting from $0$ to $\pi$ radians. The Gabor filter lambda $\lambda$

### 2022-07-29 paper/main.tex
add subsections, update images

@@
 
-This paper is organized as below:  Section 2 reviews the sculpturing
+This paper is organized as below:  Section 2 reviews the microsculpturing
 identification in ants, and image texture analysis and insect classification.
@@
 
-\section{Related work on sculpturing identification in entomology, image texture
-  analysis, and insect classification}
+\section{Related work}
 
-% \subsection{Sculpturing identification}
+In this section we review related work on microsculpturing identification, image
+texture analysis, and insect classification.
+
+\subsection{Microsculpturing identification}
 
@@
 
-% \subsection{Texture analysis}
+\subsection{Image texture analysis}
 
@@
 
-% \subsection{Insect classification}
+\subsection{Insect classification}
 
@@
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/sr103.png}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/r103.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/sr111.png}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/r111.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/ss105.png}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/s105.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/ss107.png}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/s107.png}
     \end{subfigure}
@@
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat103.png}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/r103.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat111.png}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/r111.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat105.png}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/s105.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/heat107.png}
+        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/s107.png}
     \end{subfigure}
@@
         15x15 patches. The third row shows the results of using size of 19x19
-        patches.}
+        patches. In the images, the color red represents the background class,
+        green represents the smooth class, and the blue represent the rough
+        class.}
     \label{fig:kviews_segmented}
@@
 \FloatBarrier Some classified results from the K-views are shown in Figure
-\ref{fig:kviews_segmented}. A training set of 500 patches for background, 500
-patches for rough, and 500 patches for smooth are applied with patch sizes of
-15x15 and 19x19. These patches are extracted from two rough and two smooth
-images. The K-views algorithm converges early although we set the maximum number
-of iterations to be 1000.
+\ref{fig:kviews_segmented}. A training set of 550 patches for background, 455
+patches for rough, and 493 patches for smooth are used with patch sizes of 15x15
+and 19x19. These patches are extracted from two rough and two smooth images with
+sizes of 608x608. The K-views algorithm converges early although we set the
+maximum number of iterations to be 1000. These experimental results are
+preliminary due to the minimal number of source images used and future work will
+explore a more expansive approach to the problem.
 
@@
 Additionally, a model that is pre-trained on a diverse image task such as ResNet
-can be transferred to our domain of texture analysis. However, a deep learning
-algorithm created specifically for the domain of texture analysis had the best
-results in our experiments. In future work, we will continue to explore texture
-classes present in the dataset which are not captured by the binary class
-system. All code is publicly available on GitHub
-(https://github.com/ngngardner/cuticulus).
+can be transferred to our domain of texture analysis \footnote{All code is
+    publicly available on GitHub (https://github.com/ngngardner/cuticulus)}.
+However, a deep learning algorithm created specifically for the domain of
+texture analysis had the best results in our experiments. In future work, we
+will continue to explore texture classes present in the dataset which are not
+captured by the binary class system. We will also investigate application of
+texture-based deep maching learning in image texture analysis.
 

### 2022-08-20 paper/main.tex
fix fig 7
address comments

@@
 \title{Investigation of ant cuticle dataset using image texture
-    analysis methods}
+    analysis}
 
@@
         Kennesaw, GA, USA}%
-    \addr{\affilnum{4}}{Department of Mathematics, Science, and Informatics,
+    \addr{\affilnum{4}}{Department of Informatics and Mathematics,
         Mercer University, Atlanta, GA, USA}%
@@
     In this study, we employ statistical image texture analysis and deep machine
-    learning methods to automatically group similar ant species based on
-    morphological features. We establish a public database of ant cuticle images
-    for research. We provide a comparative study of the performance of image
-    texture classification and deep machine learning methods on ant cuticle
-    images. Our results show that the deep learning methods are giving higher
-    accuracy in recognizing ant cuticle textures. Our experiments also reveal
-    that the deep learning networks designed for image texture performs better
-    than the general deep learning networks.
+    learning methods to classify similar ant species based on morphological
+    features. We establish a public database of ant cuticle images for research.
+    We provide a comparative study of the performance of image texture
+    classification and deep machine learning methods on an ant cuticle dataset.
+    Our results show that the deep learning methods give higher accuracy than
+    statistical methods in recognizing ant cuticle textures. Our experiments
+    also reveal that the deep learning networks designed for image texture
+    performs better than the general deep learning networks.
 \end{abstract}
@@
 desiccation, and serves as a canvas for advertising visual and chemical signals
-\cite{gullan_insects_2009}. Research has focused on the macrostructures and
-internal chemical components that describe the functionality of the exoskeleton
-\cite{gunderson_insect_1989}. More recent work is being done to understand the
-functional aspects of external cuticle micro sculpturing
+\cite{gullan_insects_2009}. Earlier research has focused on the macrostructures
+and internal chemical components that describe the functionality of the
+exoskeleton \cite{gunderson_insect_1989}. More recent work is being done to
+understand the functional aspects of external cuticle micro sculpturing
 \cite{muthukrishnan_insect_2020, watson_diversity_2017}.
@@
 used data types for insect analysis is image data. To develop an image-based
-system for insect analysis, we can take advantage of existing work in general
+system for insect analysis, we can utilize existing work in general
 image analysis methods.
 
-We examine ants (\textit{Formicidae}) as they display an extreme diversity of
+We study ants (\textit{Formicidae}) as they display an extreme diversity of
 cuticle micro sculpturing across all subfamilies. Microsculpture ranges from
@@
 rigidness, resist abrasion, increase internal and external surface area, resist
-microbial growth, and rear beneficial anti-biotic producing bacteria
+microbial growth, and cultivate beneficial anti-biotic producing bacteria
 \cite{johnson_effect_2011, bruckner_relationship_2017, currie_coevolved_2006}.
-These specific functions may be associated with certain sculpturing types. In
-order to analyze those functions, it is necessary to segment the image and group
-similar textures for further analysis.
+These specific functions may be associated with certain sculpturing types. To
+analyze those functions, we first segment the image and group similar textures
+for further analysis.
 
 There is a large variety of ant species and most species are diverse in terms of
-size, shape, behaviors, and cuticle textures \cite{harris_glossary_1979}.
-Based on our initial observation of the cuticle sculpturing in ants, it appears
-that some type of textural structures can be identified in images. Therefore,
-this image dataset may be worth to be explored from the image texture analysis
-point of view for the further knowledge discovery. Image texture analysis has
-been used in image processing for many decades \cite{hung_image_2019}. Due to its
-diversity of images on scales, smoothness, coarseness, microtexture, and
-macrotexture occurring in images, it is a difficult problem for image textural
-interpreters. With a rapid development of deep machine learning, deep neural
-networks have been frequently used in image texture analysis
-\cite{sun_robust_2014,liu_bow_2019,liu_pestnet_2019}.
+size, shape, behaviors, and cuticle textures \cite{harris_glossary_1979}. Based
+on our initial observation of the cuticle sculpturing in ants, it appears that
+some type of textural structures can be identified in images. Therefore, it is
+worthwhile to explore image texture analysis for further knowledge discovery.
+Image texture analysis has been used in image processing for many decades
+\cite{hung_image_2019}. Due to the diversity of images on scales, smoothness,
+coarseness, microtexture, and macrotexture occurring in images, it is a
+difficult problem. With a rapid development of deep machine learning, deep
+neural networks have been frequently used in image texture analysis in other
+fields \cite{sun_robust_2014,liu_bow_2019,liu_pestnet_2019}.
 
@@
 explore the categorization of different ant cuticle textures using image texture
-analysis methods. We share how the dataset of ant cuticle images is established
-in Section 3. Our experiment shows that the deep learning networks designed for
-image texture analysis give a better accuracy than the K-views clustering
-method.
+analysis methods. We document how the dataset of ant cuticle images is
+established. Our experiment shows that the deep learning networks designed for
+image texture analysis give a better accuracy than classical K-views clustering.
 
@@
 Section 3 describes the dataset preparation and image texture analysis methods
-used in our experiments. Section 4 gives the experimental results, visualization
-of both raw and processed images and analysis. The conclusion then follows.
+used in our experiments. Section 4 gives the experimental setup and methods
+used. Section 5 shares visualization of both raw and processed images, and
+analysis of experimental results. The conclusion then follows.
 
@@
 five categories: smooth, striate, punctate, reticulate, and tuberous
-\cite{hellenbrand_ant_2022}. Using these categories to classify 11,722 ant
-species, we found that the earliest ants were likely smooth and that cuticle
+\cite{hellenbrand_ant_2022}. We used these categories to classify 11,722 ant
+species, and found that the earliest ants were smooth and that cuticle
 microsculpturing independently evolved later in multiple lineages
-\cite{hellenbrand_ant_2022}. An even simpler classification scheme is to divide
-ants into only two categories: smooth and rough. Because microsculptuing has
-evolved numerous times in ants, this binary classification scheme is sufficient
-to detect conditions that select for any sculpturing at all compared to a smooth
-state.
+\cite{hellenbrand_ant_2022}. A simple classification scheme divides ants into
+two categories: smooth and rough. Because microsculptuing has evolved numerous
+times in ants, this binary classification scheme is sufficient.
 
-Using either a simplified or binary classification scheme, the next challenge is
-figuring out how to apply either of these schemes to large numbers of species.
-We live in an era of big data, and insect researchers have access to millions of
-images available through digital museum collections \cite{beaman_mass_2012}.
-Going through these digital collections manually is not feasible for large
-projects. Focusing only on ants, for example, there are over 14,000 described
-species, and each species has multiple worker and reproductive castes that may
-vary in their microsculpturing patterns. Even within an individual ant,
-microsculpturing typically varies among and within body segments, which
-exponentially increases the number of images that must be assessed. Applied
-computing offers a solution to this problem through developing tools that
-automate trait classification. Automated approaches would allow researchers to
-leverage existing image repositories to extract large trait datasets that can
-then be used to detect evolutionary patterns.
+The next challenge is to find out how to apply these schemes to large numbers of
+species. Insect researchers have access to millions of images available through
+digital museum collections \cite{beaman_mass_2012}. Going through these digital
+collections manually is not feasible for large projects. There are over 14,000
+described species of ants, and each species has multiple worker and reproductive
+castes that may vary in their microsculpturing patterns. Even within an
+individual ant, microsculpturing typically varies among and within body
+segments. Machine learning offers a solution to this problem through developing
+tools that automate trait classification. Automated approaches would allow
+researchers to leverage existing image repositories to extract large trait
+datasets that can then be used to detect evolutionary patterns.
 
@@
 
-Image texture analysis is an important part of image interpretation as many
-objects appearing in the textural patterns either partially or completely inside
-an image. This phenomenon appears frequently in many images we encountered in
-our environment as almost everybody is using the iPhone to capture the photos.
-Scientists are also studying their subjects depending on taking the images
-remotely such as satellite images. With an abundant number of textural images
-surrounding us, it is essential to have a reliable automatic tool for the
-textural image interpretation.
+Pattern recognition methods have been developed for textural images based on the
+spatial relationship of gray levels of pixels encoded in the texture
+\cite{haralick_image_1985,hung_image_2019}. Those methods extract texture
+features and the classification and segmentation methods are used for the
+interpretation either using the supervised or unsupervised approaches
+\cite{fukunaga_introduction_2013}. The interpretation method is usually modified
+from the traditional pattern recognition methods. For example, the gray-level
+co-occurrence matrix (GLCM) and local binary patterns (LBP) are two common
+methods for extracting features \cite{haralick_textural_1973, goos_gray_2000}.
 
-Computer methods are developed for textural images based on the spatial
-relationship of gray levels of pixels encoded in the texture
-\cite{haralick_image_1985,hung_image_2019}. Those methods by de-facto follow the
-steps used in the statistical pattern recognition
-\cite{fukunaga_introduction_2013}. Therefore, textural features are extracted,
-and the classification and segmentation methods are used for the interpretation
-either using the supervised or unsupervised approaches. The interpretation
-method is usually taken and modified from the traditional pattern recognition
-methods. For example, the gray-level co-occurrence matrix (GLCM) and local
-binary patterns (LBP) are two of methods for extracting features
-\cite{haralick_textural_1973, goos_gray_2000}.
+Classical image texture analysis methods can be grouped into four categories:
+statistical, structural, model-based, and transform-based methods
+\cite{bharati_image_2004}. The statistical method is often used by feature
+extraction methods and then fed into the classification algorithm. Markov random
+field models have also been used for textural image interpretation
+\cite{hassner_use_1981, cross_markov_1983}. Transform-based methods use some
+functions to decompose an image texture into a set of basic feature images.
+Gabor filters and Wavelet expansions are two of the widely used approaches
+\cite{bovik_multichannel_1990}.
 
-Similar to the traditional pattern recognition, classical image texture analysis
-methods can be grouped into four categories: statistical, structural,
-model-based, and transform-based methods \cite{bharati_image_2004}. Among all of
-the categories, the statistical method is often used by employing the method
-from the pattern recognition principle.  In doing so, an image texture is
-represented as feature vector using feature extraction methods and then fed into
-the algorithm. Markov random field models are also studied for textural image
-interpretation \cite{hassner_use_1981, cross_markov_1983}. The transform-based
-methods use some functions to decompose an image texture into a set of basic
-feature images. Gabor filters and Wavelet expansions are two of the widely used
-approaches \cite{bovik_multichannel_1990}.
+\subsection{Deep learning}
 
-While most feature vectors are established based on a single pixel, Liu et al.
-gave a comprehensive survey on textural characterization in which they call this
-feature extraction as texture representation \cite{liu_bow_2019}. Their survey
-on Bag of Words and Convolutional Neural Networks (CNNs)
-\cite{krizhevsky_imagenet_2017} are the spatial relationship concepts. The
-K-views was developed for taking the spatial information in the clustering
-approach \cite{hung_image_2019}.
+Liu et al. gave a comprehensive survey on textural characterization in which
+they call this feature extraction as texture representation \cite{liu_bow_2019}.
+Their survey on bag of words and convolutional neural networks (CNNs)
+\cite{krizhevsky_imagenet_2017} provides information on spatial relationship
+concepts. The K-views was developed for taking the spatial information in the
+clustering approach \cite{hung_image_2019}.
 
 In the past decade, deep learning using CNNs has emerged as the mainstream
-technology for image analysis. Following this trend, various CNN-based network
+technology for image analysis. Following this trend, various related network
 architectures have been designed to specifically characterize textural images.
-Among these, the Fisher-vector CNN descriptor (FV-CNN) proposed by Cimpoi et
-al. \cite{cimpoi_deep_2015} is widely accepted as one of the most important
+Among these, the Fisher-vector CNN descriptor (FV-CNN) proposed by Cimpoi et al.
+\cite{cimpoi_deep_2015} is widely accepted as one of the most important
 pioneering works. It applies FV pooling to deep features obtained via a CNN
-pre-trained using the ImageNet \cite{krizhevsky_imagenet_2017} to obtain encoded features
-for texture classification. Although it is capable of achieving much improved
-classification accuracy comparing to traditional hand-crafted texture features,
-FV-CNN does not support an end-to-end learning, where feature extraction,
-encoding, and classification are separated from each other. To achieve an
-end-to-end learning, Zhang et al. \cite{zhang_deep_2017} proposed the deep texture
-encoding network (DeepTEN), in which a novel texture encoding layer is added to
-a standard CNN architecture. Then, Xue et al. \cite{xue_deep_2018} constructed the
-deep encoding pooling network (DEP), which improves over DeepTEN by integrating
-local spatial characteristics into the texture representation. Based on DeepTEN
-and DEP, Hu et al. \cite{hu_multi_2019} further developed the multi-level texture
-encoding and representation (MuLTER) network, which embeds a learnable encoding
-module at each convolutional layer so that encoding is performed for both
-low-level and high-level features, yielding a multi-level texture
-representation.
+pre-trained using the ImageNet \cite{krizhevsky_imagenet_2017} to obtain encoded
+features for texture classification. FV-CNN is capable of achieving much
+improved classification accuracy comparing to traditional hand-crafted texture
+features. However, CNN-based deep learning is only used in its feature
+extraction module. Its encoding and classification modules still use traditional
+learning methods, thus not supporting an end-to-end deep learning. In the
+context of texture image classification, an end-to-end deep learning model
+refers to a neural network architecture which consists of all the modules
+(feature extraction, encoding and classification) between the initial input data
+and the final output classification result, with all the modules being trained
+simultaneously.
+
+To achieve an end-to-end learning, Zhang et al. \cite{zhang_deep_2017} proposed
+the deep texture encoding network (DeepTEN), in which a novel texture encoding
+layer is added to a standard CNN architecture. Then, Xue et al.
+\cite{xue_deep_2018} constructed the deep encoding pooling network which
+improves over DeepTEN by integrating local spatial characteristics into the
+texture representation. Based on these two methods, Hu et al. \cite{hu_multi_2019}
+further developed the multi-level texture encoding and representation (MuLTER)
+network, which embeds a learnable encoding module at each convolutional layer so
+that encoding is performed for both low-level and high-level features, yielding
+a multi-level texture representation.
 
@@
 
-\subsection{Insect classification}
+\subsection{Classification of insects}
 
@@
     feature transform} (SIFT) features \cite{gotlieb_texture_1990,
-    lowe_distinctive_2004}. The method proposed by Feng et al.
-\cite{feng_automated_2013} is used to classify 50 different moth species across
-8 families \cite{feng_automated_2013}. The results from Feng et al.
+    lowe_distinctive_2004}. The method is used to classify 50 different moth species
+across 8 families \cite{feng_automated_2013}. Their results
 \cite{feng_automated_2013} suggest that traditional feature extraction
@@
 \textit{Centruroides limpidus} and \textit{Centruroides noxius}. After applying
-background distinction based on dynamic color threshold, Urteaga et al.
-\cite{urteaga_scorpions_2016} apply feature extraction to extract features from
-the separated scorpion image such as aspect ratio, rectangularity, and
-compactness. Urteaga et al. \cite{urteaga_scorpions_2016} apply three different
-classification models to classify the image as one of the species: Artificial
-Neural Network, Regression Tree, and Random Forest classifiers
-\cite{urteaga_scorpions_2016}. The results from Urteaga et al.
-\cite{urteaga_scorpions_2016} show that after background removal,
-characteristics from the entire body of the scorpion can be used to create a
-binary classifier that can classify the image as one of the two species.
+background distinction based on dynamic color threshold, they extract features
+from the separated scorpion image such as aspect ratio, rectangularity, and
+compactness. They apply three different models to classify the image as one of
+the species: artificial neural network, regression tree, and random forest
+classifiers \cite{urteaga_scorpions_2016}. Their results show that after
+background removal, characteristics from the entire body of the scorpion can be
+used to create a binary classifier that can classify the image as one of the two
+species.
 
@@
 In this section, we describe the creation of the custom dataset used in this
-research. In our dataset, we use ant head images from AntWeb
-\cite{perrichot_antweb_2012} and define two categories for them based on the
-appearance of the cuticle texture: \textit{rough} and \textit{smooth}. We also
-refer to these ant head images, which describe the pose of the ant, as ant
-cuticule images, due to the clear cuticle texture present on the head. Some
-randomly selected images from each category are shown in Figures
+research. We use ant head images from AntWeb \cite{perrichot_antweb_2012} and
+define two categories for them based on the appearance of the cuticle texture:
+\textit{rough} and \textit{smooth}. For each species used, there exists only one
+ant head image. We also refer to these ant head images, which describe the pose
+of the ant, as ant cuticule images, due to the clear cuticle texture present on
+the head. Some randomly selected images from each category are shown in Figures
 \ref{fig:smooth-cuticle-texture} and \ref{fig:rough-cuticle-texture}.
@@
 texture: smooth, punctate, striate, reticulate, and tuberous. For simplicity, we
-work only with the two main categories. The training set identifications were
+worked only with the two main categories, and the other 4 categories were
+combined to the single rough category. The training set identifications were
 reviewed together as a group by the assistants. Once training was complete,
@@
 
-To collect the images, the assistants followed the taxonomy information
-available in the master spreadsheet to the appropriate
-AntWeb\footnote{\url{https://www.antweb.org/}} page. In many cases, there are
-multiple ant head images of the same species, and occasionally there are
-multiple image resolution available from a single image. To simplify the data
-collection process, the assistants were instructed to download the first ant
-head image of the species being identified in the highest resolution possible.
-Each image was named with an identifier that corresponds with the row number in
-the master spreadsheet. The same ant head images that were downloaded in the
-data collection phase were the same ones used in the sculpture identification
-protocol. Ant species which did not have any images of the head were excluded
-from the dataset. Additionally, ant species which only had a head image of a
-queen ant were excluded from the dataset.
+To collect the images from AntWeb\footnote{\url{https://www.antweb.org/}}, the
+assistants followed the taxonomy information available in the master spreadsheet
+to the appropriate AntWeb page. In many cases, there are multiple ant head
+images of the same species, and occasionally there are multiple image resolution
+available from a single image. To simplify the data collection process, the
+assistants were instructed to download the first ant head image of the species
+being identified in the highest resolution possible. Each image was named with
+an identifier that corresponds with the row number in the master spreadsheet.
+The same ant head images that were downloaded in the data collection phase were
+the same ones used in the sculpture identification protocol. Ant species which
+did not have any images of the head were excluded from the dataset.
+Additionally, ant species which only had a head image of a queen ant were
+excluded from the dataset.
 
-Ant specimen images taken from AntWeb \cite{perrichot_antweb_2012} are created
-by different photographers and therefore have different attributes, such as
+Ant specimen images in AntWeb \cite{perrichot_antweb_2012} are created by
+different photographers and therefore have different attributes, such as
 environment, resolution, and lighting. In the ant head images, the ant head is
@@
 
-The completed dataset of contains 2,499 images. 1072 samples of rough textured
-ant cuticle textures comprise 43\% of the dataset. The remaining 1427 samples of
-smooth textured ant cuticle textures comprise 57\% of the dataset. Due to the
-variety of the ant head image attributes, we apply simple preprocessing before
-the images are used in our model. We want the images to have a uniform size for
-simplicity in our classification process. Since the ant head images are
-typically centered in the image, we apply a center crop to each image to create
-a square image of the same size. Once the image is square, we resize each image
-to a fixed size of 256x256 pixels. We leave other discrepancies in the images
-untouched. A summary of the resulting dataset is shared in Table
-\ref{tab:dataset-summary}.
+The completed dataset\footnote{All code is publicly available on GitHub
+    \url{https://github.com/ngngardner/cuticulus}} contains 2,499 images as a
+4-dimensional array with shape of (samples, rows, columns, channels), of
+which there are 3 channels extracted from the RGB images. 1072 samples of
+rough textured ant cuticle textures comprise 43\% of the dataset. The
+remaining 1427 samples of smooth textured ant cuticle textures comprise 57\%
+of the dataset. Due to the variety of the ant head image attributes, we
+apply simple preprocessing before the images are used in our model. We want
+the images to have a uniform size for simplicity in our classification
+process. Since the ant head images are typically centered in the image, we
+apply a center crop to each image to create a square image of the same size.
+Once the image is square, we resize each image to a fixed size of 256x256
+pixels. We leave other discrepancies in the images untouched. A summary of
+the resulting dataset is shared in Table \ref{tab:dataset-summary}.
 
@@
         \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r111.png}
-        \caption*{Smooth texture}
+        \caption*{Rough texture}
     \end{subfigure}
@@
         \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s105.png}
-        \caption*{Rough texture}
+        \caption*{Smooth texture}
     \end{subfigure}
@@
 can be transferred to our domain of texture analysis \footnote{All code is
-    publicly available on GitHub (https://github.com/ngngardner/cuticulus)}.
+    publicly available on GitHub \url{https://github.com/ngngardner/cuticulus}}.
 However, a deep learning algorithm created specifically for the domain of

### 2022-08-29 paper/main.tex
address comments

@@
 \def\typeofarticle{Research Article} 
-\def\currentvolume{3} 
-\def\currentissue{1}
+\def\currentvolume{2} 
+\def\currentissue{2}
 \def\currentyear{2022} 
@@
     We provide a comparative study of the performance of image texture
-    classification and deep machine learning methods on an ant cuticle dataset.
+    classification and deep machine learning methods on this ant cuticle dataset.
     Our results show that the deep learning methods give higher accuracy than
@@
 statistical, structural, model-based, and transform-based methods
-\cite{bharati_image_2004}. The statistical method is often used by feature
-extraction methods and then fed into the classification algorithm. Markov random
-field models have also been used for textural image interpretation
-\cite{hassner_use_1981, cross_markov_1983}. Transform-based methods use some
-functions to decompose an image texture into a set of basic feature images.
-Gabor filters and Wavelet expansions are two of the widely used approaches
-\cite{bovik_multichannel_1990}.
+\cite{bharati_image_2004}. The statistical method is often used for image
+classification after feature extraction. Markov random field models have also
+been used for textural image interpretation \cite{hassner_use_1981,
+    cross_markov_1983}. Transform-based methods use some functions to decompose an
+image texture into a set of basic feature images. Gabor filters and wavelet
+expansions are two of the widely used approaches \cite{bovik_multichannel_1990}.
 
@@
 Among these, the Fisher-vector CNN descriptor (FV-CNN) proposed by Cimpoi et al.
-\cite{cimpoi_deep_2015} is widely accepted as one of the most important
-pioneering works. It applies FV pooling to deep features obtained via a CNN
-pre-trained using the ImageNet \cite{krizhevsky_imagenet_2017} to obtain encoded
-features for texture classification. FV-CNN is capable of achieving much
-improved classification accuracy comparing to traditional hand-crafted texture
-features. However, CNN-based deep learning is only used in its feature
-extraction module. Its encoding and classification modules still use traditional
-learning methods, thus not supporting an end-to-end deep learning. In the
-context of texture image classification, an end-to-end deep learning model
-refers to a neural network architecture which consists of all the modules
-(feature extraction, encoding and classification) between the initial input data
-and the final output classification result, with all the modules being trained
-simultaneously.
+\cite{cimpoi_deep_2015} is widely accepted as one of the pioneering works. It
+applies Fisher-vector pooling to deep features obtained via a CNN pre-trained
+using the ImageNet \cite{krizhevsky_imagenet_2017} to obtain encoded features
+for texture classification. FV-CNN is capable of achieving higher classification
+accuracy comparing to traditional hand-crafted texture features. However,
+CNN-based deep learning is only used in its feature extraction module. Its
+encoding and classification modules still use traditional learning methods, thus
+not supporting an \textit{end-to-end} deep learning. In the context of texture
+image classification, an end-to-end deep learning model refers to a neural
+network architecture which consists of all the modules (feature extraction,
+encoding and classification) between the initial input data and the final output
+classification result, with all the modules being trained simultaneously.
 
@@
 extract the features through the training. In addition, the deep learning
-networks achieve the higher accuracy than that of the classical approaches,
+networks can 6achieve the higher accuracy than that of the classical approaches,
 although the deep networks require much more data for training.
@@
 for training a classifier to classify an image between 10 randomly selected moth
-species.
+species. Marques et al. \cite{marques_ant_2018} propose an ensemble-based method
+to classify a large number of ant genera. They use ant images sourced from AntWeb
+in order to classify 44,806 specimens with at least 1 picture into 57 genera.
 
@@
 
-The completed dataset\footnote{All code is publicly available on GitHub
+The completed dataset\footnote{The dataset is available on GitHub
     \url{https://github.com/ngngardner/cuticulus}} contains 2,499 images as a
@@
 apply a center crop to each image to create a square image of the same size.
-Once the image is square, we resize each image to a fixed size of 256x256
+Once the image is square, we resize each image to a fixed size of 256 $\times$ 256
 pixels. We leave other discrepancies in the images untouched. A summary of
@@
         Smooth images & 1427             \\
-        Image size    & 256x256 \\
+        Image size    & 256 $\times$ 256 \\
         \bottomrule
@@
 
-\section{Methods for comparative analysis and experimental setup}
+\section{K-views, K-means, and deep learning models for experiments}
 
@@
 recognition \cite{lloyd_least_1982}. It is a single pixel based classification
-algorithm for images. The K-views is an algorithm that uses several
-characteristic views for the classification of images \cite{hung_use_2002}. The
-K-views algorithm is suitable for classifying image textures that have basic
-local patterns repeated in a periodic manner. In contrast to the K-means,
-K-views looks at the neighboring pixels for providing the spatial relationship
-for classification.  Several variations of the basic K-views model have been
-proposed \cite{yang_image_2003, lan_improved_2010}.
+algorithm for images. Based on the concept of K-means, K-views was developed,
+which uses characteristic views for image texture representation
+\cite{hung_use_2002}. The K-views algorithm is suitable for classifying image
+textures that have basic local patterns repeated in a periodic manner. In
+contrast to the K-means, K-views looks at the neighboring pixels for providing
+the spatial relationship for classification.  Several variations of the basic
+K-views model have been proposed \cite{yang_image_2003, lan_improved_2010}.
 
-For the feasibility study in this ant image datasets, we use both K-means and
-K-views in our experiments as both are statistical clustering methods. The
-experiments may indicate how well the statistical approach will do in this type
-of textural images. In pre-processing the dataset, we apply the Gabor filter to
-transform the gray-scale ant image into four Gabor-filtered channels, subtract
-the gray-scale image from each channel, take the absolute value of the resulting
-images, and then normalize them. This pre-processing will highlight some of
-characteristics features such as edge-like in the image. This step is very
-useful in feeding the datasets to the K-views algorithms. A 4-channel Gabor
-filter is applied in the pre-processing step for the K-means algorithm. The
-theta $\theta$ parameters in the Gabor filters used are multiples of $\pi/4$
-radians starting from $0$ to $\pi$ radians. The Gabor filter lambda $\lambda$
-parameter is equal to $\pi$.
+For the feasibility study in this ant image datasets, we tested the statistical
+clusterings methods and deep learning methods separately. Both K-means and
+K-views were used for statistical clustering methods. The experiments may
+indicate how well the statistical approach and deep learning networks will do in
+this type of textural images. In pre-processing the dataset for the statistical
+clustering methods, we apply the Gabor filter to transform the gray-scale ant
+image into four Gabor-filtered channels, subtract the gray-scale image from each
+channel, and take the absolute value of the resulting images. This
+pre-processing will highlight some of characteristic features in textures, such
+as edge-like in the image. This step is very useful in feeding the datasets to
+the K-views algorithms. The theta $\theta$ parameters in the Gabor filters used
+are multiples of $\pi/4$ radians starting from $0$ to $\pi$ radians. The Gabor
+filter lambda $\lambda$ parameter is equal to $\pi$.
 
@@
 
-Our first deep learning model is \textit{visual geometry group} (VGG), a
-convolutional neural network that takes advantage of very small convolutional
-filters in a deep network architecture \cite{simonyan_very_2015}. We compare
-four architectures of VGG: VGG11, VGG13, VGG16, and VGG19. The primary
-difference between the architectures is the number of layers in each model. Our
-second model is \textit{residual network} (ResNet), a deep network architecture
-that includes shortcut connections between layers (residual connections)
-\cite{he_deep_2015}. We compare three architectures of ResNet: ResNet18,
-ResNet50, and ResNet101. Again, the primary difference between the architectures
-is the number of layers in each model.
+The first deep learning model used in our experiment is \textit{visual geometry
+    group} (VGG), a convolutional neural network that takes advantage of very
+small convolutional filters in a deep network architecture
+\cite{simonyan_very_2015}. We compare four architectures of VGG11, VGG13,
+VGG16, and VGG19. The primary difference between the architectures is the
+number of layers in each model. The second deep learning model used in our
+experiments is \textit{residual network} (ResNet), a deep network
+architecture that includes shortcut connections between layers (residual
+connections) \cite{he_deep_2015}. We compare three architectures of
+ResNet18, ResNet50, and ResNet101. Again, the primary difference between the
+architectures is the number of layers in each model.
 
-For our ResNet models, we have two versions: randomized and pretrained. The
-randomized version is the same architecture, but the weights are randomly
+For the initial ResNet models, we have two versions: randomized and pretrained.
+The randomized version is the same architecture, but the weights are randomly
 initialized. The pretrained version has weights from training on the CIFAR
-dataset, an image dataset with 1000 classes. In this case, we are fine-tuning
-the pretrained model. For VGG, we are only using the randomized version. The
-base VGG architecture also has an output layer of size 1000. Since we are
-working with a binary classification problem, we modify the architecture for all
-models to have an output layer of size 2. Each model is trained over 100 epochs,
-using stochastic gradient descent with momentum. The batch size is set to 16
-images. We apply a learning rate of 0.001 and momentum parameter of 0.9.
+dataset, an image dataset with 1000 classes. Since the pretrained weights are
+readily available, we also evaluate a fine-tuned model. For VGG, we only use the
+randomized version. The base VGG architecture also has an output layer of size
+1000. Since we are working with a binary classification problem, we modify the
+architecture for all models to have an output layer of size 2. Each model is
+trained over 100 epochs, using stochastic gradient descent with momentum. The
+batch size is set to 16 images. We apply a learning rate of 0.001 and momentum
+parameter of 0.9.
 
@@
 methods specifically designed for textural images. For this examination, we
-adopt the deep residual pooling (DRP) network developed by Mao et al.
-\cite{mao_deep_2021}. The DRP framework consists of a unique residual pooling
-layer, which is formed by a residual encoding module followed by an aggregation
-module. The residual encoding module extracts relevant spatial information,
-while the aggregation module performs averaging to obtain orderless
-low-dimension features for classification.
+adopt the deep residual pooling (DRP) network \cite{mao_deep_2021}. The DRP
+framework consists of a unique residual pooling layer, which is formed by a
+residual encoding module followed by an aggregation module. The residual
+encoding module extracts relevant spatial information, while the aggregation
+module performs averaging to obtain orderless low-dimension features for
+classification.
 
@@
 classifier layer. Additionally, an auxiliary classifier layer may also be used.
-In our experiments, we follow these strategies. For each scenario, we conduct
+In our experiments, we follow the same strategy. For each scenario, we conduct
 two experiments. One experiment employs weights that are randomly initialized,
@@
 
-To handle the imbalance of the dataset, we apply undersampling for each class for
-the training dataset. By using random stratified sampling, we construct a
-training set with 800 images per class. The remaining images are randomly split
-between test and validation, which turns out to roughly a 60\%/20\%/20\% train,
-test, and validation data split. With 272 rough samples and 627 smooth samples
-left over after the stratified split, the test dataset has roughly 136 rough
-samples and 313 smooth samples. Since these leftover samples are split with code
-by 50\% there will be some rounding variance and therefore the test dataset
-built at run-time will not always have exactly the same number of samples.
+To handle the imbalance of the dataset, we apply undersampling for each class
+for the training dataset \cite{tiittanen_novel_2022}. By using random stratified
+sampling, we construct a training set with 800 images per class. The remaining
+images are randomly split between test and validation, which turns out to
+roughly a 60\%/20\%/20\% train, test, and validation data split. With 272 rough
+samples and 627 smooth samples left over after the stratified split, the test
+dataset has roughly 136 rough samples and 313 smooth samples. Since these
+leftover samples are split with code by 50\% there will be some rounding
+variance and therefore the test dataset built at run-time will not always have
+exactly the same number of samples.
 
@@
 
-We evaluate the performance of the models according to standard evaluation
-methods. Since we are working with a binary classification problem, we use a
-standard confusion matrix to evaluate the accuracy, precision, and F1 score. We
-also apply Grad-CAM with manual inspection to visualize the activation weights
-for classified images to visualize which features lead to the classification
-result. Finally, we apply t-SNE to visualize the separation learned for the
-model to further analyze the classifications made by the model.
+We evaluate the performance of the models using accuracy, precision, and F1
+score. We also apply Grad-CAM with manual inspection to visualize the activation
+weights for classified images to show which features lead to the
+classification result. Finally, we apply t-SNE to visualize the separation
+learned for the model to further analyze the classifications made by the model.
+
+\section{Experimental results and analysis}
 
@@
 
-Experiments are run on an Ubuntu 18.04 LTS Lambda Labs GPU server. The server
-contains 8 NVIDIA GeForce RTX 2080 Ti graphics cards with 12GB of memory each.
-The server uses an Intel Xeon Silver 4116 with 48 total threads and maximum
-frequency of 3.000 GHz, and has 256GB of RAM.
-
-\section{Results and analysis}
+Experiments were run on an Ubuntu 18.04 LTS Lambda Labs GPU server. The server
+contained 8 NVIDIA GeForce RTX 2080 Ti graphics cards with 12GB of memory each.
+The server used an Intel Xeon Silver 4116 with 48 total threads and maximum
+frequency of 3.000 GHz, and has 256GB of RAM. K-means algorithm comes from the
+scikit-learn\footnote{\url{https://scikit-learn.org/}} library. The K-views
+implementation was based on the K-means implementation from scikit-learn. VGG
+and ResNet models come from the
+torchvision\footnote{\url{https://pytorch.org/vision}} library. DRP model code
+was tweaked from the original author's uploaded code on
+GitHub\footnote{\url{https://github.com/maoshangbo/DRP-Texture-Recognition}}.
 
@@
 
-To begin, we share the results on the each algorithm specified in the
-methodology section using the evaluation metrics described. The results are
-shown in Table \ref{tab:results}. It should be noted that due to the class
-imbalance in the dataset, the F1 score is the preferable metric to the accuracy.
+We share the results on each algorithm specified in Section 4. Both the K-means
+and K-views algorithms are unsupervised. All of the deep learning methods used
+are supervised. The results are shown in Table \ref{tab:results}. It should be
+noted that due to the class imbalance in the dataset, the F1 score is the
+preferable metric to the accuracy.
 % Therefore, the results have been sorted by F1 score in descending order.
@@
 \FloatBarrier
-The results in the previous sections show that the fine-tuned ResNet models
-outperform the VGG and randomly initialized ResNet models on the task of ant
-head image classification. Additionally, the DRP models perform better than the
-the general deep learning models VGG and ResNet on the task of ant head image
-classification.The fine-tuned DRP single-layer model with auxiliary classifier
-performed the best with an average F1 score of 0.92. We further analyze the
-separation learned by both ResNet101 models in the following section.
+
+The results show that the fine-tuned ResNet models outperform the VGG and
+randomly initialized ResNet models on the task of ant head image classification.
+The DRP models perform better than the the general deep learning models VGG and
+ResNet on the task of ant head image classification. The fine-tuned DRP
+single-layer model with auxiliary classifier performed the best with an average
+F1 score of 0.92. We further analyze the separation learned by both ResNet101
+models in the following section.
 
@@
     \caption{t-SNE visualization of the embeddings of the second to last layer
-        of the randomly initialized ResNet101 model trained on ant head image
-        dataset.}
+        of the randomly initialized ResNet101 model trained on dataset.}
     \label{fig:resnet101_tsne}
@@
 the randomly initialized ResNet101 model using t-SNE dimensionality reduction.
-First, we run the dataset preprocessing method and initialize both models. Then,
-both models are trained according to the training parameters and the state of
-each model is saved. To visualize the deep extracted features, we modify each
-model to obtain the embeddings of the second to last layer. Then, we use the
-t-SNE algorithm to reduce the dimensionality of the embeddings to 2 dimensions.
-We plot side-by-side the ground truth and predicted labels for each model.
-Figure \ref{fig:resnet101_tsne} shows the results of the trained randomly
-initialized model and Figure \ref{fig:fresnet101_tsne} shows the results of the
-fine-tuned model. Based on the visual results of the t-SNE visualization, we can see that the
-fine-tuned model learned a stronger separation of the two classes, which
-reinforces the results that the fine-tuned model received a higher average
-accuracy.
+To visualize the deep extracted features, we modify each model to obtain the
+embeddings of the second to last layer. We plot side-by-side the ground truth
+and predicted labels for each model. Figure \ref{fig:resnet101_tsne} shows the
+results of the trained randomly initialized model and Figure
+\ref{fig:fresnet101_tsne} shows the results of the fine-tuned model. Based on
+the visual results of the t-SNE visualization, we can see that the fine-tuned
+model learned a stronger separation of the two classes, which reinforces the
+results that the fine-tuned model received a higher average accuracy.
 
@@
     \caption{t-SNE visualization of the embeddings of the second to last layer
-        of the fine-tuned ResNet101 model trained on ant head image dataset.}
+        of the fine-tuned ResNet101 model trained on dataset.}
     \label{fig:fresnet101_tsne}
@@
 
-Next, we provide some visual analysis of some correctly and
-incorrectly classified images using GradCAM. We provide two categories and two
-subcategories in our analysis. The two categories are correct and incorrect
-classification. Regardless of the feature activation map, the correctly
-classified images have the same predicted label as the ground truth, and
-incorrectly classified images have different predicted labels. The two
-subcategories are expected and non-expected feature activation. In the expected
-case, the features that are used to compute the classification are the same as
-the features used by the assistants in the sculpture identification process. In
-general, the features used by the assistants are the textures of the cuticle on
-the ant head. In the non-expected case, the features used to compute the
-classification are not from the head, for example, from the background,
-extraneous text, or the body of the ant. We used randomly selected images from
-the dataset and the fine-tuned ResNet101 model to perform the analysis. We show
-the GradCAM results in Figures \ref{fig:correct_images} and
-\ref{fig:incorrect_images}. The left image shows the preprocessed image input
-to the model. The right image shows the GradCAM output based on the
-classification. Four specimens were selected randomly from each category and
-subcategory.
+Next, we provide some visual analysis of some correctly and incorrectly
+classified images using GradCAM. In the expected case, the features that are
+used to compute the classification are the same as the features used by the
+assistants in the sculpture identification process. In general, the features
+used by the assistants are the textures of the cuticle on the ant head. In the
+non-expected case, the features used to compute the classification are not from
+the head, for example, from the background, extraneous text, or the body of the
+ant. We used randomly selected images from the dataset and the fine-tuned
+ResNet101 model to perform the analysis. We show the GradCAM results in Figures
+\ref{fig:correct_images} and \ref{fig:incorrect_images}. The left image shows
+the preprocessed image input to the model. The right image shows the GradCAM
+output based on the classification. Four specimens were selected randomly from
+each category and subcategory.
 
@@
 \newcommand{\subwidth}{0.35\textwidth}
-% \begin{figure}
-%     \centering
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
-%         \caption*{Smooth texture}
-%         \label{fig:correct_ideal_518}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
-%         \caption*{Smooth texture}
-%         \label{fig:correct_ideal_808}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
-%         \caption*{Rough texture}
-%         \label{fig:correct_ideal_842}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
-%         \caption*{Rough texture}
-%         \label{fig:correct_ideal_1091}
-%     \end{subfigure}
-%     \caption{Correctly classified images using expected features.}
-%     \label{fig:correct_ideal}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
-%         \caption*{Rough texture}
-%         \label{fig:correct_nonideal_346}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
-%         \caption*{Rough texture}
-%         \label{fig:correct_nonideal_1554}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
-%         \caption*{Rough texture}
-%         \label{fig:correct_nonideal_1694}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
-%         \caption*{Smooth texture}
-%         \label{fig:correct_nonideal_388}
-%     \end{subfigure}
-%     \caption{Correctly classified images using non-expected features.}
-%     \label{fig:correct_nonideal}
-% \end{figure}
-
 \begin{figure}
@@
     \end{subfigure}
-    \caption{Correctly classified images from fine-tuned ResNet101.}
+    \caption{GradCAM visualization for correctly classified images from
+        fine-tuned ResNet101. The red sections of the heatmap show which features
+        had higher activation for the classication.}
     \label{fig:correct_images}
 \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/84.png}
-%         \caption*{Rough texture}
-%         \label{fig:incorrect_ideal_84}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/86.png}
-%         \caption*{Smooth texture}
-%         \label{fig:incorrect_ideal_86}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/177.png}
-%         \caption*{Smooth texture}
-%         \label{fig:incorrect_ideal_177}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/197.png}
-%         \caption*{Smooth texture}
-%         \label{fig:incorrect_ideal_197}
-%     \end{subfigure}
-%     \caption{Incorrectly classified images using expected features.}
-%     \label{fig:incorrect_ideal}
-% \end{figure}
-
-% \begin{figure}
-%     \centering
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/1.png}
-%         \caption*{Rough texture}
-%         \label{fig:incorrect_nonideal_1}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/22.png}
-%         \caption*{Rough texture}
-%         \label{fig:incorrect_nonideal_22}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/61.png}
-%         \caption*{Rough texture}
-%         \label{fig:incorrect_nonideal_61}
-%     \end{subfigure}
-%     \begin{subfigure}{\subwidth}
-%         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/204.png}
-%         \caption*{Rough texture}
-%         \label{fig:incorrect_nonideal_204}
-%     \end{subfigure}
-%     \caption{Incorrectly classified images using non-expected features.}
-%     \label{fig:incorrect_nonideal}
-% \end{figure}
 
@@
     \end{subfigure}
-    \caption{Incorrectly classified images from fine-tuned ResNet101.}
+    \caption{GradCAM visualization for incorrectly classified images from
+        fine-tuned ResNet101. The red sections of the heatmap show which features
+        had higher activation for the classication.}
     \label{fig:incorrect_images}
@@
     \caption{Raw images are shown with corresponding segmented outputs from
-        K-views algorithm. The second row shows the results of using size of
-        15x15 patches. The third row shows the results of using size of 19x19
-        patches. In the images, the color red represents the background class,
-        green represents the smooth class, and the blue represent the rough
-        class.}
+        K-views algorithm. The second row shows the results of using size of 15
+        $\times$ 15 patches. The third row shows the results of using size of 19
+        $\times$ 19 patches. In the images, the color red represents the
+        background class, green represents the smooth class, and the blue
+        represent the rough class.}
     \label{fig:kviews_segmented}
@@
 \ref{fig:kviews_segmented}. A training set of 550 patches for background, 455
-patches for rough, and 493 patches for smooth are used with patch sizes of 15x15
-and 19x19. These patches are extracted from two rough and two smooth images with
-sizes of 608x608. The K-views algorithm converges early although we set the
-maximum number of iterations to be 1000. These experimental results are
-preliminary due to the minimal number of source images used and future work will
-explore a more expansive approach to the problem.
+patches for rough, and 493 patches for smooth are used with patch sizes of 15
+$\times$ 15 and 19 $\times$ 19. Hence, the three classes are utilized in the
+clustering. These patches are extracted from two rough and two smooth images
+with sizes of 608 $\times$ 608. The K-views algorithm converges early although
+we set the maximum number of iterations to be 1000. These experimental results
+are preliminary due to the minimal number of source images used and future work
+will explore a more expansive approach to the problem.
 
@@
 \section{Conclusion}
-Ant cuticle texture presumably has some function, but without the proper tools,
-evaluating the function based on thousands of species is infeasible. We have
-shown in this work that a deep learning approach and classical image texture
-analysis methods can be used to automatically categorize ants based on their
-cuticle texture, therefore supporting research on the evaluation of the function
-in future work. Our categorization system is novel in the field of automated
-insect identification due to the broad number of species captured by it.
-Additionally, a model that is pre-trained on a diverse image task such as ResNet
-can be transferred to our domain of texture analysis \footnote{All code is
-    publicly available on GitHub \url{https://github.com/ngngardner/cuticulus}}.
+We have shown in this work that a deep learning approach and classical image
+texture analysis methods can be used to automatically categorize ants based on
+their cuticle texture, therefore supporting research on the evaluation of the
+cuticle micro sculpturing function in future work. Our categorization system is
+novel in the field of automated insect identification due to the broad number of
+species captured by it. A model that is pre-trained on a diverse image task such
+as ResNet can be transferred to our domain of texture analysis\footnote{All code
+    is publicly available on GitHub \url{https://github.com/ngngardner/cuticulus}}.
 However, a deep learning algorithm created specifically for the domain of
-texture analysis had the best results in our experiments. In future work, we
+texture analysis had an F1 score of 92\% in our experiments. In future work, we
 will continue to explore texture classes present in the dataset which are not

### 2022-08-31 paper/main.tex
address comments

@@
 
-\subsection{Deep learning}
-
 Liu et al. gave a comprehensive survey on textural characterization in which
@@
 clustering approach \cite{hung_image_2019}.
+
+\subsection{Deep learning for texture analysis}
 
@@
 species. Marques et al. \cite{marques_ant_2018} propose an ensemble-based method
-to classify a large number of ant genera. They use ant images sourced from AntWeb
-in order to classify 44,806 specimens with at least 1 picture into 57 genera.
+to classify a large number of ant genera. They use ant head, profile, and dorsal
+images sourced from AntWeb in order to classify 44,806 specimens with at least
+one picture into 57 genera. Although the images used in their work are also from
+AntWeb, our goal is the classification of the ant cuticle micro sculpturing into
+broad groups based solely on the ant head images.
 
@@
 define two categories for them based on the appearance of the cuticle texture:
-\textit{rough} and \textit{smooth}. For each species used, there exists only one
-ant head image. We also refer to these ant head images, which describe the pose
-of the ant, as ant cuticule images, due to the clear cuticle texture present on
-the head. Some randomly selected images from each category are shown in Figures
+\textit{rough} and \textit{smooth}. The original AntWeb dataset was extended
+with these categories for each applicable ant specimen. The extended dataset is
+used for classifying the cuticle micro sculpturing into the broad groups. Due to
+the large number of ant species, it is quite difficult to obtain lab-quality
+specimen photos. For each species used, there exists only one ant head image. We
+also refer to these ant head images, which describe the pose of the ant, as ant
+cuticule images, due to the clear cuticle texture present on the head. Some
+randomly selected images from each category are shown in Figures
 \ref{fig:smooth-cuticle-texture} and \ref{fig:rough-cuticle-texture}.
@@
 
-Initially, the sculpture identification protocol had 5 subcategories of cuticle
-texture: smooth, punctate, striate, reticulate, and tuberous. For simplicity, we
-worked only with the two main categories, and the other 4 categories were
-combined to the single rough category. The training set identifications were
-reviewed together as a group by the assistants. Once training was complete,
-assistants were assigned the same genera of ants to identify independently each
-week. A weekly meeting was held to discuss identifications and assign new ones.
-These identifications were collected in the master spreadsheet and the
-identifications were assigned to individual ant species on a majority basis.
+Initially, the sculpture identification protocol had five subcategories of
+cuticle texture: smooth, punctate, striate, reticulate, and tuberous. For
+simplicity, we worked only with the two main categories, and the other four
+categories were combined to the single rough category. The training set
+identifications were reviewed together as a group by the assistants. Once
+training was complete, assistants were assigned the same genera of ants to
+identify independently each week. A weekly meeting was held to discuss
+identifications and assign new ones. These identifications were collected in the
+master spreadsheet and the identifications were assigned to individual ant
+species on a majority basis.
 
@@
 
-\section{K-views, K-means, and deep learning models for experiments}
+\section{Methods tested}
 
@@
 
-We evaluate the performance of the models using accuracy, precision, and F1
-score. We also apply Grad-CAM with manual inspection to visualize the activation
-weights for classified images to show which features lead to the
-classification result. Finally, we apply t-SNE to visualize the separation
-learned for the model to further analyze the classifications made by the model.
+We evaluate the performance of the models using accuracy, precision, F1 score
+and Matthew's correlation coefficient (MCC) \cite{chicco_matthews_2021}. We also
+apply Grad-CAM with manual inspection to visualize the activation weights for
+classified images to show which features lead to the classification result.
+Finally, we apply t-SNE to visualize the separation learned for the model to
+further analyze the classifications made by the model.
 
-\section{Experimental results and analysis}
+\section{Results}
 

### 2022-09-21 paper/main.tex
address minor concerns

@@
 contrast to the K-means, K-views looks at the neighboring pixels for providing
-the spatial relationship for classification.  Several variations of the basic
-K-views model have been proposed \cite{yang_image_2003, lan_improved_2010}.
+the spatial relationship for classification.
 
@@
 are multiples of $\pi/4$ radians starting from $0$ to $\pi$ radians. The Gabor
-filter lambda $\lambda$ parameter is equal to $\pi$.
+filter lambda $\lambda$ parameter is equal to $\pi$. In order to classify an
+image with K-views or K-means we count the number of pixels assigned to each
+cluster.
+
+In our experiments, we use $k=3$ clusters; one for background, and the other two
+as rough and smooth. To determine if an image is rough or smooth, we choose the
+cluster with the majority number of pixels. The number of pixels assigned to the
+background cluster are ignored for classification.
 
@@
 
-% \colorbox{green}{(TODO: DRP is best performing model, it should be included in
-%     visualization.)}
-
-% \subsection{Clustering segmentation}
-
-% \subsection{t-SNE Visualization}
-
 \begin{figure}[h]
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/518.png}
-        \caption*{Smooth texture}
         \label{fig:correct_ideal_518}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/808.png}
-        \caption*{Smooth texture}
         \label{fig:correct_ideal_808}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/842.png}
-        \caption*{Rough texture}
         \label{fig:correct_ideal_842}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_ideal/1091.png}
-        \caption*{Rough texture}
         \label{fig:correct_ideal_1091}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/346.png}
-        \caption*{Rough texture}
         \label{fig:correct_nonideal_346}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1554.png}
-        \caption*{Rough texture}
         \label{fig:correct_nonideal_1554}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/1694.png}
-        \caption*{Rough texture}
         \label{fig:correct_nonideal_1694}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/correct_nonideal/388.png}
-        \caption*{Smooth texture}
         \label{fig:correct_nonideal_388}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/84.png}
-        \caption*{Rough texture}
         \label{fig:incorrect_ideal_84}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/86.png}
-        \caption*{Smooth texture}
         \label{fig:incorrect_ideal_86}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/177.png}
-        \caption*{Smooth texture}
         \label{fig:incorrect_ideal_177}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_ideal/197.png}
-        \caption*{Smooth texture}
         \label{fig:incorrect_ideal_197}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/1.png}
-        \caption*{Rough texture}
         \label{fig:incorrect_nonideal_1}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/22.png}
-        \caption*{Rough texture}
         \label{fig:incorrect_nonideal_22}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/61.png}
-        \caption*{Rough texture}
         \label{fig:incorrect_nonideal_61}
@@
     \begin{subfigure}{\subwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{thesis_assets/gradcam/incorrect_nonideal/204.png}
-        \caption*{Rough texture}
         \label{fig:incorrect_nonideal_204}
@@
     \begin{subfigure}{\segmentedsubwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r103.png}
-        \caption*{Rough texture}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r111.png}
-        \caption*{Rough texture}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s105.png}
-        \caption*{Smooth texture}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s107.png}
-        \caption*{Smooth texture}
     \end{subfigure}
@@
 patches for rough, and 493 patches for smooth are used with patch sizes of 15
-$\times$ 15 and 19 $\times$ 19. Hence, the three classes are utilized in the
-clustering. These patches are extracted from two rough and two smooth images
-with sizes of 608 $\times$ 608. The K-views algorithm converges early although
-we set the maximum number of iterations to be 1000. These experimental results
-are preliminary due to the minimal number of source images used and future work
-will explore a more expansive approach to the problem.
+$\times$ 15 and 19 $\times$ 19. These patches are extracted from two rough and
+two smooth images with sizes of 608 $\times$ 608. The K-views algorithm
+converges early although we set the maximum number of iterations to be 1000.
+These experimental results are preliminary due to the minimal number of source
+images used and future work will explore a more expansive approach to the
+problem.
 

### 2022-09-21 paper/main.tex
proof version

--- 2022-08-31 paper/main.tex
@@
 
-\def\typeofarticle{Research Article} 
+\def\typeofarticle{Research article} 
 \def\currentvolume{2} 
@@
 \def\ppages{xx--xx} 
-\def\DOI{xxxx} 
-\def\Received{xx xx 2022} 
-\def\Revised{xx xx 2022}
-\def\Accepted{xx xx 2022} 
+\def\DOI{10.3934/aci.2022xxx} 
+\def\Received{22 July 2022} 
+\def\Revised{31 August 2022}
+\def\Accepted{01 September 2022} 
 \def\Published{xx 2022}
@@
     Haige Zhu\affil{1},
-    Zhiling Long\affil{4},
+    Zhiling Long\affil{4}, \\
     Min Wang\affil{3},
     Clint Penick\affil{2},
-    and Chih-Cheng Hung\affil{1}\corrauth
+    and Chih-Cheng Hung\affil{1,}\corrauth
 }%
@@
 }
-\corraddr{chung1@kennesaw.edu}
+\corraddr{Email: chung1@kennesaw.edu.}
 
@@
 scientists, and is expected to be a major contribution to the future of
-insect-based research \cite{martineau_survey_2017}. One of the most commonly
+insect-based research~\cite{martineau_survey_2017}. One of the most commonly
 used data types for insect analysis is image data. To develop an image-based
@@
 dataset. ImageNet is a widely used dataset of images labeled by experts with
-millions of images and thousands of categories \cite{deng_imagenet_2009}. In the
+millions of images and thousands of categories~\cite{deng_imagenet_2009}. In the
 ImageNet dataset, there are some categories that specify the class of the insect
@@
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 
+The results show that the fine-tuned ResNet models outperform the VGG and
+randomly initialized ResNet models on the task of ant head image classification.
+The DRP models perform better than the the general deep learning models VGG and
+ResNet on the task of ant head image classification. The fine-tuned DRP
+single-layer model with auxiliary classifier performed the best with an average
+F1 score of 0.92. We further analyze the separation learned by both ResNet101
+models in the following section.
+
 % \subsection{Experimental results}
@@
     \caption{Experimental results of different algorithms}
+    \small
     \input{assets/results.tex}
@@
 
-\FloatBarrier
-
-The results show that the fine-tuned ResNet models outperform the VGG and
-randomly initialized ResNet models on the task of ant head image classification.
-The DRP models perform better than the the general deep learning models VGG and
-ResNet on the task of ant head image classification. The fine-tuned DRP
-single-layer model with auxiliary classifier performed the best with an average
-F1 score of 0.92. We further analyze the separation learned by both ResNet101
-models in the following section.
+%\FloatBarrier
 
@@
 
-\begin{figure}[h]
+\begin{figure}[h!]
     \centering
@@
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r103.png}
+        \includegraphics[width=1\linewidth]{figs/r103.png}
         \caption*{Rough texture}
@@
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/r111.png}
+        \includegraphics[width=1\linewidth]{figs/r111.png}
         \caption*{Rough texture}
@@
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s105.png}
+        \includegraphics[width=1\linewidth]{figs/s105.png}
         \caption*{Smooth texture}
@@
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/s107.png}
+        \includegraphics[width=1\linewidth]{figs/s107.png}
         \caption*{Smooth texture}
@@
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/r103.png}
+        \includegraphics[width=1\linewidth]{figs/15/r103.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/r111.png}
+        \includegraphics[width=1\linewidth]{figs/15/r111.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/s105.png}
+        \includegraphics[width=1\linewidth]{figs/15/s105.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/15/s107.png}
+        \includegraphics[width=1\linewidth]{figs/15/s107.png}
     \end{subfigure}
@@
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/r103.png}
+        \includegraphics[width=1\linewidth]{figs/19/r103.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/r111.png}
+        \includegraphics[width=1\linewidth]{figs/19/r111.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/s105.png}
+        \includegraphics[width=1\linewidth]{figs/19/s105.png}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
-        \includegraphics[width=1\linewidth]{/home/ngardn10/projects/cuticulus/paper/assets/images/kviews/19/s107.png}
+        \includegraphics[width=1\linewidth]{figs/19/s107.png}
     \end{subfigure}
@@
 with sizes of 608 $\times$ 608. The K-views algorithm converges early although
-we set the maximum number of iterations to be 1000. These experimental results
-are preliminary due to the minimal number of source images used and future work
-will explore a more expansive approach to the problem.
+we set the maximum number of iterations to be 1000. In order to classify an
+image with K-views, the input image is split into patches. Then, each patch is
+assigned to a cluster based on the distance to the cluster. If there are more
+`smooth' patches, than `rough' patches, then the image will be classified as
+rough. These experimental results are preliminary due to the minimal number of
+source images used and future work will explore a more expansive approach to the
+problem.
 
-\FloatBarrier
-\section{Conclusion}
+\section{Conclusions}
 We have shown in this work that a deep learning approach and classical image
@@
 
+\section*{Conflict of interest}
+All authors declare no conflicts of interest in this paper.
+
 \bibliographystyle{AIMS}

### 2022-09-21 paper/main.tex
Proof (#22)
* proof version

* fix paragraph

@@
 patches for rough, and 493 patches for smooth are used with patch sizes of 15
-$\times$ 15 and 19 $\times$ 19. Hence, the three classes are utilized in the
-clustering. These patches are extracted from two rough and two smooth images
-with sizes of 608 $\times$ 608. The K-views algorithm converges early although
-we set the maximum number of iterations to be 1000. In order to classify an
-image with K-views, the input image is split into patches. Then, each patch is
-assigned to a cluster based on the distance to the cluster. If there are more
-`smooth' patches, than `rough' patches, then the image will be classified as
-rough. These experimental results are preliminary due to the minimal number of
-source images used and future work will explore a more expansive approach to the
+$\times$ 15 and 19 $\times$ 19. These patches are extracted from two rough and
+two smooth images with sizes of 608 $\times$ 608. The K-views algorithm
+converges early although we set the maximum number of iterations to be 1000.
+These experimental results are preliminary due to the minimal number of source
+images used and future work will explore a more expansive approach to the
 problem.

### 2022-09-21 paper/main.tex
address minor concerns (#23)

@@
     \begin{subfigure}{\segmentedsubwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{figs/r103.png}
-        \caption*{Rough texture}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
+        \caption*{Rough texture}
         \includegraphics[width=1\linewidth]{figs/r111.png}
-        \caption*{Rough texture}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{figs/s105.png}
-        \caption*{Smooth texture}
     \end{subfigure}
     \begin{subfigure}{\segmentedsubwidth}
+        \caption*{Smooth texture}
         \includegraphics[width=1\linewidth]{figs/s107.png}
-        \caption*{Smooth texture}
     \end{subfigure}

### 2022-09-21 paper/main.tex
change k-views paragraph

@@
 
-For the feasibility study in this ant image datasets, we tested the statistical
-clusterings methods and deep learning methods separately. Both K-means and
+For the feasibility study in these ant image datasets, we tested the statistical
+clustering methods and deep learning methods separately. Both K-means and
 K-views were used for statistical clustering methods. The experiments may
 indicate how well the statistical approach and deep learning networks will do in
-this type of textural images. In pre-processing the dataset for the statistical
-clustering methods, we apply the Gabor filter to transform the gray-scale ant
-image into four Gabor-filtered channels, subtract the gray-scale image from each
-channel, and take the absolute value of the resulting images. This
-pre-processing will highlight some of characteristic features in textures, such
-as edge-like in the image. This step is very useful in feeding the datasets to
-the K-views algorithms. The theta $\theta$ parameters in the Gabor filters used
-are multiples of $\pi/4$ radians starting from $0$ to $\pi$ radians. The Gabor
-filter lambda $\lambda$ parameter is equal to $\pi$. In order to classify an
-image with K-views or K-means we count the number of pixels assigned to each
-cluster.
+this type of textural images. A pre-processing step of the dataset for the
+statistical clustering methods is shown below. This pre-processing will
+highlight some of characteristic features in textures, such as edge-like in the
+image. This step is very useful in feeding the datasets to the K-views
+algorithms. In order to classify an image with K-views or K-means in which we
+count the number of pixels assigned to each cluster. A brief procedure for
+K-views experiments is given below for training and classification.
+\begin{description}
+    \item[Step 1:] Pre-processing the dataset: apply the Gabor filter to
+          transform the gray-scale ant image into four Gabor-filtered channels,
+          subtract the gray-scale image from each channel, and take the absolute
+          value of the resulting images. The theta θ parameters in the Gabor
+          filters used are multiples of π/4 radians starting from 0 to π
+          radians. The Gabor filter lambda λ parameter is equal to π.
 
-In our experiments, we use $k=3$ clusters; one for background, and the other two
-as rough and smooth. To determine if an image is rough or smooth, we choose the
-cluster with the majority number of pixels. The number of pixels assigned to the
+    \item[Step 2:] Initialize k = 3 clusters; one for background, and the other
+          two as rough and smooth.
+
+    \item[Step 3:] Extract patches for each cluster for the clustering.
+
+    \item[Step 4:] Calculate the mean of each cluster once the training
+          converges.
+
+    \item[Step 5:] Use the mean of the cluster for the classification. To
+          determine if an image is rough or smooth, we choose the cluster with
+          the majority number of pixels. The number of pixels assigned to the
           background cluster are ignored for classification.
+\end{description}
 
+\noindent Steps 2 to 4 are for training and step 5 is for the classification.
+For the K-means experiments, it will be the same as the K-views except no
+patches are used.
+\newline
 % \subsection{Deep learning models used for the experiments}

### 2022-09-22 paper/main.tex
final touches
add initials

@@
     Min Wang\affil{3},
-    Clint Penick\affil{2},
+    Clint A. Penick\affil{2},
     and Chih-Cheng Hung\affil{1,}\corrauth
@@
 highlight some of characteristic features in textures, such as edge-like in the
-image. This step is very useful in feeding the datasets to the K-views
-algorithms. In order to classify an image with K-views or K-means in which we
-count the number of pixels assigned to each cluster. A brief procedure for
-K-views experiments is given below for training and classification.
+image. This step is useful in feeding the datasets to the K-views algorithms. In
+order to classify an image with K-views or K-means in which we count the number
+of pixels assigned to each cluster. A brief procedure for K-views experiments is
+given below for training and classification.
 \begin{description}

### 2022-09-22 paper/main.tex
fix math symbols

@@
           subtract the gray-scale image from each channel, and take the absolute
-          value of the resulting images. The theta θ parameters in the Gabor
-          filters used are multiples of π/4 radians starting from 0 to π
-          radians. The Gabor filter lambda λ parameter is equal to π.
+          value of the resulting images. The theta $\theta$ parameters in the
+          Gabor filters used are multiples of $\pi/4$ radians starting from 0 to
+          $\pi$ radians. The Gabor filter lambda $\lambda$ parameter is equal to
+          $\pi$.
 
-    \item[Step 2:] Initialize k = 3 clusters; one for background, and the other
-          two as rough and smooth.
+    \item[Step 2:] Initialize $k = 3$ clusters; one for background, and the
+          other two as rough and smooth.
 

### 2022-09-23 cuticulus/core/datasets/downloader.py
refactor downloader,

@@ -1,14 +1,10 @@
 """Download raw dataset."""
-
 import logging
 from pathlib import Path
-from zipfile import PyZipFile
 
-import gdown
-import questionary
+from juliacall import Main as jl
 
 from cuticulus import const
-from cuticulus.console import console
 
 log = logging.getLogger('rich')
 
@@ -17,48 +13,33 @@ class Downloader(object):
     """Download raw dataset."""
 
     def __init__(self):
-        """Initialize Download class."""
-        self.raw = Path.cwd() / 'dataset.zip'
-        self.base_path = Path.cwd() / 'dataset'
-        self.download()
-
-    def download_zip(self):
-        """Download raw dataset .zip file."""
-        if not self.raw.exists():
-            console.log('Dataset .zip not found.')
-
-            prompt = 'Okay to download dataset to: "{0}"?'.format(self.raw)
-            if questionary.confirm(prompt).ask():
-                log.info('Downloading dataset...')
-                gdown.download(const.DATASET_URL, str(self.raw))
-                log.info('Downloaded dataset!')
-            else:
-                console.log('Download cancelled.')
-
-    def unzip_dataset(self):
-        """Unzip dataset to the proper path."""
-        if not self.base_path.exists():
-            console.log('Dataset directory not found.')
-
-            prompt = 'Okay to unzip dataset to: "{0}"?'.format(self.base_path)
-            if questionary.confirm(prompt).ask():
-                log.info('Extracting dataset...')
-                PyZipFile(self.raw).extractall(str(self.base_path))
-                log.info('Extracted dataset!')
-            else:
-                console.log('Extraction cancelled.')
-
-    def download(self):
-        """Download dataset.
-
-        Raises:
-            FileNotFoundError: cannot find zip or dataset
+        """Initialize downloader."""
+        name = 'cuticulus-dataset-raw'
+        description = """
+        Dataset: Cuticle texture ant head image dataset (raw)
+        License: MIT
+        Website: {website}
         """
-        self.download_zip()
-        self.unzip_dataset()
+        description = description.format(website=const.DATASET_URL)
 
-        if not self.raw.exists():
-            raise FileNotFoundError
+        register = """
+        register(DataDep(
+            "{name}",
+            "{description}",
+            "{url}",
+            "{hash}",
+            post_fetch_method = unpack,
+        ))
+        """
+        register = register.format(
+            name=name,
+            description=description,
+            url=const.DATASET_URL,
[…]

### 2022-09-23 cuticulus/__main__.py
script generates file 'dataset.zip'

@@ -1,24 +1,36 @@
 """Main entrypoint."""
+import shutil
+from pathlib import Path
 
-from cuticulus.core.collages import build_collage
-from cuticulus.datasets import RoughSmoothSub
+from dotenv import load_dotenv
+
+from cuticulus.datasets import RoughSmoothFull
 
 
 def main():
     """Run main function."""
-    patch_size = (32, 32)
-    ds = RoughSmoothSub(
-        size=patch_size,
-        rebuild=True,
+    # load environment variables
+    if not Path('.env').exists():
+        shutil.copy('.env.example', '.env')
+    load_dotenv()
+
+    # build the dataset
+    image_size = (256, 256)
+    ds = RoughSmoothFull(
+        size=image_size,
     )
 
-    build_collage(
-        ds.labels,
-        ds.images,
-        name='rs',
-        rows=10,
-        cols=10,
-    )
+    # create output directory
+    output_path = Path('dataset')
+    if output_path.exists():
+        shutil.rmtree(output_path)
+    output_path.mkdir(exist_ok=True)
+    shutil.copytree(ds.dir_path, output_path / ds.dir_path.name)
+
+    # create output zip file
+    if Path('{0}.zip').exists():
+        shutil.rmtree(Path('{0}.zip'))
+    shutil.make_archive(output_path, 'zip', output_path)
 
 
 if __name__ == '__main__':

### 2022-09-23 cuticulus/const.py
Refactor (#24)
* drop nix

* docs: installation

* update deps, add julia package

* refactor downloader,

* script generates file 'dataset.zip'

* github action

@@ -1,7 +1,7 @@
 """Global constants module."""
 
-GDRIVE_URL = 'https://drive.google.com'
-DATASET_URL = '{0}/uc?id=1xABlWE790uWmT0mMxbDjn9KZlNUB6Y57'.format(GDRIVE_URL)
+DATASET_URL = 'http://cuticulus-dataset.storage.googleapis.com/dataset.zip'
+DATASET_HASH = 'dd16dc2f12c25c830acbe57001f8cad5e9d57bd26283412f7469be93527e5b1f'  # noqa
 
 # dataset
 LABEL_COLS = ('Jp', 'Becca', 'Katy')

### 2022-09-23 cuticulus/__main__.py
instantiate pkg

@@ -3,8 +3,11 @@ import shutil
 from pathlib import Path
 
 from dotenv import load_dotenv
+from juliacall import Main as jl
 
 from cuticulus.datasets import RoughSmoothFull
+
+jl.seval('using Pkg; Pkg.instantiate(); Pkg.resolve()')
 
 
 def main():

### 2022-09-23 cuticulus/__main__.py
remove sanity check

@@ -3,11 +3,8 @@ import shutil
 from pathlib import Path
 
 from dotenv import load_dotenv
-from juliacall import Main as jl
 
 from cuticulus.datasets import RoughSmoothFull
-
-jl.seval('using Pkg; Pkg.instantiate(); Pkg.resolve()')
 
 
 def main():
